---
title: "Calibration plot"
author: "Grace Lyden"
date: "2025-06-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
# base::load("mortality-model.RData")
base::load("mortality_model_test_data.RData")
base::load("eight-sets-valid.Rdata")

library(dplyr)
library(haven)
library(survival)
library(ggplot2)
library(splines)
library(lubridate)
library(tidyr)
```

For calibration, we can use time-varying weights by week, similar to the mortality model. First, we refit the IPW models for the probability of transplant in the test set, to account for the test-set censoring distribution. Then, we expand the validation/test set of calendar cross-sections (which are one row per PX_ID per cross-section) to include up to 6 rows for the 6 weeks follow-up, with time-varying weights per row.

## Refit transplant models for IPW
Fit the numerator and denominator models for the IPW in the test set. These are based on probabilities of receiving a transplant.
```{r}
## create probability for numerator of stabilized weights ##
# we want Pr(transplant) given variables at start of 6-week window
# for simplicity (and to avoid correlated data), we will model this in non-overlapping 6-week windows of test_data

# create 6-week windows of test_data
test_data = test_data %>%
  group_by(PX_ID) %>%
  mutate(window=ceiling(row_number()/6),
         week_from_baseline = (row_number() - 6*(window-1))) %>%
  ungroup()
# test_data %>% select(PX_ID,interval_start,interval_stop,transplant,window,week_from_baseline) %>% head(20)
test_data = test_data %>%
  group_by(PX_ID,window) %>%
  mutate(albumin_at_start = first(albumin),
         bilirubin_at_start = first(bilirubin),
         eGFR_at_start = first(eGFR),
         sodium_at_start = first(sodium),
         BNP_at_start = first(BNP),
         BNP_NT_Pro_at_start = first(BNP_NT_Pro),
         durable_LVAD_at_start = first(durable_LVAD),
         short_mcs_ever_at_start = first(short_MCS_ever)) %>%
  ungroup()

# fit transplant model
## just US-CRS variables bc anything in this numerator model has to be in the mortality model
tx_model_first = glm(
   transplant ~ 
    ns(albumin_at_start, df = 3) + ns(bilirubin_at_start, df = 3) + ns(eGFR_at_start, df = 3) + ns(sodium_at_start, df = 3) + ns(BNP_at_start, df = 3)*BNP_NT_Pro_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + factor(week_from_baseline), data = test_data, family = binomial)

# remove new variables from test_data
test_data = test_data %>% select(-(window:short_mcs_ever_at_start))

## create probability for denominator of stabilized weights ##
# need prob_no_tx variable in test_data

# fit chosen model
tx_status_hemo_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + 
     ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + 
     status + ns(cpo, df = 3) + ns(api, df = 3) + ns(papi, df = 3)), 
   data = test_data, family = binomial)
test_data$fit_prob_hemo <- predict(tx_status_hemo_model, newdata=test_data, type = "response")

# with chosen model, set probability of not receiving a transplant each week
test_data = test_data %>%
  mutate(prob_no_tx = 1 - fit_prob_hemo)
```

## Set up test data with rows after delisting, by week
```{r test}
## Create week variable
test_data = test_data %>% 
  group_by(PX_ID) %>%
  mutate(
    week = row_number()
  ) %>% relocate(week, .after = CAN_LISTING_CTR_CD) %>% ungroup


# Adjust delisting time variables to censor at 6 weeks
# Depending on whether the week is even or odd, this translates to 4 or 5 weeks of follow up after delisting
# If candidates dies more than 4/5 weeks after delisting, they are censored
test_data = test_data %>% group_by(PX_ID) %>% 
  
  mutate(
    # if we have full follow-up for the week, follow for censoring (floor)
    time_until_censor_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_censor_after_delisting) ~ 
                  floor(pmax(0,removal_time + time_until_censor_after_delisting - interval_stop)/7)*7,
                T ~ NA),
    # follow for death even w/o full week (ceiling)
    time_until_death_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_death_after_delisting) ~ 
                  ceiling(pmax(0,removal_time + time_until_death_after_delisting - interval_stop)/7)*7,
                T ~ NA)) %>%
  
  fill(time_until_censor_after_delisting,time_until_death_after_delisting,.direction="downup") %>%
  
  ## Note: This is intentionally different from mortality_model_data - 5 weeks regardless of even/odd
  # censor at 6 weeks
  ungroup() %>%
  mutate(time_until_censor_after_delisting = case_when(
    time_until_censor_after_delisting > 35  |
      time_until_death_after_delisting > 35 ~ 35,
    TRUE ~ time_until_censor_after_delisting
  ),
  # recode time to death as missing if censored
  time_until_death_after_delisting = ifelse(!is.na(time_until_censor_after_delisting),NA,
                                            time_until_death_after_delisting),
  outcome = ifelse(!is.na(time_until_censor_after_delisting),0,outcome)
  
  )

## Determine delisting row and # of duplicates we need after delisting, until death or censoring
test_data = test_data %>% group_by(PX_ID) %>% mutate(
  delisting = case_when(
    (!is.na(time_until_death_after_delisting) | !is.na(time_until_censor_after_delisting)) & row_number() == n() ~ 1,
    TRUE ~ 0)
) %>% ungroup() %>% mutate(
  duplicates_after_delisting = case_when(
    delisting == 1 & time_until_death_after_delisting > 0 ~
      time_until_death_after_delisting/7,
    delisting == 1 & time_until_censor_after_delisting > 0 ~ 
      time_until_censor_after_delisting/7,
    TRUE ~ 0))

## repeat rows for the duplicates after delisting
# and recode interval_start, interval_stop
test_data = test_data %>%
  slice(rep(row_number(), duplicates_after_delisting+1)) %>%
  group_by(PX_ID) %>%
  mutate(week=row_number()) %>%
  group_by(PX_ID,duplicates_after_delisting) %>%
  mutate(dup=row_number()-1) %>%
  ungroup() %>%
  mutate(interval_start=ifelse(duplicates_after_delisting>0,interval_start+7*dup,interval_start),
         current_date = as.Date(can_listing_dt) + interval_start,
         interval_stop=ifelse(duplicates_after_delisting>0,interval_stop+7*dup,interval_stop)) %>%
  select(-dup)

# code outcome for last row - now outcome is correct
test_data = test_data %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    row_number()==n() & !is.na(time_until_death_after_delisting) & is.na(time_until_censor_after_delisting) ~ 1,
    T ~ 0))
table(test_data$outcome)
table(test_data$outcome,test_data$CAN_REM_CD)

## Adjust prob_no_tx variable after delisting
# set prob_no_tx to 1 after delisting
test_data = test_data %>%
  group_by(PX_ID) %>%
  mutate(delisting2 = cumsum(delisting),
         prob_no_tx = ifelse(delisting2>1,1,prob_no_tx)) %>%
  ungroup
table(test_data$transplant,test_data$delisting) # quick check
```

## Set up validation data for calibration analysis

Join `test_data` to the eight calendar cross-sections and include 6 weeks of follow-up per PX_ID per cross-section. Calculate IPW.
```{r join}
# first, combine the eight_sets_valid datasets - save predicted probs from different models as variables
## not saving int; just going to pool all 8 sets for calibration analysis
eight_sets_valid = full_join(eight_sets_valid_modA %>% select(PX_ID,current_date,prob_outcome_6wk_modA=prob_outcome_6wk),
                       eight_sets_valid_modB %>% select(PX_ID,current_date,prob_outcome_6wk_modB=prob_outcome_6wk)) %>%
  full_join(.,eight_sets_valid_modC %>% select(PX_ID,current_date,prob_outcome_6wk_modC=prob_outcome_6wk)) %>%
  full_join(.,eight_sets_valid_modD %>% select(PX_ID,current_date,prob_outcome_6wk_modD=prob_outcome_6wk)) %>%
  full_join(.,eight_sets_valid_modE %>% select(PX_ID,current_date,prob_outcome_6wk_modE=prob_outcome_6wk)) %>%
  full_join(.,eight_sets_valid_modF %>% select(PX_ID,current_date,prob_outcome_6wk_modF=prob_outcome_6wk)) %>%
  full_join(.,eight_sets_valid_modG %>% select(PX_ID,current_date,prob_outcome_6wk_modG=prob_outcome_6wk)) %>%
  full_join(.,eight_sets_valid_modH %>% select(PX_ID,current_date,prob_outcome_6wk_modH=prob_outcome_6wk))
names(eight_sets_valid)

# confirm eight_sets and test_data are both unique by PX_ID and current_date
nrow(eight_sets_valid);eight_sets_valid %>% distinct(PX_ID,current_date) %>% nrow
nrow(test_data);test_data %>% distinct(PX_ID,current_date) %>% nrow

# function to extend eight_sets to include 6 rows follow-up per PX_ID (per cross-section) and calculate IPW (sweight)
## eight_sets input should be a very limited dataset with PX_ID, current_date, and variables not in test_data (predicted probabilities of 6-week survival)
extend_6wk = function(eight_sets) {
  # left join to test_data
  joindat = left_join(test_data,
                      eight_sets %>%
                        mutate(include=T),
                      by=c("PX_ID","current_date"))

  # create 6 weeks of follow-up for each row included in eight_sets
  joindat = joindat %>%
    mutate(int_start = case_when(include ~ 1, T ~ 0)) %>%
    group_by(PX_ID) %>%
    mutate(
      grp = cumsum(int_start)
    ) %>%
    group_by(PX_ID,grp) %>%
    mutate(week_from_baseline = row_number()) %>%
    ungroup() %>%
    filter(week_from_baseline<=6,grp>0)
  # joindat %>% select(PX_ID,grp,week_from_baseline,current_date,interval_start,interval_stop,outcome)
  # table(joindat$include)
  # nrow(eight_sets)
  # summary(joindat$current_date[joindat$week_from_baseline==1])
  # table(joindat$outcome)
  # table(eight_sets$outcome)
  
  ## Calculate (unstabilized) weights
  # Create weight variable based on prob_no_tx
  joindat = joindat %>% group_by(PX_ID, grp) %>% mutate(
    weight = 1/(cumprod(prob_no_tx))
  ) %>% ungroup()
  
  # remove transplant weeks
  joindat = joindat %>% filter(transplant == 0)
  
  ## look at weights
  # with(joindat,tapply(weight,week_from_baseline,summary))
  
  ## calculate stabilized weights
  
  # get variables at start
  joindat = joindat %>% 
    group_by(PX_ID,grp) %>% 
    mutate(albumin_at_start = first(albumin),
           bilirubin_at_start = first(bilirubin),
           eGFR_at_start = first(eGFR),
           sodium_at_start = first(sodium),
           BNP_at_start = first(BNP),
           BNP_NT_Pro_at_start = first(BNP_NT_Pro),
           durable_LVAD_at_start = first(durable_LVAD),
           short_mcs_ever_at_start = first(short_MCS_ever),
           status_at_start=first(status),
           excep_at_start=first(Exception)) %>% 
    ungroup %>%
    mutate(excep_at_start=ifelse(is.na(excep_at_start),0,excep_at_start))
  
  # calculate numerator
  joindat = joindat %>% mutate(
    prob_no_tx_start = 1 - 
      predict(tx_model_first, newdata = joindat, type = "response"),
    prob_no_tx_start = ifelse(delisting2>1,1,prob_no_tx_start)) %>%
    group_by(PX_ID, grp) %>% mutate(
    sweight = cumprod(prob_no_tx_start)/(cumprod(prob_no_tx))
  ) %>% ungroup
  # with(joindat,tapply(sweight,week_from_baseline,summary))
  
  # # truncate weights
  # joindat = joindat %>% group_by(week_from_baseline) %>% mutate(
  #   tweight = ifelse(sweight > quantile(sweight,probs=.98,na.rm=T), 
  #                    quantile(sweight,probs=.98,na.rm=T), sweight),
  #   tweight = ifelse(sweight < quantile(sweight,probs=.02,na.rm=T), 
  #                    quantile(sweight,probs=.02,na.rm=T), tweight)) %>%
  #   ungroup
  # with(joindat,tapply(tweight,week_from_baseline,summary))
  
  return(joindat)
}
eight_sets_valid_long = extend_6wk(eight_sets_valid)
nrow(eight_sets_valid);nrow(eight_sets_valid_long)
```

## Calculate calibration statistic in validation data
Average squared difference between observed 6-week mortality within status and average predicted 6-week mortality within status, across medical urgency statuses in validation data (9/1/23-8/31/24).

```{r}
makeCalibData = function(data,q=10,risk_score="prob_outcome_6wk") {
  # group data into ntiles by expected risk
  data1 = data %>% filter(week_from_baseline==1)
  data1$risk_grp = ntile(data1[[risk_score]],n = q)
  data = left_join(data,data1 %>% select(PX_ID,grp,risk_grp),by=c("PX_ID","grp"))
  
  # get average predicted risk in each ntile
  df = data1 %>% group_by(risk_grp) %>% 
    summarise(exp=mean(!!sym(risk_score)),
              exp.25=quantile(!!sym(risk_score),.25),
              exp.75=quantile(!!sym(risk_score),.75))
  
  # get observed risk in each ntile
  obs_risks = lapply(1:q,function(i) {
    ipw.surv = survfit(Surv(start,stop,outcome) ~ 1,
                      data=data %>% filter(risk_grp==i) %>%
                        mutate(start=week_from_baseline-1,
                               stop=week_from_baseline,
                               id=interaction(PX_ID,grp)),
                      type='kaplan-meier', weights=sweight,
                      cluster=id)
    ipw.CI = summary(ipw.surv,times=6,extend=T)[c("surv","lower","upper")]
    # convert to mortality
    mort.CI = lapply(ipw.CI,function(surv) {1-surv})
    return(mort.CI)
  })
  obs_df = do.call(rbind.data.frame,obs_risks) %>%
    rename(obs=surv,upper=lower,lower=upper)
  output_df = cbind.data.frame(df,obs_df)
  return(output_df)
}

# function to calculate calibration statistic by status
## joindat should be the extended dataset by week
## risk_score is the name of the predicted 6-week mortality
calibStatStatus = function(joindat,risk_score) {
  # get overall observed and expected by status
  calib_status = lapply(1:6,function(i) {
    makeCalibData(joindat %>% filter(status_at_start==i),q=1,risk_score) %>%
      mutate(status=i) %>% relocate(status,.before=exp) %>% select(-risk_grp)
  })
  
  # # get overall observed and expected by status and exception
  # levels = joindat %>% count(status_at_start,excep_at_start)
  # calib_status = lapply(1:nrow(levels),
  #                       function(xx) {
  #   i = levels$status_at_start[xx]
  #   j = levels$excep_at_start[xx]
  #   makeCalibData(joindat %>% filter(status_at_start==i,excep_at_start==j),q=1,risk_score) %>%
  #     mutate(status=i,exception=j) %>% relocate(status,exception,.before=exp) %>% select(-risk_grp)
  # })
  calib_status = do.call(rbind.data.frame,calib_status)
  # return(calib_status %>% mutate(sqdiff=(100*(obs-exp))^2,diff=100*(obs-exp)) %>% select(status,obs,exp,sqdiff,diff))
  
  # calculate average squared difference bw obs and exp across statuses
  ## do this in percent (*100) for bigger numbers
  return(with(calib_status,mean((100*(obs-exp))^2)))
}
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modA")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modB")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modC")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modD")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modE")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modF")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modG")
calibStatStatus(eight_sets_valid_long,"prob_outcome_6wk_modH")


```





# Working

## Make calibration plot
```{r calib}

calibdat = makeCalibData(joindat,20)
# calibdat1.0 = makeCalibData(joindat,20,"uscrs")

library(ggforce)
ggplot(calibdat,aes(x=exp,y=obs)) +
  # geom_smooth(method = "lm", se = FALSE, linetype="dashed", fullrange=T,
  #             linewidth=.5,alpha=.75) +
  geom_abline(intercept=0,slope=1,color="darkgrey") +
  geom_errorbar(aes(ymin=lower,ymax=upper)) +
  geom_point(size=2,alpha=.8) + 
  scale_x_continuous(limits=c(0,.15)) +
  scale_y_continuous(limits=c(0,.15),
                     oob = scales::rescale_none) +
  # coord_cartesian(xlim=c(0,.15),ylim=c(0,.15)) +
  facet_zoom(xlim=c(0,.05),ylim=c(0,.05),zoom.size=1,show.area=F) +  
  theme_bw() +
  theme(text=element_text(size=14)) +
  labs(y="Observed 6-week mortality probability",x="Predicted 6-week mortality probability (US-CRS 2.0)")
ggsave("calib-overall-NEW.png",width=8,height=4)

# by status
joindat = joindat %>% group_by(PX_ID,grp) %>% mutate(status_at_start=first(status)) %>% ungroup
table(joindat$status_at_start)
table(joindat$status)

# get overall observed and expected by status
calib_status = lapply(1:6,function(i) {
  makeCalibData(joindat %>% filter(status_at_start==i),q=1) %>%
    mutate(status=i) %>% relocate(status,.before=exp) %>% select(-risk_grp)
})
calib_status = do.call(rbind.data.frame,calib_status)

# make bar plot by status
calib_status_long = calib_status %>%
  pivot_longer(cols = c(exp, obs), names_to = "type", values_to = "value") %>%
  mutate(lower=ifelse(type=="exp",exp.05,lower),
         upper=ifelse(type=="exp",exp.95,upper),
         type_fact = factor(type,levels=c("exp","obs"),
                            labels=c("Predicted\n(US-CRS 2.0)",
                                     "Observed")))
ggplot(calib_status_long,
       aes(x = factor(status), y = value, fill = type_fact)) +
  geom_col(position = position_dodge(width = 0.9), color = "black") +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  coord_cartesian(ylim=c(0,.4)) +
  scale_fill_manual(values = c("Predicted\n(US-CRS 2.0)" = "white", "Observed" = "gray40")) +
  labs(x = "Status", y = "6-Week Mortality Probability", fill = NULL) +
  theme_bw() + theme(text=element_text(size=14))
ggsave("calib-by-status-NEW.png",width=7.5,height=6.5)
```


## Calibration by status and exception
```{r excepplot}
# get exception at start
joindat = joindat %>% group_by(PX_ID,grp) %>% mutate(excep_at_start=first(Exception),excep_at_start=ifelse(is.na(excep_at_start),0,excep_at_start)) %>% ungroup
joindat %>% count(status_at_start,excep_at_start)

# get overall observed and expected by status
levels = joindat %>% count(status_at_start,excep_at_start)
calib_status = lapply(1:nrow(levels),
                      function(xx) {
  i = levels$status_at_start[xx]
  j = levels$excep_at_start[xx]
  makeCalibData(joindat %>% filter(status_at_start==i,excep_at_start==j),q=1) %>%
    mutate(status=i,exception=j) %>% relocate(status,exception,.before=exp) %>% select(-risk_grp)
})
calib_status = do.call(rbind.data.frame,calib_status)

# make bar plot by status
calib_status_long = calib_status %>%
  pivot_longer(cols = c(exp, obs), names_to = "type", values_to = "value") %>%
  mutate(lower=ifelse(type=="exp",exp.05,lower),
         upper=ifelse(type=="exp",exp.95,upper),
         type_fact = factor(type,levels=c("exp","obs"),
                            labels=c("Predicted\n(US-CRS 2.0)",
                                     "Observed")))
ggplot(calib_status_long,
       aes(x = factor(paste(status,exception,sep=".")), y = value, fill = type_fact)) +
  geom_col(position = position_dodge(width = 0.9), color = "black") +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  coord_cartesian(ylim=c(0,.4)) +
  scale_fill_manual(values = c("Predicted\n(US-CRS 2.0)" = "white", "Observed" = "gray40")) +
  labs(x = "Status.Exception (0=standard, 1=exception)", y = "6-Week Mortality Probability", fill = NULL) +
  theme_bw() + theme(text=element_text(size=14))
ggsave("calib-by-excep-NEW.png",width=9.5,height=6.5)
```
