---
title: "Create 1-week discrete-time dataset"
output: html_document
date: "2024-06-06"
---

## Load data and necessary libraries
## Dataset can be constructed from the heart_pipeline_final.Rmd file
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
# Molly:
load("C:/Users/mhwhi/Downloads/heart_post_policy.RData")
df_cand_thor <- haven::read_sas('C:/Users/mhwhi/Downloads/pubsaf2403/cand_thor.sas7bdat', col_select=c(PX_ID, CAN_GENDER, CAN_RACE, CAN_HGT_CM, CAN_WGT_KG, CAN_ECMO))

# Grace:
# load("heart_post_policy.RData")
# groups(df) # PX_ID grouping
# curr_data_release = "//srtr-fs-p01/SAF/External/Pub_SAF/saf2403_Q1/pubsaf2403"
# df_cand_thor <- haven::read_sas(file.path(curr_data_release,'cand_thor.sas7bdat'),
#                                 col_select=c(PX_ID, CAN_GENDER, CAN_RACE, CAN_HGT_CM, CAN_WGT_KG, CAN_ECMO))

library(renv)
library(dplyr)
library(tidyr)
library(lubridate)
library(table1)
library(purrr)
library(ggplot2)

# Set cohort dates and censoring dates
censor_date_end = mdy("12-31-2023")
censor_date_train = mdy("11-15-2022")
filter_date_start = mdy("01-01-2019")
filter_date_end_train = mdy("10-01-2022")

# Retain listing date, transplant date for waiting time calculations
df = df %>% mutate(can_listing_dt = if_else(grepl("CAN_LISTING_DT", unique_event), unique_date, NA),
                   transplant_date = if_else(grepl("REC_TX_DT", unique_event), unique_date, NA)) %>%
  fill(can_listing_dt, transplant_date, .direction = "downup")

# gets earliest date ever at each status
df = df %>%
  group_by(PX_ID, status) %>%  # Assuming PX_ID is the patient identifier
  mutate(
    status_start = as.Date(first(unique_date))
  ) %>% mutate(
    status_start_1 = if_else(status == 1, status_start, NA),
    status_start_2 = if_else(status == 2, status_start, NA),
    status_start_3 = if_else(status == 3, status_start, NA),
    status_start_4 = if_else(status == 4, status_start, NA),
    status_start_5 = if_else(status == 5, status_start, NA),
    status_start_6 = if_else(status == 6, status_start, NA)
  ) %>%
  ungroup()

df = df %>% group_by(PX_ID) %>%
  mutate(
  durable_LVAD_old=durable_LVAD,
  durable_LVAD = case_when(
    durable_LVAD == 1 ~ 1,
    durable_lvad_implant_dt <= unique_date ~ 1,
    is.na(durable_lvad_implant_dt) & any(durable_LVAD == 1) ~ 1,
    TRUE ~ 0
  )) %>%
  ungroup()
# is this a mistake to recode durable_LVAD=1 early, for people like this?
df %>% filter(PX_ID==936345) %>% 
  select(PX_ID,can_listing_dt,unique_event,unique_date,status,durable_lvad_implant_dt,
         durable_LVAD,durable_LVAD_old) %>%
  head(20)
```

## Make composite outcome (death or deterioration) and censor candidates on December 31st, 2023
```{r}
df = df  %>% 
  filter(unique_date <= censor_date_end) %>% 
  group_by(PX_ID) %>%
  mutate(
    outcome = ifelse(!is.na(CAN_REM_CD) & removal_time == t_stop & CAN_REM_CD == 13, 1, outcome),
    any_tx=any(transplant==1),
    outcome = ifelse(any_tx,0,outcome),
    outcome_sum = cumsum(cumsum(outcome))) %>%
  ungroup() %>% filter(outcome_sum<=1) %>% select(-outcome_sum) %>%
  group_by(PX_ID) %>%
  mutate(
    need_to_censor = row_number() == n() & t_stop - t_start >  censor_date_end - as.Date(unique_date),
    t_stop = ifelse(need_to_censor, t_start + censor_date_end - as.Date(unique_date), t_stop),
    outcome = ifelse(need_to_censor, 0, outcome)
  )
table(df$t_stop >= df$t_start)

# some numbers
dim(table(df$PX_ID)) # 24531
table(df$transplant) # 16963  
table(df$outcome) # 2560
df %>% filter(outcome==1,!any_tx) %>% nrow() # 2560 pretx deaths


```

## Identify candidates who were delisted
## Determine time to death / censor for these candidates
```{r}

 # few people with rows/death after removal
df %>% filter(cumsum(cumsum(grepl("CAN_REM_DT", unique_event)))>1) %>% select(PX_ID,unique_event,outcome)

# confirm that all transplant recipients have CAN_REM_DT in unique_event
df %>% filter(transplant==1,!grepl("CAN_REM_DT",unique_event))

# create delisting indicator
df = df %>% mutate(
  delisting = ifelse(grepl("CAN_REM_DT", unique_event) & !grepl("REC_TX_DT", unique_event) &
                       !(CAN_REM_CD %in% 13), 1, 0)
)

# create delisting_dt variable and fill
df = df %>% group_by(PX_ID) %>% mutate(
  delisting_dt = if_else(delisting == 1, unique_date, NA)
) %>% fill(delisting_dt, .direction = "downup") %>% ungroup()

# similar to row from transplant to death/censoring, make one row from delisting to death
df = df %>% group_by(PX_ID) %>%
  mutate(delisting = cumsum(cumsum(delisting)),
         t_stop = ifelse(delisting == 1, last(t_stop), t_stop),
         outcome = ifelse(delisting == 1, last(outcome), outcome)) %>%
  filter(delisting <= 1) %>%
  ungroup

# make time from delisting to death/censor
df = df %>% mutate(
  time_until_death_after_delisting = ifelse(delisting == 1 & outcome == 1, t_stop - t_start, NA),
  time_until_censor_after_delisting = ifelse(delisting == 1 & outcome == 0, censor_date_end - as.Date(unique_date), NA)) %>% group_by(PX_ID) %>% fill(time_until_death_after_delisting, time_until_censor_after_delisting, .direction = "up") %>% ungroup

# including posttx deaths in time_until_death_after_delisting? no
df %>% filter(transplant==1,!is.na(time_until_death_after_delisting)) %>% nrow()

# censoring anyone who dies? no
df %>% filter(outcome==1,!is.na(time_until_censor_after_delisting)) %>% nrow()

# count transplant, death, delisting
df %>% count(transplant,delisting,outcome)
```


## Define short term MCS ever variables
These variables need to be created before we subset to the first row per 1-week period, to ensure we are capturing all VADs.
```{r}

df <- df %>%
  group_by(PX_ID) %>%
  
  mutate(ECMO_ever = case_when(ECMO == '1' ~ '1', TRUE ~ NA_character_),
         temp_surg_ever = case_when(temp_surg_LVAD == '1' ~ '1', TRUE ~ NA_character_),
         BiVAD_no_discharge_ever = case_when(BiVAD_no_discharge == '1' ~ '1', TRUE ~ NA_character_)) %>%
  
  fill(ECMO_ever, .direction = 'down') %>%
  fill(temp_surg_ever, .direction = 'down') %>%
  fill(BiVAD_no_discharge_ever, .direction = 'down') %>%

  mutate(ECMO_ever = case_when(is.na(ECMO_ever) ~ '0', TRUE ~ ECMO_ever),
         temp_surg_ever = case_when(is.na(temp_surg_ever) ~ '0', TRUE ~ temp_surg_ever),
         BiVAD_no_discharge_ever = case_when(is.na(BiVAD_no_discharge_ever) ~ '0', TRUE ~ BiVAD_no_discharge_ever))


```

## Discrete time data construction
Expand data into 1-week discrete-time format, from listing to removal, taking variables at the start of each week. Note that `outcome` (death or deterioration) can be 1 for multiple rows after this step, if a death occurred at the end of the period. Also, `outcome` can include posttransplant deaths (just a few, due to data discrepancies with death before transplant).

The important fields created by this step are `interval_start` and `interval_stop`, which are always multiples of 7. That is, `interval_stop` will be the end of the 1-week interval regardless of when death or transplant occurred. Note that `t_start` and `interval_start` are 0 on the candidate's listing date.
```{r}
# create tx dataset
df_tx = df %>% filter(transplant == 1 | (CAN_REM_CD == 4 & grepl("CAN_REM_DT",unique_event)))

# remove rows after transplant, removal (including removal row)
df0 = df
df = df %>% filter(cumsum(grepl("CAN_REM_DT", unique_event))<1 |
                     n()==1)

# if removal is the only row, removed on listing date
df %>% ungroup %>% filter(grepl("CAN_REM_DT", unique_event),
                          !grepl("LISTING",unique_event))
df %>% ungroup %>% filter(grepl("CAN_REM_DT", unique_event),
                          !t_start==0)

# recode t_stop if removed on listing date
df = df %>%
  mutate(t_stop = ifelse(grepl("CAN_REM_DT",unique_event),1,t_stop))

# count up transplants, deaths
dim(table(df$PX_ID)) # 24531
nrow(df_tx) # 16963
dim(table(df_tx$PX_ID)) # 16963
df %>% filter(!is.na(time_until_death_after_delisting) | outcome==1) %>% distinct(PX_ID) %>% nrow() # 2560

# look at people with t_start = t_stop
df %>% filter(t_start==t_stop) %>%
  select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,transplant)
df %>% filter(t_start==t_stop,n()!=1) %>%
  select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,transplant)
df %>% filter(t_start==t_stop,outcome==1) %>%
  select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,transplant)
df %>% filter(t_start==t_stop,delisting==1) %>%
  select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,delisting)

# recode t_stop as 1 if t_stop=0
## otherwise, remove when t_start=t_stop
df = df %>%
  mutate(t_stop=ifelse(t_stop==0,1,t_stop)) %>%
  filter(t_start!=t_stop)
dim(table(df$PX_ID)) # 24531

# CHECK THIS
# some rows after removal_time?? Be careful with last_active_date / last_inactive_date. The max is not the last day waiting.
df %>% select(PX_ID,unique_event,unique_date,t_start,t_stop,removal_time) %>% filter(t_start > removal_time)
df %>% ungroup %>% filter(t_start > removal_time) %>%
  count(outcome,transplant,delisting)
df %>% filter(any(t_start > removal_time)) %>% ungroup %>%
  count(outcome,transplant,delisting)
df %>% ungroup %>% filter(t_start > removal_time) %>% # count(is.na(CAN_REM_CD)) ## all still waiting, not removed
  select(PX_ID,unique_event,unique_date,t_start,t_stop,CAN_REM_CD,removal_time,can_listing_dt,last_active_date,last_inactive_date) %>% head

# define period to group observations that occur within the same week
# take first observation within each week
# duplicate each observation based on how many days there are until the next period
  ## NOTE: duplication causes some people to have outcome=1 multiple times, if the interval ending in death lasted >1 week
# define interval_start, interval_stop as the 7 day periods

df_final = df %>% mutate(
  period = ceiling(t_stop/7)
) %>% group_by(PX_ID, period) %>% 
  # make outcome=1 for entire period if outcome=1
  mutate(outcome = ifelse(sum(outcome)>0,1,outcome)) %>%
  # taking the first row per period
  filter(row_number() == 1) %>% 
  mutate(
    t_start = ceiling(t_start/7)*7,
    t_stop = ceiling(t_stop/7)*7, 
    duplicates = (t_stop - t_start)/7
  ) %>% ungroup() %>% uncount(duplicates) %>%
  group_by(PX_ID) %>%
  mutate(interval_start = seq(0, max(t_stop) - 7, by = 7),
         interval_stop = interval_start + 7) %>%
  ungroup()

# count deaths
df_final %>% filter(!is.na(time_until_death_after_delisting) | outcome==1) %>% distinct(PX_ID) %>% nrow() # 2560
```

## Add dates back into discrete dataset
Create a `current_date` variable for the date on `interval_start`.
```{r}
# set current_date 
df_final = df_final %>% 
  mutate(
    current_date = as.Date(can_listing_dt) + interval_start
    )
summary(df_final$current_date)

# confirm no rows after delisting
df_final %>% filter(!is.na(delisting_dt)) %>% # all delisted candidates
  filter(removal_time < (current_date-as.Date(can_listing_dt))) %>% # removal time never before current interval
  select(PX_ID,current_date,can_listing_dt,removal_time,interval_start,interval_stop,unique_date,unique_event)
```

## Set transplant = 1 for PX_ID in transplant list, at their last observation on the waitlist
```{r}
tx_patients = unique(df_tx$PX_ID)

# set transplant=1 on last row of df_final
df_final = df_final %>% group_by(PX_ID) %>%
  mutate(
    transplant = ifelse(row_number() == n() & PX_ID %in% tx_patients, 1, 0)
    ) %>% 
  select(-c(t_start, t_stop, period)) %>% # , unique_event, unique_date
  relocate(interval_start, interval_stop, .after=CAN_LISTING_CTR_CD) %>%
  ungroup
table(df_final$transplant) # 16963 
length(tx_patients) # 16963 

# confirm no rows after transplant
df_final %>% filter(transplant==1,is.na(transplant_time)) # always have a transplant_time
df_final %>% # transplant_time never before current interval
  filter(transplant_time < (current_date-as.Date(can_listing_dt))) %>%
  select(PX_ID,current_date,can_listing_dt,removal_time,interval_start,interval_stop,unique_date,unique_event)
```

## Create train/test split on 1/1/2023
For tables on missingness, subset discrete dataset to all weeks that started on or after 1/1/2019. Then separate into training and test sets, with training follow-up through 11/15/2022 for all candidates listed by 10/01/2022.
```{r}
# check number of transplants, px_id
table(df_final$transplant) # 16963 
dim(table(df_final$PX_ID)) # 24531

# reduce data to cohort, create train/testing split
df_final = df_final %>% 
  filter(current_date >= filter_date_start) %>% 
  mutate(
      train_test_split = ifelse(current_date <= censor_date_train & 
                                  can_listing_dt <= filter_date_end_train, "train", "test"),
      train_test_split = as.factor(train_test_split)
  )

# check number of transplants, px_id - lost some
table(df_final$transplant) # 16931  
dim(table(df_final$PX_ID)) # 24487

# check that it worked
with(df_final,tapply(current_date,train_test_split,summary))

# how many rows per patient?
df_final %>% group_by(PX_ID,train_test_split) %>%
  mutate(rown=n()) %>%
  ungroup %>%
  with(.,tapply(rown,train_test_split,summary))
# $test
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    1.00   27.00   50.00   42.08   59.00   65.00 
# 
# $train
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     1.0    48.0    94.0   101.1   153.0   203.0 
```

# Missing data and out-of-range values

## Missingness before out of range fix
```{r}
table1(~  albumin + bilirubin + creatinine + sodium + BNP + cardiac_output + diastolicBP + systolicBP + PCWP + PASP + PADP + central_venous_pressure + factor(durable_LVAD) + factor(ECMO_ever) | train_test_split, data=df_final)

# initial week, so that candidates appear only once
table1(~  albumin + bilirubin + creatinine + sodium + BNP + cardiac_output + diastolicBP + systolicBP + PCWP + PASP + PADP + central_venous_pressure + factor(durable_LVAD) + factor(ECMO_ever) | train_test_split, data=df_final %>% group_by(PX_ID) %>% slice(1) %>% ungroup)
```

## Recode out-of-range values
```{r}
# out of range value flag
df_final = df_final %>% mutate(
  out_of_range_flag = case_when(
    sodium < 100 | sodium > 200 ~ 1,
    BNP_NT_Pro == 0 & BNP > 20000 ~ 1,
    BNP_NT_Pro == 1 & BNP > 40000 ~ 1,
    bilirubin > 50 ~ 1,
    albumin < 0.5 | albumin > 9.9 ~ 1,
    cardiac_output < 0.2 | cardiac_output > 15 ~ 1,
    diastolicBP > 110 ~ 1,
    systolicBP < 10 | systolicBP > 300 ~ 1,
    PCWP > 100 ~ 1,
    PASP < 10 | PASP > 120 ~ 1,
    PADP > 110 ~ 1,
    central_venous_pressure > 110 ~ 1,
    creatinine < 0.1 | creatinine > 40 ~ 1,
    TRUE ~ 0)
)

# set out of range values to missing
df_final = df_final %>% mutate(
  sodium = ifelse(sodium < 100 | sodium > 200, NA, sodium),
  BNP = case_when(
    BNP_NT_Pro == 0 & BNP > 20000 ~ NA,
    BNP_NT_Pro == 1 & BNP > 40000 ~ NA,
    TRUE ~ BNP
  ),
  bilirubin = ifelse(bilirubin > 50, NA, bilirubin),
  albumin = ifelse(albumin < 0.5 | albumin > 9.9, NA, albumin),
  cardiac_output = ifelse(cardiac_output < 0.2 | cardiac_output > 15, NA, cardiac_output),
  diastolicBP = ifelse(diastolicBP > 110, NA, diastolicBP),
  systolicBP = ifelse(systolicBP < 10 | systolicBP > 300, NA, systolicBP),
  PCWP = ifelse(PCWP > 100, NA, PCWP),
  PASP = ifelse(PASP < 10 | PASP > 120, NA, PASP),
  PADP = ifelse(PADP > 110, NA, PADP),
  central_venous_pressure = ifelse(central_venous_pressure > 110, NA, central_venous_pressure),
  creatinine = ifelse(creatinine < 0.1 | creatinine > 40, NA, creatinine)
)

```

## Data summary before imputation
```{r}
df_final_preImpute = df_final

table1(~  albumin + bilirubin + creatinine + sodium + BNP + cardiac_output + diastolicBP + systolicBP + PCWP + PASP + PADP + central_venous_pressure  + factor(out_of_range_flag) + factor(durable_LVAD) + factor(ECMO_ever) | train_test_split, data=df_final) # very low percent missing except w/ BNP
summary(df_final$BNP[df_final$BNP_NT_Pro %in% 0]) # should be between 0 and 20000
summary(df_final$BNP[df_final$BNP_NT_Pro %in% 1]) # should be between 0 and 40000
```

## Body surface area calculation
```{r}
# impute missing candidate values based on median in our cohort
cand_thor_impute = df_cand_thor %>%
  filter(PX_ID %in% df_final$PX_ID) %>%
  # summarise(med_hgt=median(CAN_HGT_CM,na.rm=T),
  #           med_wgt=median(CAN_WGT_KG,na.rm=T))
  mutate(CAN_HGT_CM = ifelse(is.na(CAN_HGT_CM),median(CAN_HGT_CM,na.rm=T),CAN_HGT_CM),
         CAN_WGT_KG = ifelse(is.na(CAN_WGT_KG),median(CAN_WGT_KG,na.rm=T),CAN_WGT_KG),
         bsa = 0.007184 * (CAN_HGT_CM ^ 0.725) * (CAN_WGT_KG ^ 0.425))

# merge df_final w/ cand_thor
df_final <- left_join(df_final, cand_thor_impute, by='PX_ID')
```

### Fill missing arterial lactate, LDH, Hemoglobin, BNP, sodium, albumin, bilirubin, creatinine
```{r}
# fill missing arterial lactate, LDH, Hemoglobin, BNP, sodium, albumin, bilirubin, creatinine
with(df_final, median(BNP[BNP_NT_Pro == '0'],na.rm=T))
with(df_final, median(BNP[BNP_NT_Pro == '1'],na.rm=T))
df_final %>% select(sodium,albumin,bilirubin,creatinine) %>% apply(.,2,median,na.rm=T)
df_final <- df_final %>%
  mutate(
    arterial_lactate = ifelse(is.na(arterial_lactate), 1, arterial_lactate),
    LDH = ifelse(is.na(LDH), 220, LDH),
    HemoHemoglobin = case_when(
      is.na(HemoHemoglobin) & CAN_GENDER == 'M' ~ 14.9,
      is.na(HemoHemoglobin) & CAN_GENDER == 'F' ~ 13.3,
      TRUE ~ HemoHemoglobin
    ),
    BNP_NT_Pro = case_when(
      is.na(BNP_NT_Pro) ~ 0,
      TRUE ~ BNP_NT_Pro
    ),
    BNP = case_when(
      is.na(BNP) & BNP_NT_Pro == '0' ~ median(BNP[BNP_NT_Pro == '0'], na.rm = TRUE),
      is.na(BNP) & BNP_NT_Pro == '1' ~ median(BNP[BNP_NT_Pro == '1'], na.rm = TRUE),
      TRUE ~ BNP
    ),
    sodium = ifelse(is.na(sodium), median(sodium, na.rm = TRUE), sodium),
    albumin = ifelse(is.na(albumin), median(albumin, na.rm = TRUE), albumin),
    bilirubin = ifelse(is.na(bilirubin), median(bilirubin, na.rm = TRUE), bilirubin),
    creatinine = ifelse(is.na(creatinine), median(creatinine, na.rm = TRUE), creatinine)
  )
```

## Fill missing cardiac_output, central_venous_pressure, PASP, PADP, PCWP, systolicBP, diastolicBP, resting_HR
```{r}
df_final <- df_final %>%
  ungroup() %>%
  mutate(
    cardiac_output = case_when(
      is.na(cardiac_output) & durable_LVAD == '1' ~ 2.5 * bsa,
      is.na(cardiac_output) & durable_LVAD != '1' ~ 2.2 * bsa,
      TRUE ~ cardiac_output
    ), 
    central_venous_pressure = case_when(
      is.na(central_venous_pressure) & durable_LVAD == '1' ~ 10,
      is.na(central_venous_pressure) & durable_LVAD != '1' ~ 12,
      TRUE ~ central_venous_pressure
    ),
    PASP = case_when(
      is.na(PASP) & durable_LVAD == '1' ~ 35,
      is.na(PASP) & durable_LVAD != '1' ~ 38,
      TRUE ~ PASP
    ),
    PADP = case_when(
      is.na(PADP) & durable_LVAD == '1' ~ 15,
      is.na(PADP) & durable_LVAD != '1' ~ 18,
      TRUE ~ PADP
    ),
    PCWP = case_when(
      is.na(PCWP) & durable_LVAD == '1' ~ 15,
      is.na(PCWP) & durable_LVAD != '1' ~ 18,
      TRUE ~ PCWP
    ),
    systolicBP = case_when(
      is.na(systolicBP) & durable_LVAD == '1' ~ 100,
      is.na(systolicBP) & durable_LVAD != '1' ~ 110,
      TRUE ~ systolicBP
    ),
    diastolicBP = ifelse(is.na(diastolicBP), 80, diastolicBP),
    resting_HR = ifelse(is.na(resting_HR), 8.5, resting_HR)
  )
```

## Calculate eGFR, define short mcs ever
```{r}
df_final <- df_final %>%
  mutate(eGFR = case_when(
    CAN_GENDER == 'F' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.7), 1)^(-0.241)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^(age_at_listing) * 1.012,
    
    CAN_GENDER == 'M' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.9), 1)^(-0.302)) *
      (pmax((creatinine / 0.9), 1)^(-1.2)) * 0.9938^(age_at_listing)),
    short_MCS_ever_old = 
         case_when(ECMO_ever == '1' | temp_surg_ever == '1' | BiVAD_no_discharge_ever == '1' ~ 1,
           TRUE ~ 0),
    short_MCS_ever = 
         case_when(ECMO_ever == '1' | temp_surg_ever == '1' | BiVAD_no_discharge_ever == '1' | CAN_ECMO %in% 1 ~ 1,
           TRUE ~ 0))
with(df_final,table(short_MCS_ever_old,short_MCS_ever,useNA="ifany"))


```

## Define hemodynamics, impute missing values
```{r}
# define hemodynamics, impute missing values
df_final <- df_final %>%
  mutate(
    cpo = (1/451) * cardiac_output * (((1/3)*systolicBP) + ((2/3)*diastolicBP)),
    api = (systolicBP - diastolicBP) / PCWP,
    papi = (PASP - PADP) / central_venous_pressure) %>%
  
  mutate(cpo = as.numeric(cpo),
         api = as.numeric(api),
         papi = as.numeric(papi)) %>%
  mutate(api = ifelse(PCWP == 0, 1, api),
         papi = case_when(
            central_venous_pressure == 0 ~ 1,
            papi < 1 & !is.na(papi) ~ 1,
            TRUE ~ papi
    ))
```

## Table 1, with missing data
```{r}
## before imputation
table_data = df_final_preImpute
label(table_data$albumin) = "Albumin"
label(table_data$bilirubin) = "Bilirubin"
label(table_data$creatinine) = "Creatinine"
label(table_data$sodium) = "Sodium"
label(table_data$BNP) = "BNP"
label(table_data$cardiac_output) = "Cardiac output"
label(table_data$diastolicBP) = "Diastolic BP"
label(table_data$systolicBP) = "Systolic BP"
label(table_data$PCWP) = "PCWP"
label(table_data$PASP) = "PASP"
label(table_data$PADP) = "PADP"
label(table_data$central_venous_pressure) = "Central venous pressure"
table_data$out_of_range_flag = factor(table_data$out_of_range_flag)
label(table_data$out_of_range_flag) = "Any data out of range"
table_data$durable_LVAD = factor(table_data$durable_LVAD)
label(table_data$durable_LVAD) = "Durable LVAD"
table_data$ECMO_ever = factor(table_data$ECMO_ever)
label(table_data$ECMO_ever) = "Ever on ECMO"
table1(~  albumin + bilirubin + creatinine + sodium + BNP + cardiac_output + diastolicBP + systolicBP + PCWP + PASP + PADP + central_venous_pressure  + durable_LVAD + ECMO_ever + out_of_range_flag | train_test_split, data=table_data) # very low percent missing except w/ BNP

# summary of number of candidates
dim(table(df_final_preImpute$PX_ID)) # 24852
df_final_preImpute %>% distinct(PX_ID,train_test_split) %>%
  count(train_test_split) %>%
  mutate(prop=n/dim(table(df_final_preImpute$PX_ID)))
#   train_test_split     n  prop
#   <fct>            <int> <dbl>
# 1 test              8052 0.329
# 2 train            18877 0.771

# summary of number of rows per candidate
df_final_preImpute %>% group_by(PX_ID,train_test_split) %>%
    mutate(rown=n()) %>%
    ungroup %>%
    # with(.,tapply(rown,train_test_split,summary))
    with(.,tapply(rown,train_test_split,sd))
df_final_preImpute %>% group_by(PX_ID) %>%
    mutate(rown=n()) %>%
    ungroup %>%
    pull(rown) %>% summary # sd

## after imputation - supplement
table_data = df_final
label(table_data$albumin) = "Albumin"
label(table_data$bilirubin) = "Bilirubin"
label(table_data$creatinine) = "Creatinine"
label(table_data$sodium) = "Sodium"
label(table_data$BNP) = "BNP"
label(table_data$cardiac_output) = "Cardiac output"
label(table_data$diastolicBP) = "Diastolic BP"
label(table_data$systolicBP) = "Systolic BP"
label(table_data$PCWP) = "PCWP"
label(table_data$PASP) = "PASP"
label(table_data$PADP) = "PADP"
label(table_data$central_venous_pressure) = "Central venous pressure"
table_data$out_of_range_flag = factor(table_data$out_of_range_flag)
label(table_data$out_of_range_flag) = "Any data out of range"
table_data$durable_LVAD = factor(table_data$durable_LVAD)
label(table_data$durable_LVAD) = "Durable LVAD"
table_data$ECMO_ever = factor(table_data$ECMO_ever)
label(table_data$ECMO_ever) = "Ever on ECMO"
table1(~  albumin + bilirubin + creatinine + sodium + BNP + cardiac_output + diastolicBP + systolicBP + PCWP + PASP + PADP + central_venous_pressure  + durable_LVAD + ECMO_ever + out_of_range_flag | train_test_split, data=table_data) # very low percent missing except w/ BNP
```

## Calculate wait times
```{r}
calculate_wait_times <- function(status_vector, current_date, status_start_1, status_start_2, status_start_3, status_start_4, status_start_5, status_start_6) {
  n <- length(status_vector)
  
  # Convert to character if not already
  status_vector <- as.character(status_vector)
  
  # Initialize wait time vectors
  wait_1 <- numeric(n)
  wait_2 <- numeric(n)
  wait_3 <- numeric(n)
  wait_4 <- numeric(n)
  wait_5 <- numeric(n)
  wait_6 <- numeric(n)
  
  # Flags to check if first time the status is encountered
  first_time_1 <- TRUE
  first_time_2 <- TRUE
  first_time_3 <- TRUE
  first_time_4 <- TRUE
  first_time_5 <- TRUE
  first_time_6 <- TRUE
  
  # Initialize cumulative wait times
  cum_wait_1 <- 0
  cum_wait_2 <- 0
  cum_wait_3 <- 0
  cum_wait_4 <- 0
  cum_wait_5 <- 0
  cum_wait_6 <- 0
  
  # Loop through each time point
  for (i in 1:n) {
    current_status <- status_vector[i]
    
    if (is.na(current_status) || current_status == "inactive") {
      # Freeze wait times during inactive period, no accrual but keep the last values
      wait_1[i] <- cum_wait_1
      wait_2[i] <- cum_wait_2
      wait_3[i] <- cum_wait_3
      wait_4[i] <- cum_wait_4
      wait_5[i] <- cum_wait_5
      wait_6[i] <- cum_wait_6  # Ensure that the last value of wait_6 is retained
    } else {
      # Increment wait times only when the current status allows
      
      # Status 1 logic
      if (current_status == "1") {
        if (first_time_1) {
          cum_wait_1 <- as.numeric(current_date[i] - status_start_1[i])
          first_time_1 <- FALSE
        } else {
          cum_wait_1 <- cum_wait_1 + 7
        }
      }
      
      # Status 2 logic
      if (current_status <= "2") {
        if (first_time_2) {
          cum_wait_2 <- as.numeric(current_date[i] - status_start_2[i])
          first_time_2 <- FALSE
        } else {
          cum_wait_2 <- cum_wait_2 + 7
        }
      }
      
      # Status 3 logic
      if (current_status <= "3") {
        if (first_time_3) {
          cum_wait_3 <- as.numeric(current_date[i] - status_start_3[i])
          first_time_3 <- FALSE
        } else {
          cum_wait_3 <- cum_wait_3 + 7
        }
      }
      
      # Status 4 logic
      if (current_status <= "4") {
        if (first_time_4) {
          cum_wait_4 <- as.numeric(current_date[i] - status_start_4[i])
          first_time_4 <- FALSE
        } else {
          cum_wait_4 <- cum_wait_4 + 7
        }
      }
      
      # Status 5 logic
      if (current_status <= "5") {
        if (first_time_5) {
          cum_wait_5 <- as.numeric(current_date[i] - status_start_5[i])
          first_time_5 <- FALSE
        } else {
          cum_wait_5 <- cum_wait_5 + 7
        }
      }
      
      # Status 6 logic
      if (current_status == "6") {  # Only accrue when status is exactly 6
        if (first_time_6) {
          cum_wait_6 <- as.numeric(current_date[i] - status_start_6[i])
          first_time_6 <- FALSE
        } else {
          cum_wait_6 <- cum_wait_6 + 7
        }
      }
      
      # Assign cumulative wait times to the wait vectors
      wait_1[i] <- cum_wait_1
      wait_2[i] <- cum_wait_2
      wait_3[i] <- cum_wait_3
      wait_4[i] <- cum_wait_4
      wait_5[i] <- cum_wait_5
      wait_6[i] <- cum_wait_6
    }
  }
  
  # Return the wait time vectors as a data frame
  data.frame(wait_1, wait_2, wait_3, wait_4, wait_5, wait_6)
}



df_final <- df_final %>%
  group_by(PX_ID) %>%
  do({
    wait_times <- calculate_wait_times(
      status_vector = as.character(.$status),
      current_date = .$current_date,
      status_start_1 = .$status_start_1,
      status_start_2 = .$status_start_2,
      status_start_3 = .$status_start_3,
      status_start_4 = .$status_start_4,
      status_start_5 = .$status_start_5,
      status_start_6 = .$status_start_6
    )
    cbind(., wait_times)  # Combine the original dataframe with the new wait times
  }) %>%
  ungroup()  # Ungroup after processing

# check that wait time freezes when inactive
df_final  %>% select(PX_ID, status, interval_start, interval_stop, wait_1, wait_2, wait_3, wait_4, wait_5, wait_6)
```

## Durable LVAD checks
```{r}
df_final %>% select(PX_ID, current_date, durable_LVAD, durable_lvad_implant_dt, status, status_criteria) %>% filter(!is.na(durable_lvad_implant_dt) & durable_LVAD == 1 & current_date > "2018-10-18" & current_date < durable_lvad_implant_dt) %>% arrange(PX_ID)

df_final %>% filter(PX_ID == 654863 & durable_LVAD == 1)

table(df_final$status_criteria)
df_final %>% filter(status_criteria == "CriteriaLvadSupport")
df_final %>% mutate(durable_lvad_implant_dt = as.Date(durable_lvad_implant_dt)) %>% filter(status_criteria == "CriteriaLvadSupport" & !(status %in% c(4,6, "inactive"))) %>% select(PX_ID, status, durable_LVAD, current_date,  durable_lvad_implant_dt)

df_final %>% filter(durable_LVAD == 1 & is.na(durable_lvad_implant_dt)) %>% select(PX_ID, status, status_criteria, current_date, durable_lvad_implant_dt, durable_LVAD)

```

## Calculate durable LVAD time
```{r}
df_final = df_final %>% mutate(
  durable_lvad_implant_dt = as.Date(durable_lvad_implant_dt))

df_final = df_final %>% mutate(
  durable_lvad_time = case_when(
    is.na(durable_lvad_implant_dt) & durable_LVAD == 1 ~ as.numeric(current_date - as.Date(can_listing_dt)),
    durable_lvad_implant_dt <= current_date ~ as.numeric(current_date - durable_lvad_implant_dt),
    TRUE ~ 0
  )
)
df_final  %>% filter(durable_LVAD == 1) %>% select(durable_lvad_implant_dt, current_date, can_listing_dt, durable_lvad_time)

```

## Save discrete time data
```{r}
# check numbers
dim(table(df_final$PX_ID)) # 24487
table(df_final$transplant) # 16931

# save dataset
save("df_final", file =  "discrete_final_data.RData")
```

## Durable LVAD summary
```{r}

df_final = df_final %>%
  group_by(PX_ID) %>% # Group by PX_ID
  mutate(
    first_lvad_date = min(current_date[status_criteria == "CriteriaLvadSupport"], na.rm = TRUE) # Find the first date
  ) %>%
  ungroup() 

filtered = df_final %>%
  group_by(PX_ID) %>%
  filter(any(status_criteria == "CriteriaLvadSupport")) %>% mutate(
  lvad_time_test = first_lvad_date - max(as.Date(can_listing_dt), durable_lvad_implant_dt, na.rm = T) 
)

ggplot(filtered, aes(x = lvad_time_test)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of LVAD Time Test",
       x = "LVAD Time Test",
       y = "Frequency") +
  theme_minimal()

filtered = as.data.frame(filtered)
base::summary(filtered$lvad_time_test)
?summary
variable = filtered$lvad_time_test
filtered = filtered %>% select(PX_ID, lvad_time_test) %>%
  distinct()
summary(as.numeric(variable))
as.numericvariable
max(df_final$lvad_time_test, na.rm = T)

filtered %>% filter(!is.na(lvad_time_test)) %>% select(PX_ID, current_date, can_listing_dt ,durable_lvad_implant_dt, lvad_time_test)

df_final %>%
  group_by(PX_ID) %>%
  filter(any(status_criteria) == "CriteriaLvadSupport")

df_final %>% filter(status_criteria == "CriteriaLvadSupport")

df_final %>% filter(PX_ID == 542942 & status_criteria == "CriteriaLvadSupport")

filtered %>% select(lvad_time_test)
filtered %>% filter(as.numeric(lvad_time_test) == 77)

df_final %>% filter(PX_ID == 654863) %>% select(PX_ID, status, current_date, durable_LVAD, can_listing_dt, durable_lvad_implant_dt, first_lvad_date, status_criteria)

df_final %>% filter(as.numeric(lvad_time_test) > 0 & current_date > "2019-01-01")
```

```{r}
df %>% filter(is.na(durable_lvad_implant_dt) & durable_LVAD == 1) %>% count(PX_ID)
df %>% filter(PX_ID == 1107906) %>% select(durable_LVAD, durable_lvad_implant_dt, status)
```









