---
title: "Create 1-week discrete-time dataset"
output: html_document
date: "2026-01-07"
editor_options: 
  chunk_output_type: console
---

## Load data and necessary libraries
## Dataset can be constructed from the heart_pipeline_final.Rmd file
```{r setup, include=FALSE}
rm(list = ls())
gc()
knitr::opts_chunk$set(echo = TRUE)

grace=T
if (!grace) {
  # Molly:
  load("C:/Users/mhwhi/Downloads/heart_post_policy.RData")
  df_cand_thor <- haven::read_sas('C:/Users/mhwhi/Downloads/pubsaf2403/cand_thor.sas7bdat', col_select=c(PX_ID, CAN_GENDER, CAN_RACE, CAN_HGT_CM, CAN_WGT_KG, CAN_ECMO))
} else {
  load("heart_post_policy.RData")
  groups(df)
  curr_data_release = "//srtr-fs-p01/SAF/Internal/SAF2512/thor"
  df_cand_thor <- haven::read_sas(file.path(curr_data_release,'cand.sas7bdat'),
                                  col_select=c(PX_ID, CAN_GENDER, CAN_RACE, CAN_HGT_CM, CAN_WGT_KG, CAN_ECMO))
}

library(renv)
library(dplyr)
library(tidyr)
library(lubridate)
library(table1)
library(purrr)
library(ggplot2)

### Cohort dates ###
# Train: 1/1/19-6/30/23 (censor on 8/15/23)
# Validation: 9/1/23-8/31/24
# Test: 9/1/24-8/31/25

# Set cohort dates and censoring dates
censor_date_end = mdy("8-31-2025")
censor_date_train = mdy("6-30-2023") # max current_date in training
filter_date_start = mdy("01-01-2019")
filter_date_end_train = mdy("5-15-2023") # max listing_date in training
test_date_start = mdy("9-1-2023")
test2_date_start = mdy("9-1-2024")

# All 1-week intervals that began between 1/1/2019 and 6/30/2023, for candidates listed by 5/15/2023, were retained in the model training data, while all other 1-week intervals were held out for model assessment (validation and test).

# Retain listing date, transplant date for waiting time calculations
df = df %>% mutate(can_listing_dt = if_else(grepl("CAN_LISTING_DT", unique_event), unique_date, NA),
                   transplant_date = if_else(grepl("REC_TX_DT", unique_event), unique_date, NA)) %>%
  group_by(PX_ID) %>%
  fill(can_listing_dt, transplant_date, .direction = "downup")

# gets earliest date ever at each status
df = df %>%
  group_by(PX_ID, status) %>%  # Assuming PX_ID is the patient identifier
  mutate(
    status_start = as.Date(first(unique_date))
  ) %>% mutate(
    status_start_1 = if_else(status == 1, status_start, NA),
    status_start_2 = if_else(status == 2, status_start, NA),
    status_start_3 = if_else(status == 3, status_start, NA),
    status_start_4 = if_else(status == 4, status_start, NA),
    status_start_5 = if_else(status == 5, status_start, NA),
    status_start_6 = if_else(status == 6, status_start, NA)
  ) %>%
  ungroup()

df = df %>% group_by(PX_ID) %>%
  mutate(
  durable_LVAD_old=durable_LVAD,
  durable_LVAD = case_when(
    durable_LVAD == 1 ~ 1,
    durable_lvad_implant_dt <= unique_date ~ 1,
    is.na(durable_lvad_implant_dt) & any(durable_LVAD == 1) ~ 1,
    TRUE ~ 0
  )) %>%
  ungroup()
# # is this a mistake to recode durable_LVAD=1 early, for people like this? no, it's LVAD at listing.
# df %>% filter(PX_ID==936345) %>% 
#   select(PX_ID,can_listing_dt,unique_event,unique_date,status,durable_lvad_implant_dt,
#          durable_LVAD,durable_LVAD_old) %>%
#   head(20)
```

## Make composite outcome (death or deterioration) and censor candidates on August 31, 2025 (Dec 2025 SAF)
```{r}
df = df  %>% 
  filter(unique_date <= censor_date_end) %>% 
  group_by(PX_ID) %>%
  mutate(
    outcome = ifelse(!is.na(CAN_REM_CD) & removal_time == t_stop & CAN_REM_CD == 13, 1, outcome),
    any_tx=any(transplant==1),
    outcome = ifelse(any_tx,0,outcome),
    outcome_sum = cumsum(cumsum(outcome))) %>%
  ungroup() %>% filter(outcome_sum<=1) %>% select(-outcome_sum) %>%
  group_by(PX_ID) %>%
  mutate(
    need_to_censor = row_number() == n() & t_stop - t_start >  censor_date_end - as.Date(unique_date),
    t_stop = ifelse(need_to_censor, t_start + censor_date_end - as.Date(unique_date), t_stop),
    outcome = ifelse(need_to_censor, 0, outcome)
  )
table(df$t_stop >= df$t_start)

# some numbers
dim(table(df$PX_ID)) # 34001
table(df$transplant) # 23960  
table(df$outcome) # 3460
df %>% filter(outcome==1,!any_tx) %>% nrow() # 3460 pretx deaths


```

## Identify candidates who were delisted
## Determine time to death / censor for these candidates
```{r}

 # few people with rows/death after removal
df %>% filter(cumsum(cumsum(grepl("CAN_REM_DT", unique_event)))>1) %>% select(PX_ID,unique_event,outcome)

# confirm that all transplant recipients have CAN_REM_DT in unique_event
df %>% filter(transplant==1,!grepl("CAN_REM_DT",unique_event))

# create delisting indicator
df = df %>% mutate(
  delisting = ifelse(grepl("CAN_REM_DT", unique_event) & !grepl("REC_TX_DT", unique_event) &
                       !(CAN_REM_CD %in% 13), 1, 0)
)

# create delisting_dt variable and fill
df = df %>% group_by(PX_ID) %>% mutate(
  delisting_dt = if_else(delisting == 1, unique_date, NA)
) %>% fill(delisting_dt, .direction = "downup") %>% ungroup()

# similar to row from transplant to death/censoring, make one row from delisting to death
df = df %>% group_by(PX_ID) %>%
  mutate(delisting = cumsum(cumsum(delisting)),
         t_stop = ifelse(delisting == 1, last(t_stop), t_stop),
         outcome = ifelse(delisting == 1, last(outcome), outcome)) %>%
  filter(delisting <= 1) %>%
  ungroup

# make time from delisting to death/censor
df = df %>% mutate(
  time_until_death_after_delisting = ifelse(delisting == 1 & outcome == 1, t_stop - t_start, NA),
  time_until_censor_after_delisting = ifelse(delisting == 1 & outcome == 0, censor_date_end - as.Date(unique_date), NA)) %>% group_by(PX_ID) %>% fill(time_until_death_after_delisting, time_until_censor_after_delisting, .direction = "up") %>% ungroup

# including posttx deaths in time_until_death_after_delisting? no
df %>% filter(transplant==1,!is.na(time_until_death_after_delisting)) %>% nrow()

# censoring anyone who dies? no
df %>% filter(outcome==1,!is.na(time_until_censor_after_delisting)) %>% nrow()

# count transplant, death, delisting
df %>% count(transplant,delisting,outcome)
```


## Define short term MCS ever variables
These variables need to be created before we subset to the first row per 1-week period, to ensure we are capturing all VADs.
```{r}

df <- df %>%
  group_by(PX_ID) %>%
  
  mutate(ECMO_ever = case_when(ECMO == '1' ~ '1', TRUE ~ NA_character_),
         temp_surg_ever = case_when(temp_surg_LVAD == '1' ~ '1', TRUE ~ NA_character_),
         BiVAD_no_discharge_ever = case_when(BiVAD_no_discharge == '1' ~ '1', TRUE ~ NA_character_)) %>%
  
  fill(ECMO_ever, .direction = 'down') %>%
  fill(temp_surg_ever, .direction = 'down') %>%
  fill(BiVAD_no_discharge_ever, .direction = 'down') %>%

  mutate(ECMO_ever = case_when(is.na(ECMO_ever) ~ '0', TRUE ~ ECMO_ever),
         temp_surg_ever = case_when(is.na(temp_surg_ever) ~ '0', TRUE ~ temp_surg_ever),
         BiVAD_no_discharge_ever = case_when(is.na(BiVAD_no_discharge_ever) ~ '0', TRUE ~ BiVAD_no_discharge_ever))


```

## Discrete time data construction
Expand data into 1-week discrete-time format, from listing to removal, taking variables at the start of each week. Note that `outcome` (death or deterioration) can be 1 for multiple rows after this step, if a death occurred at the end of the period.

The important fields created by this step are `interval_start` and `interval_stop`, which are always multiples of 7. That is, `interval_stop` will be the end of the 1-week interval regardless of when death or transplant occurred. Note that `t_start` and `interval_start` are 0 on the candidate's listing date.
```{r}
# create tx dataset
df_tx = df %>% filter(transplant == 1 | (CAN_REM_CD == 4 & grepl("CAN_REM_DT",unique_event)))

# remove rows after transplant, removal (including removal row)
df0 = df
df = df %>% filter(cumsum(grepl("CAN_REM_DT", unique_event))<1 |
                     n()==1)

# if removal is the only row, removed on listing date
df %>% ungroup %>% filter(grepl("CAN_REM_DT", unique_event),
                          !grepl("LISTING",unique_event))
df %>% ungroup %>% filter(grepl("CAN_REM_DT", unique_event),
                          !t_start==0)

# recode t_stop if removed on listing date
df = df %>%
  mutate(t_stop = ifelse(grepl("CAN_REM_DT",unique_event),1,t_stop))

# count up transplants, deaths
dim(table(df$PX_ID)) # 34001
nrow(df_tx) # 23960
dim(table(df_tx$PX_ID)) # 23960
df %>% filter(!is.na(time_until_death_after_delisting) | outcome==1) %>% distinct(PX_ID) %>% nrow() # 3460

# # look at people with t_start = t_stop
# df %>% filter(t_start==t_stop) %>%
#   select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,transplant)
# df %>% filter(t_start==t_stop,n()!=1) %>%
#   select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,transplant)
# df %>% filter(t_start==t_stop,outcome==1) %>%
#   select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,transplant)
# df %>% filter(t_start==t_stop,delisting==1) %>%
#   select(PX_ID,unique_event,unique_date,t_start,t_stop,death_time,outcome,delisting)

# recode t_stop as 1 if t_stop=0
## otherwise, remove when t_start=t_stop
df = df %>%
  mutate(t_stop=ifelse(t_stop==0,1,t_stop)) %>%
  filter(t_start!=t_stop)
dim(table(df$PX_ID)) # 34001

# # IMPORTANT
# # some rows after removal_time?? Be careful with last_active_date / last_inactive_date. The max is not the last day waiting.
# df %>% select(PX_ID,unique_event,unique_date,t_start,t_stop,removal_time) %>% filter(t_start > removal_time)
# df %>% ungroup %>% filter(t_start > removal_time) %>%
#   count(outcome,transplant,delisting)
# df %>% filter(any(t_start > removal_time)) %>% ungroup %>%
#   count(outcome,transplant,delisting)
# df %>% ungroup %>% filter(t_start > removal_time) %>% # count(is.na(CAN_REM_CD)) ## all still waiting, not removed
#   select(PX_ID,unique_event,unique_date,t_start,t_stop,CAN_REM_CD,removal_time,can_listing_dt,last_active_date,last_inactive_date) %>% head
# ## conclusion: not a problem for what we're doing w/ the removal_time variable

# define period to group observations that occur within the same week
# take first observation within each week
# duplicate each observation based on how many days there are until the next period
  ## NOTE: duplication causes some people to have outcome=1 multiple times, if the interval ending in death lasted >1 week
# define interval_start, interval_stop as the 7 day periods

df_final = df %>% mutate(
  period = ceiling(t_stop/7)
) %>% group_by(PX_ID, period) %>% 
  # make outcome=1 for entire period if outcome=1
  mutate(outcome = ifelse(sum(outcome)>0,1,outcome)) %>%
  # taking the first row per period
  filter(row_number() == 1) %>% 
  mutate(
    t_start = ceiling(t_start/7)*7,
    t_stop = ceiling(t_stop/7)*7, 
    duplicates = (t_stop - t_start)/7
  ) %>% ungroup() %>% uncount(duplicates) %>%
  group_by(PX_ID) %>%
  mutate(interval_start = seq(0, max(t_stop) - 7, by = 7),
         interval_stop = interval_start + 7) %>%
  ungroup()

# count deaths
df_final %>% filter(!is.na(time_until_death_after_delisting) | outcome==1) %>% distinct(PX_ID) %>% nrow() # 3460
```

## Add dates back into discrete dataset
Create a `current_date` variable for the date on `interval_start`.
```{r}
# set current_date 
df_final = df_final %>% 
  mutate(
    current_date = as.Date(can_listing_dt) + interval_start
    )
summary(df_final$current_date)

# confirm no rows after delisting
df_final %>% filter(!is.na(delisting_dt)) %>% # all delisted candidates
  filter(removal_time < (current_date-as.Date(can_listing_dt))) %>% # removal time never before current interval
  select(PX_ID,current_date,can_listing_dt,removal_time,interval_start,interval_stop,unique_date,unique_event)
```

## Set transplant = 1 for PX_ID in transplant list, at their last observation on the waitlist
```{r}
tx_patients = unique(df_tx$PX_ID)

# set transplant=1 on last row of df_final
df_final = df_final %>% group_by(PX_ID) %>%
  mutate(
    transplant = ifelse(row_number() == n() & PX_ID %in% tx_patients, 1, 0)
    ) %>% 
  select(-c(t_start, t_stop, period)) %>% # , unique_event, unique_date
  relocate(interval_start, interval_stop, .after=CAN_LISTING_CTR_CD) %>%
  ungroup
table(df_final$transplant) # 23960 
length(tx_patients) # 23960 

# confirm no rows after transplant
df_final %>% filter(transplant==1,is.na(transplant_time)) # always have a transplant_time
df_final %>% # transplant_time never before current interval
  filter(transplant_time < (current_date-as.Date(can_listing_dt))) %>%
  select(PX_ID,current_date,can_listing_dt,removal_time,interval_start,interval_stop,unique_date,unique_event)
```

## Create train/test split on 6/30/2023
For tables on missingness, subset discrete dataset to all weeks that started on or after 1/1/2019. Then separate into training and test sets, with training follow-up through 6/30/2023 for all candidates listed by 5/15/2023.
```{r}
# check number of transplants, px_id
table(df_final$transplant) # 23960  
dim(table(df_final$PX_ID)) # 34001

# reduce data to cohort, create train/testing split
df_final = df_final %>% 
  filter(current_date >= filter_date_start) %>% 
  mutate(
      train_test_split = ifelse(current_date <= censor_date_train & 
                                  can_listing_dt <= filter_date_end_train, "train", "test"),
      train_test_split = as.factor(train_test_split)
  )

# check number of transplants, px_id - lost some
table(df_final$transplant) # 23928   
dim(table(df_final$PX_ID)) # 33961

# check that it worked
with(df_final,tapply(current_date,train_test_split,summary))

# how many rows per patient?
df_final %>% group_by(PX_ID,train_test_split) %>%
  mutate(rown=n()) %>%
  ungroup %>%
  with(.,tapply(rown,train_test_split,summary))
# $test
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#    1.00   33.00   68.00   67.13  110.00  120.00 
# 
# $train
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     1.0    49.0    99.0   108.2   160.0   235.0 
```

# Missing data and out-of-range values

## Recode out-of-range values
```{r}
# set out of range values to missing
df_final = df_final %>% mutate(
  sodium = ifelse(sodium < 100 | sodium > 200, NA, sodium),
  BNP = case_when(
    BNP_NT_Pro == 0 & BNP > 20000 ~ NA,
    BNP_NT_Pro == 1 & BNP > 40000 ~ NA,
    TRUE ~ BNP
  ),
  bilirubin = ifelse(bilirubin > 50, NA, bilirubin),
  albumin = ifelse(albumin < 0.5 | albumin > 9.9, NA, albumin),
  cardiac_output = ifelse(cardiac_output < 0.2 | cardiac_output > 15, NA, cardiac_output),
  diastolicBP = ifelse(diastolicBP > 110, NA, diastolicBP),
  systolicBP = ifelse(systolicBP < 10 | systolicBP > 300, NA, systolicBP),
  PCWP = ifelse(PCWP > 100, NA, PCWP),
  PASP = ifelse(PASP < 10 | PASP > 120, NA, PASP),
  PADP = ifelse(PADP > 110, NA, PADP),
  central_venous_pressure = ifelse(central_venous_pressure > 110, NA, central_venous_pressure),
  creatinine = ifelse(creatinine < 0.1 | creatinine > 40, NA, creatinine)
)

```

## Body surface area calculation
```{r}
# impute missing candidate values based on median in our cohort
cand_thor_impute = df_cand_thor %>%
  filter(PX_ID %in% df_final$PX_ID) %>%
  # summarise(med_hgt=median(CAN_HGT_CM,na.rm=T),
  #           med_wgt=median(CAN_WGT_KG,na.rm=T))
  mutate(#CAN_HGT_CM = ifelse(is.na(CAN_HGT_CM),median(CAN_HGT_CM,na.rm=T),CAN_HGT_CM),
         #CAN_WGT_KG = ifelse(is.na(CAN_WGT_KG),median(CAN_WGT_KG,na.rm=T),CAN_WGT_KG),
         bsa = 0.007184 * (CAN_HGT_CM ^ 0.725) * (CAN_WGT_KG ^ 0.425))

# merge df_final w/ cand_thor
df_final <- left_join(df_final, cand_thor_impute, by='PX_ID')

# use BSA to calculate cardiac output
df_final <- df_final %>%
  ungroup() %>%
  mutate(
    cardiac_output = case_when(
      is.na(cardiac_output) & durable_LVAD == '1' ~ 2.5 * bsa,
      is.na(cardiac_output) & durable_LVAD != '1' ~ 2.2 * bsa,
      TRUE ~ cardiac_output
    )
  )
```

## Define short mcs ever
```{r}
df_final <- df_final %>%
  mutate(
    short_MCS_ever_old = 
         case_when(ECMO_ever == '1' | temp_surg_ever == '1' | BiVAD_no_discharge_ever == '1' ~ 1,
           TRUE ~ 0),
    short_MCS_ever = 
         case_when(ECMO_ever == '1' | temp_surg_ever == '1' | BiVAD_no_discharge_ever == '1' | CAN_ECMO %in% 1 ~ 1,
           TRUE ~ 0))
with(df_final,table(short_MCS_ever_old,short_MCS_ever,useNA="ifany"))
```

## Calculate durable LVAD time
```{r}
df_final = df_final %>% mutate(
  durable_lvad_implant_dt = as.Date(durable_lvad_implant_dt))

df_final = df_final %>% mutate(
  durable_lvad_time = case_when(
    is.na(durable_lvad_implant_dt) & durable_LVAD == 1 ~ as.numeric(current_date - as.Date(can_listing_dt)),
    durable_lvad_implant_dt <= current_date ~ as.numeric(current_date - durable_lvad_implant_dt),
    TRUE ~ 0
  )
)
df_final  %>% filter(durable_LVAD == 1) %>% select(durable_lvad_implant_dt, current_date, can_listing_dt, durable_lvad_time)


### from mortality_model_data.Rmd
### fix death to be pretransplant only and just one row ###

# fix extra rows w/ outcome=1
table(df_final$outcome)
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    T ~ 0))
table(df_final$outcome)

# recode outcome as 0 if patient received transplant
## want outcome to be an indicator of pretx death
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(any_tx=any(transplant==1)) %>%
  ungroup() %>%
  mutate(outcome = ifelse(any_tx,0,outcome))
table(df_final$outcome)

# date for end of interval
# df_final = df_final %>% mutate(
#   current_date_end = current_date + pmin(last_active_date - current_date, 7)
# )

# time on LVAD at current_date
df_final = df_final %>% mutate(
  durable_LVAD = case_when(
    durable_lvad_time > 0 ~ 1,
    T ~ durable_LVAD
  ),
  durable_lvad_time = case_when(
    durable_LVAD == 1 & durable_lvad_implant_dt > current_date ~ as.numeric(current_date - as.Date(can_listing_dt)),
    TRUE ~ durable_lvad_time
  ),
  lvad_years = durable_lvad_time/365.25
)
```

## Bring in justification form data
```{r excep}
library(haven)

## note: only works for Grace
curr_data_release = "//srtr-fs-p01/SAF/Internal/SAF2512/thor/ThoracicRegistration"
just_form_hr = read_sas(here::here(curr_data_release, "JustFormHR.sas7bdat"), NULL) %>%  
    zap_formats() %>% zap_labels()
just_form_hr_data_link = 
  read_sas(here::here(curr_data_release, "JustFormHRDataLink.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
risk_strat_data_hr = 
  read_sas(here::here(curr_data_release, "RiskStratDataHR.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
status_just_episode = 
  read_sas(here::here(curr_data_release, "StatusJustEpisode.sas7bdat"), NULL) %>%
  zap_formats() %>% zap_labels()
just_form_hr_stat1 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat1.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
just_form_hr_stat2 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat2.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
just_form_hr_stat3 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat3.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
just_form_hr_stat4 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat4.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
thor_support_device = 
    read_sas(here::here(curr_data_release, "ThorSupportDevice.sas7bdat"), NULL) %>% 
    zap_formats() %>% zap_labels()

# MOLLY ADDED - data loading
# just_form_hr = read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/JustFormHR.sas7bdat") %>%  
#     zap_formats() %>% zap_labels()
# just_form_hr_data_link = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/JustFormHRDataLink.sas7bdat") %>%  
#   zap_formats() %>% zap_labels()
# risk_strat_data_hr = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/RiskStratDataHR.sas7bdat") %>%  
#   zap_formats() %>% zap_labels()
# status_just_episode = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/StatusJustEpisode.sas7bdat") %>%
#   zap_formats() %>% zap_labels()
# just_form_hr_stat1 = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/JustFormHRStat1.sas7bdat") %>%  
#   zap_formats() %>% zap_labels()
# just_form_hr_stat2 = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/JustFormHRStat2.sas7bdat") %>%  
#   zap_formats() %>% zap_labels()
# just_form_hr_stat3 = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/JustFormHRStat3.sas7bdat") %>%  
#   zap_formats() %>% zap_labels()
# just_form_hr_stat4 = 
#   read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/JustFormHRStat4.sas7bdat") %>%  
#   zap_formats() %>% zap_labels()
# thor_support_device = 
#     read_sas("C:/Users/mwhite10/Downloads/Supplement2603/Thoracic Registration/ThorSupportDevice.sas7bdat") %>% 
#     zap_formats() %>% zap_labels()

# join everything
justdat = inner_join(risk_strat_data_hr %>% select(px_id,wl_id,WlregAuditId),
                     just_form_hr_data_link %>% select(WlregAuditId,JustId),
                     by="WlregAuditId") %>%
  inner_join(.,
             status_just_episode %>% select(JustId,BeginDate,EndDate),
             by="JustId") %>%
  inner_join(.,
             just_form_hr %>% select(JustId,RequestedCandStatCd,Exception,ThorSupportDevicePrimary,ThorSupportDeviceSecondary),
             by="JustId") %>%
  left_join(.,
             bind_rows(just_form_hr_stat2 %>% select(JustId,CriteriaIabpSupport),
                       just_form_hr_stat3 %>% select(JustId,CriteriaIabpSupport)),
             by="JustId") %>%
  left_join(.,thor_support_device %>% select(DeviceId,DeviceTypeId,VadBrandId,TahBrandId,PercuBrandId,VentricularSupportId,ImplantDt,OtherSpecify),
            by=c("ThorSupportDevicePrimary"="DeviceId")) %>%
  mutate(BrandId = coalesce(VadBrandId,TahBrandId,PercuBrandId)) %>%
  # EndDate is missing for people who were still at that status when data were pulled
  mutate(EndDate = case_when(is.na(EndDate) ~ max(EndDate,na.rm=T),T ~ EndDate))

# look at Impellas in OtherSpecify
thor_support_device %>% filter(grepl("imp",OtherSpecify,ignore.case=T)) %>% count(OtherSpecify) %>% print(n=100)

# special dataset for iabp and impella
# 237=Impella CP, 238=Impella RP, 239=Impella 5.5, 225=Impella Recover 2.5, 226=Impella Recover 5.0
justdat_device = justdat %>%
  mutate(iabp = case_when(DeviceTypeId %in% 8 ~ 1,T ~ 0),
         impella = case_when(BrandId %in% c(237,238,239,225,226) ~ 1,T ~ 0),
         impella = case_when(grepl("imp",OtherSpecify,ignore.case=T) ~ 1,T ~ impella)) %>%
  arrange(px_id,BeginDate,EndDate) %>%
  group_by(px_id) %>%
  mutate(iabp_ever = ifelse(cumsum(iabp)>0,1,0),
         impella_ever = ifelse(cumsum(impella)>0,1,0)) %>%
  mutate(rown=row_number()==n()) %>%
  ungroup() %>%
  # define a BeginDate-EndDate range until the next form (including inactive time and status 5-6)
  mutate(EndDate = case_when(!rown ~ lead(BeginDate),
                             T ~ EndDate))
  
# keep justification forms for patients in study, on current_date (end of day)
justdat_join = justdat %>% right_join(.,df_final %>% select(PX_ID,current_date),by=c("px_id"="PX_ID")) %>%
  mutate(current_date_yms=current_date+days(1)-seconds(1)) %>% 
  filter(current_date_yms>=BeginDate,current_date_yms<=EndDate) %>% distinct # note: distinct not needed

# keep devices from forms for patients in study, on current_date (end of day)
justdat_device_join = justdat_device %>% right_join(.,df_final %>% select(PX_ID,current_date),by=c("px_id"="PX_ID")) %>%
  mutate(current_date_yms=current_date+days(1)-seconds(1)) %>% 
  filter(current_date_yms>=BeginDate,current_date_yms<=EndDate) %>% distinct # note: distinct not needed
justdat_device_join %>% count(iabp,iabp_ever,impella,impella_ever)

# join justification forms to df_final
names(justdat_join)[names(justdat_join) %in% names(df_final)]
names(justdat_device_join)[names(justdat_device_join) %in% names(df_final)]
df_final2 = df_final %>% rename(JustId_old=JustId,RequestedCandStatCd_old=RequestedCandStatCd,Exception_old=Exception) %>%
  left_join(.,justdat_join,by=c("PX_ID"="px_id","current_date")) %>%
  left_join(.,justdat_device_join %>% select(px_id,current_date,iabp,impella,iabp_ever,impella_ever),
            by=c("PX_ID"="px_id","current_date"))

## sanity checks
nrow(df_final2);nrow(df_final)
df_final2 %>% count(status,RequestedCandStatCd)
df_final2 %>% count(status,Exception)
df_final2 %>% count(iabp,impella)
df_final2 %>% count(iabp,iabp_ever,impella,impella_ever)
df_final2 %>% filter(is.na(iabp),is.na(impella)) %>% count(status)

# people missing both iabp and impella are people w/ no previous Status 1-4 form
justdat %>% right_join(.,df_final2 %>% filter(is.na(iabp),is.na(impella)) %>%
                                select(PX_ID,current_date_yms),by=c("px_id"="PX_ID")) %>%
  filter(BeginDate <= current_date_yms) %>%
  nrow

# recode missing iabp and impella as 0
df_final2 = df_final2 %>%
  mutate(iabp=ifelse(is.na(iabp),0,iabp),
         iabp_ever=ifelse(is.na(iabp_ever),0,iabp_ever),
         impella=ifelse(is.na(impella),0,impella),
         impella_ever=ifelse(is.na(impella_ever),0,impella_ever))
df_final2 %>% count(iabp,iabp_ever,impella,impella_ever)
df_final2 %>% count(status,iabp,impella) %>% print(n=30) # carry forward into next status, even if inactive, 5, or 6

# save df_final2 as df_final
df_final=df_final2
rm(df_final2)
rm(just_form_hr,just_form_hr_data_link,risk_strat_data_hr,status_just_episode,just_form_hr_stat1,just_form_hr_stat2,just_form_hr_stat3,just_form_hr_stat4,thor_support_device,justdat,justdat_device,justdat_join,justdat_device_join)
gc()
```

# Calculate diagnosis group variables
```{r diag}
# diagnosis group - CHD
  ## the four big ones including CHD
  # restrictive cardiomyopathy are ineligible to receive certain devices
# definitely include retransplant
table(df_final$diagnosis,useNA="ifany")

# read in SAF to get CAN_DGN
## only works for Grace
curr_data_release = "//srtr-fs-p01/SAF/Internal/SAF2512/thor"
df_cand_thor <- haven::read_sas(file.path(curr_data_release,'cand.sas7bdat'),
                                  col_select=c(PX_ID, CAN_DGN, CAN_DGN2))
# MOLLY ADDED - read in SAF
# df_cand_thor <- haven::read_sas("C:/Users/mwhite10/Downloads/pubsaf2603/cand_thor.sas7bdat",
#                                   col_select=c(PX_ID, CAN_DGN, CAN_DGN2))
df_final = left_join(df_final,df_cand_thor,by="PX_ID")

# look at CAN_DGN for each diagnosis
df_final %>% filter(CAN_DGN %in% c(1208,1209,1100,1199)) %>% count(diagnosis,CAN_DGN)

# recode 1100 and 1199 as Re-Transplant
df_final = df_final %>%
  mutate(diagnosis = case_when(CAN_DGN %in% c(1100,1199) ~ "Re-Transplant",
                               T ~ diagnosis))
table(df_final$diagnosis,useNA="ifany")

# recode 1208 and 1209 as Dilated_CM
df_final = df_final %>%
  mutate(diagnosis = case_when(CAN_DGN %in% c(1208,1209) ~ "Dilated_CM",
                               T ~ diagnosis))
table(df_final$diagnosis,useNA="ifany")

# look at CAN_DGN for each diagnosis
df_final %>% filter(CAN_DGN %in% c(1208,1209,1100,1199)) %>% count(diagnosis,CAN_DGN)

# confirm no missing CAN_DGN being coded as Other
df_final %>% filter(is.na(CAN_DGN)) %>% nrow

```

## Split data into training / testing sets
All 1-week intervals that began between 1/1/2019 and 6/30/2023, for candidates listed by 5/15/2023, were retained in the model training data, while all other 1-week intervals were held out in validation and test sets for model assessment.
```{r}
# confirm listing dates and current dates
with(df_final,tapply(can_listing_dt,train_test_split,summary))
with(df_final,tapply(current_date,train_test_split,summary))

#temporally split training and testing sets
df_final_FULL = df_final
df_final = df_final %>% filter(train_test_split == "train")
```

## MICE
```{r}
## need to impute labs and hemos:
df_final %>% select(albumin,bilirubin,sodium,BNP,BNP_NT_Pro,
                    durable_LVAD,short_MCS_ever,status,
                    CAN_GENDER,creatinine,age_at_listing,
                    cardiac_output,systolicBP,diastolicBP,PCWP,
                    PASP,PADP,central_venous_pressure,
                    impella,iabp,impella_ever,iabp_ever,diagnosis,lvad_years) %>%
  apply(.,2,function(col) sum(is.na(col)))





```


## Calculate eGFR
```{r}
df_final <- df_final %>%
  mutate(eGFR = case_when(
    CAN_GENDER == 'F' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.7), 1)^(-0.241)) *
      (pmax((creatinine / 0.7), 1)^(-1.2)) * 0.9938^(age_at_listing) * 1.012,
    
    CAN_GENDER == 'M' & !is.na(creatinine) ~ 142 * (pmin((creatinine / 0.9), 1)^(-0.302)) *
      (pmax((creatinine / 0.9), 1)^(-1.2)) * 0.9938^(age_at_listing))
  )
```

## Define hemodynamics
```{r}
# define hemodynamics
df_final <- df_final %>%
  mutate(
    cpo = (1/451) * cardiac_output * (((1/3)*systolicBP) + ((2/3)*diastolicBP)),
    api = (systolicBP - diastolicBP) / PCWP,
    papi = (PASP - PADP) / central_venous_pressure) %>%
  
  mutate(cpo = as.numeric(cpo),
         api = as.numeric(api),
         papi = as.numeric(papi)) %>%
  mutate(api = case_when(PCWP == 0 ~ 1, T ~ api),
         papi = case_when(
            central_venous_pressure == 0 ~ 1,
            papi < 1 & !is.na(papi) ~ 1,
            TRUE ~ papi
    ))


# fix API when negative (recode as missing)
summary(df_final$api)
df_final = df_final %>%
  mutate(api = ifelse(api<0,NA,api),
         api = ifelse(is.na(api),median(api,na.rm=T),api))
summary(df_final$api)
```

## Save discrete time data
```{r}
# check numbers
dim(table(df_final$PX_ID)) # 33961
table(df_final$transplant) # 23928 

# save dataset
save("df_final", file =  "discrete_final_data_MICE.RData")
```