---
title: "mortality model assessment"
output: html_document
date: "2024-07-15"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list = ls())
# load("C:/Users/mhwhi/Downloads/mortality-model.RData")
# load("C:/Users/mhwhi/Downloads/mortality_model_test_data.RData")
# Grace
base::load("mortality-model-options.Rdata")
base::load("mortality-model-lasso-options.Rdata")
base::load("mortality_model_test_data.RData")
library(rms)
library(tidyr)
library(dplyr)
library(haven)
library(tidyr)
library(purrr)
library(riskRegression)
library(survival)
library(ggplot2)
library(splines)
```

```{r}
test_data = test_data %>% mutate(can_listing_dt = as.Date(can_listing_dt),
                                 transplant_date = as.Date(transplant_date))

table(test_data$status,useNA="ifany")
```

# Code waiting time at status
```{r}
# wait time determined by status
test_data = test_data %>%
  mutate(wait_time = case_when(
    status == 1 ~ wait_1,
    status == 2 ~ wait_2,
    status == 3 ~ wait_3,
    status == 4 ~ wait_4,
    status == 5 ~ wait_5,
    status == 6 ~ wait_6
  ))


```

# Rename variables "at_start" for prediction
```{r}
#start variables for the model testing
test_data = test_data %>% mutate(
  bilirubin_at_start = bilirubin,
  albumin_at_start = albumin,
  eGFR_at_start = eGFR,
  BNP_NT_Pro_at_start = ifelse(BNP_NT_Pro == 1, as.character(1), as.character(0)),
  BNP_at_start = BNP,
  sodium_at_start = sodium,
  durable_LVAD_at_start = durable_LVAD,
  short_mcs_ever_at_start = short_MCS_ever,
  cpo_at_start = cpo,
  api_at_start = api,
  papi_at_start = papi,
  lvad_years_at_start = lvad_years,
  impella_at_start = impella,
  iabp_at_start = iabp,
  iabp_ever_at_start = iabp_ever,
  impella_ever_at_start = impella_ever,
  diagnosis_at_start = diagnosis
)
```

# Calculate time to event
Also: Remove inactive rows and trim continuous variables, for prediction.
```{r}
# Time to event based on death, removal for deterioration, tx, or censoring

test_data = test_data %>% mutate(
  time_to_death = case_when(
    !is.na(death_time) ~ death_time - interval_start,
    TRUE ~ NA
  ),
  time_to_removal_deteriorate = case_when(
    !is.na(CAN_REM_CD) & CAN_REM_CD == 13 ~ removal_time - interval_start,
    TRUE ~ NA
  ),
  time_to_transplant = as.numeric(case_when(
    any(transplant == 1) ~ as.Date(transplant_date) - current_date,
    TRUE ~ NA
  ))
) 

test_data = test_data %>% 
  mutate(
    time = pmin(time_to_transplant, time_to_death, time_to_removal_deteriorate, 42, na.rm = T)
  )

# define follow up time in weeks, outcome dependent on whether it occurred in interval
test_data = test_data %>% 
  mutate(
    follow_up_int = case_when(
      time == time_to_transplant ~ floor(time/7),
      time == time_to_death ~ ceiling(time/7),
      time == time_to_removal_deteriorate ~ ceiling(time/7),
      TRUE ~ 6
    ),
    outcome = case_when(
      time == time_to_transplant ~ 0,
      time == time_to_death ~ 1,
      time == time_to_removal_deteriorate ~ 1,
      TRUE ~ 0
    )
  )

#filter out inactives and post delisting dates
test_data_actives = test_data %>% filter(!(status %in% c("inactive", "off_waitlist")) & !is.na(status))

# truncate at cutoffs from modeling data
trim=read.csv("trim_values.csv") %>%
  mutate(var=paste0(var,"_at_start"))
test_data_actives = test_data_actives %>%
  mutate(across(c(albumin_at_start,bilirubin_at_start,eGFR_at_start,sodium_at_start,cpo_at_start,api_at_start,papi_at_start), ~ {
    var <- cur_column()
    min_val <- trim$min[trim$var == var]
    max_val <- trim$max[trim$var == var]
    pmin(pmax(., min_val), max_val)
  })) %>%
  mutate(lvad_years_at_start = ifelse(lvad_years_at_start>10,10,lvad_years_at_start)) %>%
  mutate(BNP_at_start = case_when(
           BNP_NT_Pro_at_start==0 & BNP_at_start<trim$min[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==0] ~ trim$min[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==0],
           BNP_NT_Pro_at_start==0 & BNP_at_start>trim$max[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==0] ~ trim$max[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==0],
           BNP_NT_Pro_at_start==1 & BNP_at_start<trim$min[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==1] ~ trim$min[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==1],
           BNP_NT_Pro_at_start==1 & BNP_at_start>trim$max[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==1] ~ trim$max[trim$var=="BNP_at_start" & trim$BNP_NT_Pro_at_start==1],
           T ~ BNP_at_start)
  )

```

# Create nonoverlapping test intervals by calendar date
The validation dataset is from 9/1/23-8/31/24, so intervals should start on: 9/1/23, 10/15/23, 12/1/23, 1/15/24, 3/1/24, 4/15/24, 6/1/24, 7/15/24.

The test dataset is from 9/1/24-8/31/25, so intervals should start on: 9/1/24, 10/15/24, 12/1/24, 1/15/25, 3/1/25, 4/15/25, 6/1/25, 7/15/25.

```{r}
library(lubridate)
valid_dates = c(ymd("2023-09-01"),ymd("2023-10-15"),ymd("2023-12-01"),ymd("2024-01-15"),ymd("2024-03-01"),ymd("2024-04-15"),ymd("2024-06-01"),ymd("2024-07-15"))
test_dates = c(ymd("2024-09-01"),ymd("2024-10-15"),ymd("2024-12-01"),ymd("2025-01-15"),ymd("2025-03-01"),ymd("2025-04-15"),ymd("2025-06-01"),ymd("2025-07-15"))

# Subset data at certain calendar dates
# Calculate wait time at specific calendar date
# Calculate "rank" based on status (1-6) and waiting time
  ## If status and wait time are the same, set the rank to be equal

make_eight_ints = function(dates) {
  if (length(dates)!=8) {
    stop("dates vector does not have length 8")
  }
  
  first_int = test_data_actives %>% 
    filter(current_date <= dates[1] & current_date > (dates[1]-days(7))) %>% 
    mutate(int = 1,
           wait_time = wait_time + as.numeric(dates[1] - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup
  
  second_int = test_data_actives %>% 
    filter(current_date <= dates[2] & current_date > (dates[2]-days(7))) %>% 
    mutate(int = 2,
           wait_time = wait_time + as.numeric(as.Date("2023-02-15") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup  
  
  third_int = test_data_actives %>% 
    filter(current_date <= dates[3] & current_date > (dates[3]-days(7))) %>% 
    mutate(int = 3,
           wait_time = wait_time + as.numeric(as.Date("2023-04-01") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup 
  
  fourth_int = test_data_actives %>% 
    filter(current_date <= dates[4] & current_date > (dates[4]-days(7))) %>% 
    mutate(int = 4,
           wait_time = wait_time + as.numeric(as.Date("2023-05-15") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup 
  
  fifth_int = test_data_actives %>% 
    filter(current_date <= dates[5] & current_date > (dates[5]-days(7))) %>% 
    mutate(int = 5,
           wait_time = wait_time + as.numeric(as.Date("2023-07-01") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup 
  
  sixth_int = test_data_actives %>% 
    filter(current_date <= dates[6] & current_date > (dates[6]-days(7))) %>% 
    mutate(int = 6,
           wait_time = wait_time + as.numeric(as.Date("2023-08-15") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup 
  
  seventh_int = test_data_actives %>% 
    filter(current_date <= dates[7] & current_date > (dates[7]-days(7))) %>% 
    mutate(int = 7,
           wait_time = wait_time + as.numeric(as.Date("2023-10-01") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup  
  
  eighth_int = test_data_actives %>% 
    filter(current_date <= dates[8] & current_date > (dates[8]-days(7))) %>% 
    mutate(int = 8,
           wait_time = wait_time + as.numeric(as.Date("2023-11-15") - current_date)) %>%
    ungroup() %>% arrange(status, desc(wait_time)) %>% 
    mutate(rank = row_number()) %>% 
    group_by(status, wait_time) %>%
    mutate(rank = first(rank)) %>% ungroup 
  
  
  # Combine into dataset for pooled AUC
  eight_sets = rbind(first_int, second_int, third_int, fourth_int, fifth_int, sixth_int, seventh_int, eighth_int)
  return(eight_sets)
}
eight_sets_valid = make_eight_ints(valid_dates)
eight_sets_test = make_eight_ints(test_dates)

# check coding
eight_sets_valid %>%
  filter(int==1) %>%
  filter(status == 4) %>% arrange(desc(wait_time)) %>% 
  select(PX_ID, status, wait_time, rank) %>% head
```

# Function to calculate US-CRS 1.0 log odds
```{r uscrs1}
uscrs1 = function(albumin,bilirubin,eGFR,sodium,lvad,mcs,BNP,NTpro) {
  return(-.656*albumin
         +.617*log(bilirubin+1) 
         -.012*eGFR
         -.077*sodium
         -.377*lvad
         +1.092*mcs
         +.433*log(BNP+1)*(NTpro==0)
         +.224*log(BNP+1)*(NTpro==1)
         )
}

```

# Flip rankings so that higher = higher risk
```{r}
# Flip the predictors to be higher when the candidate is higher risk of death

eight_sets_valid = eight_sets_valid %>% 
  ungroup() %>% mutate(
  rank = 1/rank,
  uscrs = uscrs1(albumin,bilirubin,eGFR,sodium,durable_LVAD,short_MCS_ever,BNP,BNP_NT_Pro),
  uscrs_trim = uscrs1(albumin_at_start,bilirubin_at_start,eGFR_at_start,sodium_at_start,durable_LVAD_at_start,short_mcs_ever_at_start,BNP_at_start,BNP_NT_Pro_at_start),
  status_flipped = 1/as.numeric(status)
)
eight_sets_test = eight_sets_test %>% 
  ungroup() %>% mutate(
  rank = 1/rank,
  uscrs = uscrs1(albumin,bilirubin,eGFR,sodium,durable_LVAD,short_MCS_ever,BNP,BNP_NT_Pro),
  uscrs_trim = uscrs1(albumin_at_start,bilirubin_at_start,eGFR_at_start,sodium_at_start,durable_LVAD_at_start,short_mcs_ever_at_start,BNP_at_start,BNP_NT_Pro_at_start),
  status_flipped = 1/as.numeric(status)
)
```

# Calculate predicted 6-week survival probabilities with US-CRS 2.0
Try out all of the potential models (models A-G) in validation data (`eight_sets_valid`). Try out Model C and Model E (with linear splines) in test data (`eights_sets_test`).
```{r}
# load the function to calculate predicted 1-week hazard from a binomial glm and also calculate 6-week predicted survival
base::load("mort-pred-fun.Rdata")

# apply to eight_sets_valid
eight_sets_valid_modA = pred_data_fun(eight_sets_valid,mortality_modelA)
eight_sets_valid_modB = pred_data_fun(eight_sets_valid,mortality_modelB)
eight_sets_valid_modC = pred_data_fun(eight_sets_valid,mortality_modelC)
eight_sets_valid_modD = pred_data_fun(eight_sets_valid,mortality_modelD)
eight_sets_valid_modE = pred_data_fun(eight_sets_valid,mortality_modelE)
eight_sets_valid_modF = pred_data_fun(eight_sets_valid,mortality_modelF)
eight_sets_valid_modG = pred_data_fun(eight_sets_valid,mortality_modelG)
eight_sets_valid_modH = pred_data_fun(eight_sets_valid,mortality_modelH,lasso=T)

# new variables created by pred_data_fun
names(eight_sets_valid_modA)[!(names(eight_sets_valid_modA)%in%names(eight_sets_valid))]

# save results
save(eight_sets_valid_modA,eight_sets_valid_modB,
     eight_sets_valid_modC,eight_sets_valid_modD,
     eight_sets_valid_modE,eight_sets_valid_modF,
     eight_sets_valid_modG,eight_sets_valid_modH,
     file="eight-sets-valid.Rdata")

# apply to eight_sets_test
eight_sets_test_modC = pred_data_fun(eight_sets_test,mortality_modelC)
base::load("mortality-modelE-linear.Rdata")
eight_sets_test_modE = pred_data_fun(eight_sets_test,mortality_modelE_linear,linear_spline=T)
save(eight_sets_test_modC,eight_sets_test_modE,file="eight-sets-test.Rdata")
```

# Calculate AUC in each interval of validation set and take average
Calculate "Mean 6-Week AUC" in validation dataset (across 8 calendar cross-sections) for each potential model.
```{r}
getMeanAUC = function(moddata) {
  auc_results = sapply(1:dim(table(moddata$int)),function(i) {
    data_i <- moddata %>% filter(int==i)
    auc_tmp <- Score(
      list("uscrs2" = data_i$prob_outcome_6wk),
      formula = Hist(time, outcome) ~ albumin + log(bilirubin + 1) + 
                eGFR + sodium + log(BNP + 1) * BNP_NT_Pro + 
                durable_LVAD + short_MCS_ever + status,
      data = data_i %>% 
                mutate(time = ifelse(outcome == 0, time + 0.001, time)),
      cens.method = "ipcw", cens.model = "cox",
      conf.int = TRUE,
      times = 42,
      metrics = c("auc"),
      conservative = TRUE
    )
    return(auc_tmp$AUC$score[["AUC"]])
  })
  return(mean(auc_results))
}
getMeanAUC(eight_sets_valid_modA)
getMeanAUC(eight_sets_valid_modB)
getMeanAUC(eight_sets_valid_modC)
getMeanAUC(eight_sets_valid_modD)
getMeanAUC(eight_sets_valid_modE)
getMeanAUC(eight_sets_valid_modF)
getMeanAUC(eight_sets_valid_modG)
getMeanAUC(eight_sets_valid_modH)
getMeanAUC(eight_sets_test_modC) # .807
getMeanAUC(eight_sets_test_modE) # .798
```

# Calculate AUC in each interval of test set
```{r}
# create one test set with both model predictions
eight_sets_test_mod = full_join(eight_sets_test_modC %>% select(PX_ID,current_date,int,
                                                                rank,uscrs,uscrs_trim,status_flipped,
                                                                time,outcome,albumin,bilirubin,
                                                                eGFR,sodium,BNP,BNP_NT_Pro,
                                                                durable_LVAD,short_MCS_ever,status,
                                                                prob_outcome_6wk_modC=prob_outcome_6wk,
                                                                haz_dth_modC=haz_dth_nowk),
                                eight_sets_test_modE %>% select(PX_ID,current_date,int,
                                                                prob_outcome_6wk_modE=prob_outcome_6wk,
                                                                haz_dth_modE=haz_dth_nowk),
                                by=c("PX_ID","current_date","int"))
summary(eight_sets_test_mod$prob_outcome_6wk_modC)
summary(eight_sets_test_mod$prob_outcome_6wk_modE)

# initialize lists
auc_results <- list()
p_val <- list()
roc_plot <- list()

# Loop through datasets to compute AUC
for (i in 1:length(test_dates)) {
  data_i <- eight_sets_test_mod %>% filter(int==i)
  auc_tmp <- Score(
    list("uscrs 2.0 simple" = data_i$prob_outcome_6wk_modC, "uscrs 2.0 complex" = data_i$prob_outcome_6wk_modE, "waiting time + status" = data_i$rank, "status only" = data_i$status_flipped, "uscrs 1.0" = data_i$uscrs),
    formula = Hist(time, outcome) ~ albumin + log(bilirubin + 1) + 
              eGFR + sodium + log(BNP + 1) * BNP_NT_Pro + 
              durable_LVAD + short_MCS_ever + status,
    data = data_i %>% 
              mutate(time = ifelse(outcome == 0, time + 0.001, time)),
    cens.method = "ipcw", cens.model = "cox",
    conf.int = TRUE,
    times = 42,
    metrics = c("auc"),
    conservative = TRUE,
    plots = "roc"
  )
  roc_plot[[i]] <- auc_tmp$ROC$plotframe %>% mutate(Date=test_dates[i])
  p_val[[i]] <- auc_tmp$AUC$contrasts
  auc_results[[i]] <- auc_tmp$AUC$score %>%
    mutate(interval = paste0("Interval_", i))  # Add dataset identifier
}
pval_df <- do.call(rbind.data.frame,p_val)
auc_df <- bind_rows(auc_results) %>%
  mutate(Model=factor(model,levels=c("status only","waiting time + status","uscrs 2.0 simple","uscrs 2.0 complex","uscrs 1.0"),labels=c("Status (1-6)","Status (1-6) +\nWaiting Time","US-CRS 2.0 -\nSimple","US-CRS 2.0 -\nComplex","US-CRS 1.0")))
roc_df <- do.call(rbind.data.frame,roc_plot)

# add dates
auc_df$Date <- rep(test_dates, each = dim(table(auc_df$Model)))
pval_df$Date <- rep(test_dates, each = choose(dim(table(auc_df$Model)),2))

# plot results
ggplot(auc_df %>% filter(model!="uscrs 1.0"), 
       aes(x = Date, y = AUC, ymin = lower, ymax = upper, color = Model, shape=Model)) +
  geom_line(position = position_dodge(width = 24),alpha=.85) +
  geom_point(position = position_dodge(width = 24),size=2.5) +
  geom_errorbar(position = position_dodge(width = 24),width=15,alpha=.85) +
  labs(
    #title = "Figure 1: AUC of US-CRS 2.0 compared to\nstatus quo (6-status with waiting time tiebreakers)",
    x = "Cross-section start date",
    y = "Time-dependent AUC (6-week mortality)"
  ) +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = NA), 
    
    # Increase font sizes
    # plot.title = element_text(size = 18),
    # axis.title.x = element_text(size = 16), 
    # axis.title.y = element_text(size = 16),
    
    # axis.text = element_text(size = 10),
    # legend.title = element_text(size = 14),
    # legend.text = element_text(size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    
    text=element_text(size=12),
    legend.key.spacing.y = unit(.5, "lines")
  ) +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_continuous(
    breaks = as.numeric(test_dates),
    labels = format(test_dates, "%m/%d/%Y"),
    minor_breaks=F
  ) +
  scale_shape_manual(values=c(16,17,15,12)) +
  scale_color_viridis_d(direction=-1)
  # scale_color_manual(values = c("darkgrey","darkorange","darkblue")) #+
  # scale_linetype_manual(values = c("dashed","solid","solid"))
ggsave("uscrs_auc_comparison-test.png", width = 7, height = 4, dpi = 300)

# look for significant differences in AUC
pval_df %>% filter(p<0.05)
pval_df %>% filter(model=="uscrs 2.0 simple" | reference=="uscrs 2.0 simple",
                   model!="uscrs 1.0",reference!="uscrs 1.0") %>%
  select(model,reference,delta.AUC,p,Date) %>%
  mutate(signif=p<0.05) %>% arrange(desc(signif))
pval_df %>% filter(model=="uscrs 2.0 complex" | reference=="uscrs 2.0 complex",
                   model!="uscrs 1.0",reference!="uscrs 1.0") %>%
  select(model,reference,delta.AUC,p,Date) %>%
  mutate(signif=p<0.05) %>% arrange(desc(signif))
# the USCRS 2.0 models are basically the same - significantly better than status or status + waiting time on 3/8 dates

# make faceted ROC plots

roc_df = roc_df %>% mutate(Model=factor(model,levels=c("status only","waiting time + status","uscrs 2.0 simple","uscrs 2.0 complex","uscrs 1.0"),labels=c("Status (1-6)","Status (1-6) +\nWaiting Time","US-CRS 2.0 -\nSimple","US-CRS 2.0 -\nComplex","US-CRS 1.0"))) %>% filter(model!="uscrs 1.0")
ggplot(roc_df, 
       aes(x = FPR, y = TPR, color = Model)) +
  geom_line(linewidth=.95,alpha=.85) +
  geom_point(data=roc_df %>% filter(model=="status only"),aes(x=FPR,y=TPR),size=1.75) +
  scale_color_viridis_d(direction=-1) +
  scale_x_continuous(breaks=seq(0,1,.25),labels=c("0",".25",".5",".75","1")) +
  facet_wrap(~Date,nrow=2) +
  labs(x = "False Positive Rate", y = "True Positive Rate", color="Model") +
  theme_bw() + 
  theme(text=element_text(size=12),
        legend.key.spacing.y = unit(.5, "lines")) +
  ylim(c(0,1))
ggsave("uscrs_roc_comparison-test.png", width = 7, height = 4, dpi = 300)
```

<!-- # Calibration -->

```{r calib}
# calib <- Score(
#   list("uscrs" = eight_sets$prob_outcome_6wk),
#   formula = Hist(time, outcome) ~ 1,#albumin + log(bilirubin+1) + eGFR + sodium + log(BNP+1)*BNP_NT_Pro + durable_LVAD + short_MCS_ever + status,
#   data = eight_sets %>% mutate(time = ifelse(outcome==0,time+0.001,time)),
#   # cens.method = "ipcw", cens.model = "cox",
#   # conf.int = TRUE,
#   times = 42,
#   # metrics = c("auc"),
#   conservative = T,
#   plots = "calibration"
# )
# plotCalibration(calib,
#                 xlim=c(0,.1),ylim=c(0,.1),
#                 method="quantile",cens.method="local",q=10,
#                 auc.in.legend=F,brier.in.legend=F)
# plotCalibration(calib,bars=T,ylim=c(0,.1),cens.method="local")
```



```{r}
# Function to calculate AUC using predicted probabilities directly
# calculate_auc = function(data, prob_column) {
#   auc_value <- Score(list(data[[prob_column]]), 
#                      formula = death6week ~ 1,  # Only the response variable
#                      data = data, 
#                      metrics = c("AUC"), 
#                      model = FALSE)  # Set model to FALSE
#   
#   # Extract AUC and CI
#   auc <- auc_value$AUC$score$AUC
#   ci_low <- auc_value$AUC$score$lower  # Lower CI
#   ci_high <- auc_value$AUC$score$upper # Upper CI
#   
#   return(c(AUC = auc, CI_Lower = ci_low, CI_Upper = ci_high))
# }
# 
# # List of datasets
# datasets <- list(
#   first_int = first_int,
#   second_int = second_int,
#   third_int = third_int,
#   fourth_int = fourth_int,
#   fifth_int = fifth_int,
#   sixth_int = sixth_int,
#   seventh_int = seventh_int,
#   eighth_int = eighth_int
# )
# 
# # Specify the column names for the predicted probabilities and rank
# prob_column_name <- "uscrs_score"  # Column for first predictor
# rank_column_name <- "rank"  # Column for second predictor
# 
# # Initialize a data frame to store results
# results <- data.frame(Model = character(), 
#                       AUC = numeric(), 
#                       CI_Lower = numeric(), 
#                       CI_Upper = numeric(),
#                       Predictor = character())  # To differentiate predictors
# 
# # Loop through datasets and compute AUCs for both predictors
# for (i in seq_along(datasets)) {
#   data_name <- paste0("Model ", i)  # Set model name as Model 1, Model 2, etc.
#   data <- datasets[[i]]
#   
#   # Calculate AUC for prob_surv_sweights
#   auc_result_prob <- calculate_auc(data, prob_column_name)
#   results <- rbind(results, data.frame(Model = data_name, 
#                                         AUC = auc_result_prob["AUC"], 
#                                         CI_Lower = auc_result_prob["CI_Lower"], 
#                                         CI_Upper = auc_result_prob["CI_Upper"],
#                                         Predictor = "US-CRS 2.0"))
#   
#   # Calculate AUC for rank
#   auc_result_rank <- calculate_auc(data, rank_column_name)
#   results <- rbind(results, data.frame(Model = data_name, 
#                                         AUC = auc_result_rank["AUC"], 
#                                         CI_Lower = auc_result_rank["CI_Lower"], 
#                                         CI_Upper = auc_result_rank["CI_Upper"],
#                                         Predictor = "Status Quo"))
# }
# 
# # Convert Model to a factor with specified levels for ordering
# results$Model <- factor(results$Model, levels = paste0("Model ", 1:8))
# dates = as.Date(c("2023-01-01", "2023-02-15", "2023-04-01", "2023-05-15", 
#                           "2023-07-01", "2023-08-15", "2023-10-01", "2023-11-15"))
# results$Date <- rep(dates, each = 2)
# 
# 
# # Plot AUC values with confidence intervals
# ggplot(results, aes(x = Date, y = AUC, ymin = CI_Lower, ymax = CI_Upper, color = Predictor)) +
#   geom_point() +
#   geom_line() +
#   geom_errorbar(alpha = 0.3) +
#   labs(title = "AUC Comparison 2023",
#        x = "Date",
#        y = "AUC Value") +
#   theme_minimal() +
#   scale_y_continuous(limits = c(0,1)) + 
#   scale_x_continuous(breaks = as.numeric(dates), 
#                      labels = c("1/1/2023", "2/15/2023", "4/1/2023", "5/15/2023", 
#                                 "7/1/2023", "8/15/2023", "10/1/2023", "11/15/2023")) # Adjust x-axis formatting as needed
```

```{r}
# Load necessary libraries
# library(pROC)
# 
# # Select a single dataset (time point)
# selected_data <- second_int  # Change this to select a different time point
# 
# # Specify the predicted probability column names
# prob_column_name <- "prob_surv_sweights"  # Predictor 1 (US-CRS 2.0)
# rank_column_name <- "rank"  # Predictor 2 (Status Quo)
# 
# # Compute ROC curves for both predictors
# roc_prob <- roc(selected_data$death6week, selected_data[[prob_column_name]])  # US-CRS 2.0
# roc_rank <- roc(selected_data$death6week, selected_data[[rank_column_name]])  # Status Quo
# 
# # Extract ROC data for both models
# roc_data_prob <- data.frame(
#   specificity = rev(roc_prob$specificities),
#   sensitivity = rev(roc_prob$sensitivities),
#   model = "US-CRS 2.0"
# )
# 
# roc_data_rank <- data.frame(
#   specificity = rev(roc_rank$specificities),
#   sensitivity = rev(roc_rank$sensitivities),
#   model = "Status Quo"
# )
# 
# # Combine data for both models
# roc_data <- rbind(roc_data_prob, roc_data_rank)
# 
# # Plot ROC curves using ggplot2
# ggplot(roc_data, aes(x = 1 - specificity, y = sensitivity, color = model)) +
#   geom_line(size = 1) +  # ROC curves
#   geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +  # Diagonal line
#   labs(title = "ROC Curve on Feb 15, 2023",
#        x = "False positive rate",
#        y = "True positive rate",
#        color = "Predictor") +
#   scale_color_manual(values = c("blue", "red")) +  # Custom colors for differentiation
#   theme_minimal()

```

```{r}
# getHR = function(mod,vars,comb=rep(1,length(vars)),cov=vcov(mod),alpha=0.05) {
#   # get beta from mod object
#   beta = coef(mod)
#   beta[is.na(beta)] = 0
#   
#   # get combination vector
#   c = rep(0,length(beta))
#   names(c) = names(beta)
#   c[vars] = comb
#   
#   # compute the HR
#   xb = c %*% beta
#   hr = exp(xb)
#   
#   if (!all.equal(names(c),colnames(cov))) {
#     stop("Mismatched names bw beta and var-cov matrix")
#   }
#   
#   # compute the 95% CI
#   se = sqrt(t(c) %*% cov %*% c)
#   lowCI = exp(xb - qnorm(1-alpha/2)*se)
#   highCI = exp(xb + qnorm(1-alpha/2)*se)
#   p = pnorm(abs(xb/se),lower.tail=F)*2
#   return(data.frame(hr,lowCI,highCI,p))
# }
```


```{r}
# library(dplyr)
# library(splines)
# library(tidyr)
# library(ggplot2)
# library(rms)
# library(survey) # for svydesign and svyCreateTableOne
# library(tableone)
# library(sandwich)
# library(ggplot2)
# library(dplyr)
# 
# # Summary with robust CIs
# mortality_model_sweights_VCOV = vcovCL(mortality_model_sweights_lvad_time_lvad_time, cluster = ~ PX_ID)
# mortality_model_sweights_lvad_sum = do.call(rbind.data.frame, lapply(names(coef(mortality_model_sweights_lvad_time_lvad_time)), function(var_interest) {
#   bind_cols(var = var_interest, getHR(mortality_model_sweights_lvad_time_lvad_time, vars = var_interest, cov = mortality_model_sweights_VCOV))
# }))
# 
# # Filter out intercept
# plotdat = mortality_model_sweights_lvad_sum %>%
#   filter(var != "(Intercept)") %>%
#   mutate(model = "Stabilized Weights + LVAD Time")
# 
# plotdat = plotdat %>% filter(!grepl("rcs", var))
# # Hazard Ratio Forest Plot
# ggplot(plotdat, aes(x = hr, y = var)) +
#   geom_point(position = position_dodge(width = 0.5)) +
#   geom_errorbarh(aes(xmin = lowCI, xmax = highCI), height = 0.2, position = position_dodge(width = 0.5)) +
#   labs(x = "Hazard Ratio (95% CI)", y = "Term", title = "Model Hazard Ratios") +
#   geom_vline(xintercept = 1, linetype = "dashed") 
# 
# # Save plot
# ggsave("stabilized_weights_lvad_time_HR.png", height = 7, width = 16)
# plotdat
```
```{r}
# library(dplyr)
# library(ggplot2)
# library(splines) # for cubic splines
# library(sandwich) # for robust standard errors
# library(lmtest) # for coefficient testing
# 
# # Step 1: Define Reference Value for LVAD Years
# ref_lvad_years <- 0 # Baseline category (e.g., 0 years)
# lvad_years_values <- c(0, 1, 2, 5)  # Define key points for estimation
# 
# # Step 2: Create a Data Frame for Prediction
# newdata <- expand.grid(
#   albumin_at_start = median(combined_df$albumin_at_start, na.rm = TRUE),
#   bilirubin_at_start = median(combined_df$bilirubin_at_start, na.rm = TRUE),
#   eGFR_at_start = median(combined_df$eGFR_at_start, na.rm = TRUE),
#   BNP_at_start = median(combined_df$BNP_at_start, na.rm = TRUE),
#   BNP_NT_Pro_at_start = median(combined_df$BNP_NT_Pro_at_start, na.rm = TRUE),
#   sodium_at_start = median(combined_df$sodium_at_start, na.rm = TRUE),
#   durable_LVAD_at_start = median(combined_df$durable_LVAD_at_start, na.rm = TRUE),
#   short_mcs_ever_at_start = median(combined_df$short_mcs_ever_at_start, na.rm = TRUE),
#   cpo_at_start = median(combined_df$cpo_at_start, na.rm = TRUE),
#   api_at_start = median(combined_df$api_at_start, na.rm = TRUE),
#   papi_at_start = median(combined_df$papi_at_start, na.rm = TRUE),
#   lvad_years_at_start = lvad_years_values # Vary only this variable
# )
# 
# # Step 3: Predict HRs from Model
# predictions <- predict(mortality_model_sweights_lvad_time_lvad_time, newdata, type = "link", se.fit = TRUE)
# 
# # Step 4: Convert to Hazard Ratios
# newdata <- newdata %>%
#   mutate(
#     log_hr = predictions$fit,
#     hr = exp(log_hr),
#     lowCI = exp(log_hr - 1.96 * predictions$se.fit),
#     highCI = exp(log_hr + 1.96 * predictions$se.fit)
#   )
# 
# # Step 5: Normalize Against Reference (LVAD Years = 0)
# ref_hr <- newdata %>% filter(lvad_years_at_start == ref_lvad_years) %>% pull(hr)
# newdata <- newdata %>%
#   mutate(
#     norm_hr = hr / ref_hr,
#     norm_lowCI = lowCI / ref_hr,
#     norm_highCI = highCI / ref_hr
#   )
# 
# newdata = newdata %>% filter(lvad_years_at_start != 0)
# # Step 6: Plot the Forest Plot
# ggplot(newdata, aes(x = norm_hr, y = as.factor(lvad_years_at_start))) +
#   geom_point(position = position_dodge(width = 0.5)) +
#   geom_errorbarh(aes(xmin = norm_lowCI, xmax = norm_highCI), height = 0.2, position = position_dodge(width = 0.5)) +
#   labs(x = "Hazard Ratio (95% CI)", y = "LVAD Years", title = "Hazard Ratio for LVAD Years") +
#   geom_vline(xintercept=1, linetype="dashed") 
# 
# 

```
```{r}
# newdata
# newdata = newdata %>% select(lvad_years_at_start, norm_hr, norm_lowCI, norm_highCI) %>%
#   rename(hr = norm_hr, lowCI = norm_lowCI, highCI = norm_highCI, var = lvad_years_at_start)
# newdata = newdata %>% mutate(
#     var = as.character(var)
#   )
# plotdat = plotdat %>% select(var, hr, lowCI, highCI)
# forest_data = bind_rows(newdata, plotdat) %>% mutate(
#   var = case_when(
#     var == "1" ~ "LVAD Years = 1 [v. 0]",
#     var == "2" ~ "LVAD Years = 2 [v. 0]",
#     var == "5" ~ "LVAD Years = 5 [v. 0]",
#     var == "albumin_at_start" ~ "Albumin",
#     var == "log(bilirubin_at_start + 1)" ~ "Log Bilirubin",
#     var == "eGFR_at_start" ~ "eGFR",
#     var == "sodium_at_start" ~ "Sodium",
#     var == "durable_LVAD_at_start" ~ "Durable LVAD",
#     var == "short_mcs_ever_at_start" ~ "Short-term MCS",
#     var == "cpo_at_start" ~ "CPO",
#     var == "api_at_start" ~ "API",
#     var == "papi_at_start" ~ "PAPI",
#     var == "log(BNP_at_start + 1):BNP_NT_Pro_at_start0" ~ "Log BNP, Regular BNP",
#     var == "log(BNP_at_start + 1):BNP_NT_Pro_at_start1" ~ "Log BNP, NT-Pro BNP"
#   )
# )
# 
# forest_data
```
```{r}
# 
# ggplot(forest_data, aes(x = hr, y = var)) +
#   geom_point(position = position_dodge(width = 0.5), size = 2) +
#   geom_errorbarh(aes(xmin = lowCI, xmax = highCI), height = 0.3, position = position_dodge(width = 0.5)) +
#   labs(x = "Hazard Ratio (95% CI)", y = "Term", title = "Model Hazard Ratios") +
#   geom_vline(xintercept = 1, linetype = "dashed")  +
#   theme_minimal() +  # Clean theme
#   theme(
#     plot.title = element_text(size = 18),  # Title larger & bold
#     axis.title.x = element_text(size = 16),  # X-axis label
#     axis.title.y = element_text(size = 16),  # Y-axis label
#     axis.text.x = element_text(size = 14),  # X-axis tick labels
#     axis.text.y = element_text(size = 14)# Legend title size
#   )
# 
# ggsave("uscrs_forest_ro1.png", width = 12, height = 7)
```

