---
title: "mortality_model_dataset"
output: html_document
date: "2024-06-24"
editor_options: 
  chunk_output_type: console
---

## Load data and libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# rm(list=ls())
load("discrete_final_data_MICE.RData") # df_final - created in discrete_data_prep.Rmd
# load("df_intervals_setup.RData") # df_intervals - US-CRS 1.0
library(dplyr)
library(splines)
library(tidyr)
library(ggplot2)
library(rms)
library(survey) # for svydesign and svyCreateTableOne
library(tableone) # for CreateTableOne
library(haven)
library(lubridate)
library(sandwich)
```


# Fit transplant models for IPW
The denominator of the inverse probability weights is a function of the time-varying probability of transplant. We estimated this probability by a discrete-time survival model (ie, logistic regression) for time to transplant, using the training discrete-time dataset from listing to removal and allowing patient characteristics to update at the beginning of each week.

We are fitting this model in the training data, even though no mortality information are used, because transplant probabilities shift over time and we have data over a long time period, from 2019 through August 2025. Weights for model evaluation should therefore be based on transplant models fit to the validation and/or test data.

We considered four specifications of the model for the denominator of the weights:

   - US-CRS 1.0 variables
   - US-CRS 1.0 variables + status
   - US-CRS 1.0 variables + status + hemodynamics
   - US-CRS 1.0 variables + status + hemodynamics + time since listing
   
The denominator weights should adjust for all time-varying confounders of transplant and mortality, and balance the distribution of confounders at each timepoint. Status was considered a potential confounder because it is a strong predictor of transplant under the current heart allocation system and may provide additional information about mortality.

We then assessed covariate balance after weighting each week of the discrete-time dataset by the inverse probability of receiving the observed treatment (transplant or no transplant) that week. The weighted distributions of all variables (US-CRS 1.0 variables + status + hemodynamics + time since listing) were compared between the groups "transplant" and "no transplant", where patients could contribute to both groups, joining the "transplant" group on their week of receiving transplant.

The model with US-CRS 1.0 variables + status + hemodynamics had the best covariate balance.

## Compare denominator models
```{r}
## Model with US-CRS 1.0 vars, status, and hemodynamics
tx_status_hemo_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status + ns(cpo, df = 3) + ns(api, df = 3) + ns(papi, df = 3)), data = train_data, family = binomial)
tx_status_hemo_model
summary(tx_status_hemo_model)

## calculate weights for covariate balance assessment ##

## Predict probability of transplant
train_data$fit_prob_hemo <- predict(tx_status_hemo_model, newdata=train_data, type = "response")

## Construct weight variables for covariate balance assessment
train_data <- train_data %>%
  mutate(weight_hemo = if_else(transplant == 1, 1 / fit_prob_hemo, 1 / (1 - fit_prob_hemo)))

# with chosen model, set probability of not receiving a transplant each week
train_data = train_data %>%
  mutate(prob_no_tx = 1 - fit_prob_hemo)
```

## Fit numerator model for stabilized weights
For stabilized weights, we also need to model the time-varying probability of transplant based on variables at baseline, which is, for us, at the start of a 6-week window. Any variables in the numerator model have to be included in the marginal structural model, so we restricted to the variables in US-CRS 1.0 (the base model) for this step. Again, we estimated this probability by a discrete-time survival model (ie, logistic regression) for time to transplant, using the training discrete-time dataset from listing to removal, but divided into non-overlapping 6-week windows, with predictors collected at the start of the window.

Note that rows after removal/transplant are not needed for modeling the probability of transplant, because after removal, the probability of transplant is 0.
```{r}

## create numerator of stabilized weights ##
# we want Pr(transplant) given variables at start of 6-week window
# for simplicity (and to avoid correlated data), we will model this in non-overlapping 6-week windows of train_data

# create 6-week windows of train_data
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(window=ceiling(row_number()/6),
         week_from_baseline = (row_number() - 6*(window-1))) %>%
  ungroup()
# train_data %>% select(PX_ID,interval_start,interval_stop,transplant,window,week_from_baseline) %>% head(20)
train_data = train_data %>%
  group_by(PX_ID,window) %>%
  mutate(albumin_at_start = first(albumin),
         bilirubin_at_start = first(bilirubin),
         eGFR_at_start = first(eGFR),
         sodium_at_start = first(sodium),
         BNP_at_start = first(BNP),
         BNP_NT_Pro_at_start = first(BNP_NT_Pro),
         durable_LVAD_at_start = first(durable_LVAD),
         short_mcs_ever_at_start = first(short_MCS_ever)) %>%
  ungroup()

# fit transplant model
## just US-CRS variables bc anything in this numerator model has to be in the mortality model
tx_model_first = glm(
   transplant ~ 
    ns(albumin_at_start, df = 3) + ns(bilirubin_at_start, df = 3) + ns(eGFR_at_start, df = 3) + ns(sodium_at_start, df = 3) + ns(BNP_at_start, df = 3)*BNP_NT_Pro_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + factor(week_from_baseline), data = train_data, family = binomial)
summary(tx_model_first)

# remove new variables from train_data
train_data = train_data %>% select(-(window:short_mcs_ever_at_start))

```





## Create week variable
```{r}
train_data = train_data %>% 
  group_by(PX_ID) %>%
  mutate(
    week = row_number()
) %>% relocate(week, .after = CAN_LISTING_CTR_CD) %>% ungroup
```


## Censor the training data on 8/15/2023
Note: This code just censors the follow-up after removal for delisted candidates.
```{r}
censor_date_train = mdy("8-15-2023")
train_data = train_data %>%
  mutate(death_after_delisting_dt = 
           as.Date(can_listing_dt) + removal_time + time_until_death_after_delisting,
         delisting_dt = ifelse(!is.na(time_until_censor_after_delisting) |
                                 !is.na(time_until_death_after_delisting),
           as.Date(can_listing_dt) + removal_time,ymd(NA)),
         ## fix these
         # code as missing if delisted after censor_date_train
         time_until_death_after_delisting =
           ifelse(is.na(delisting_dt) |
                    death_after_delisting_dt > censor_date_train |
                    delisting_dt > censor_date_train,
                  NA,time_until_death_after_delisting),
         time_until_censor_after_delisting =
           ifelse(!is.na(delisting_dt) & delisting_dt <= censor_date_train &
                    is.na(time_until_death_after_delisting) &
                    !(CAN_REM_CD %in% 13),
                  censor_date_train - delisting_dt,NA)
          )
  
```


## Adjust delisting time variables to censor at 6 weeks and create additional rows after delisting
```{r}

# Adjust delisting time variables to censor at 6 weeks
# Depending on whether the week is even or odd, this translates to 4 or 5 weeks of follow up after delisting
# If candidates dies more than 4/5 weeks after delisting, they are censored
train_data = train_data %>% group_by(PX_ID) %>% 
  
  mutate(
    # if we have full follow-up for the week, follow for censoring (floor)
    time_until_censor_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_censor_after_delisting) ~ 
                  floor(pmax(0,removal_time + time_until_censor_after_delisting - interval_stop)/7)*7,
                T ~ NA),
    # follow for death even w/o full week (ceiling)
    time_until_death_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_death_after_delisting) ~ 
                  ceiling(pmax(0,removal_time + time_until_death_after_delisting - interval_stop)/7)*7,
                T ~ NA)) %>%
  
  fill(time_until_censor_after_delisting,time_until_death_after_delisting,.direction="downup") %>%

  # censor at 6 weeks
  mutate(time_until_censor_after_delisting = case_when(
    time_until_censor_after_delisting > 28 & max(week) %% 2 == 0 ~ 28,
    time_until_censor_after_delisting > 35 & max(week) %% 2 != 0 ~ 35,
    time_until_death_after_delisting > 28 & max(week) %% 2 == 0 ~ 28,
    time_until_death_after_delisting > 35 & max(week) %% 2 != 0 ~ 35,
    TRUE ~ time_until_censor_after_delisting
  ),
    # recode time to death as missing if censored
    time_until_death_after_delisting = ifelse(!is.na(time_until_censor_after_delisting),NA,
                                              time_until_death_after_delisting),
    outcome = ifelse(!is.na(time_until_censor_after_delisting),0,outcome)
    
)

## Determine delisting row and # of duplicates we need after delisting, until death or censoring
train_data = train_data %>% group_by(PX_ID) %>% mutate(
  delisting = case_when(
    (!is.na(time_until_death_after_delisting) | !is.na(time_until_censor_after_delisting)) & row_number() == n() ~ 1,
    TRUE ~ 0)
  ) %>% ungroup() %>% mutate(
  duplicates_after_delisting = case_when(
    delisting == 1 & time_until_death_after_delisting > 0 ~
      time_until_death_after_delisting/7,
    delisting == 1 & time_until_censor_after_delisting > 0 ~ 
      time_until_censor_after_delisting/7,
  TRUE ~ 0))

## repeat rows for the duplicates after delisting
# and recode interval_start, interval_stop
train_data = train_data %>%
  slice(rep(row_number(), duplicates_after_delisting+1)) %>%
  group_by(PX_ID) %>%
  mutate(week=row_number()) %>%
  group_by(PX_ID,duplicates_after_delisting) %>%
  mutate(dup=row_number()-1) %>%
  ungroup() %>%
  mutate(interval_start=ifelse(duplicates_after_delisting>0,interval_start+7*dup,interval_start),
         interval_stop=ifelse(duplicates_after_delisting>0,interval_stop+7*dup,interval_stop)) %>%
  select(-dup)

# code outcome for last row - now outcome is correct
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    row_number()==n() & !is.na(time_until_death_after_delisting) & is.na(time_until_censor_after_delisting) ~ 1,
    T ~ 0))
```


## Adjust prob_no_tx variable after delisting
```{r}
# set prob_no_tx to 1 after delisting
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(delisting2 = cumsum(delisting),
         prob_no_tx = ifelse(delisting2>1,1,prob_no_tx))
```



# Create duplicate observations for each landmark
```{r}
# First 4 weeks don't have 3 landmarks, so they have fewer duplicates
train_data = train_data %>% 
  ungroup() %>%
  mutate(
  duplicate_amount = case_when(
    week %in% c(1,2) ~ 0,
    week %in% c(3,4) ~ 1,
    TRUE ~ 2
  ))
  
# create duplicate rows
train_data = train_data %>%
  slice(rep(row_number(), duplicate_amount+1)) %>% relocate(duplicate_amount, .after = CAN_LISTING_CTR_CD)
```

## Define landmark variable as function of PX_ID, week, and duplicate number
Also, remove transplant rows, when candidates were censored mid-week.
```{r}

combined_df = train_data %>% group_by(PX_ID, week) %>%
  mutate(rown = row_number()) %>% ungroup() %>%
  mutate(landmark = case_when(
    week %in% c(1,2) ~ 1,
    rown == 1 & week %in% c(3,4,5,6) ~ 1,
    rown == 1 & !(week %in% c(1,2,3,4,5,6)) ~ ceiling(week/2) - 2,
    rown == 2 & week %in% c(3,4,5,6) ~ 2,
    rown == 2 & !(week %in% c(3,4,5,6)) ~ ceiling(week/2) - 1,
    rown == 3 & week %in% c(5,6) ~ 3,
    rown == 3 ~ ceiling(week/2)
  )) %>% arrange(PX_ID, landmark) %>% 
  relocate(landmark, .after = CAN_LISTING_CTR_CD) %>%
  group_by(PX_ID,landmark) %>%
  mutate(week_from_baseline = row_number(),
         rem=first(delisting2)>1) %>% ungroup() %>% filter(!rem) %>% select(-rem)

# Create weight variable based on prob_no_tx
combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  weight = 1/(cumprod(prob_no_tx))
) %>% ungroup()

# remove transplant weeks
combined_df = combined_df %>% filter(transplant == 0)

```

## Landmark start variables and calculate stabilized weights
```{r}

combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  albumin_at_start = first(albumin),
  bilirubin_at_start = first(bilirubin),
  eGFR_at_start = first(eGFR),
  sodium_at_start = first(sodium),
  BNP_at_start = first(BNP),
  BNP_NT_Pro_at_start = first(BNP_NT_Pro),
  durable_LVAD_at_start = first(durable_LVAD),
  short_mcs_ever_at_start = first(short_MCS_ever),
  status_at_start = first(status),
  cpo_at_start = first(cpo),
  api_at_start = first(api),
  papi_at_start = first(papi), 
  lvad_years_at_start = first(lvad_years),
  impella_at_start = first(impella),
  iabp_at_start = first(iabp),
  impella_ever_at_start = first(impella_ever),
  iabp_ever_at_start = first(iabp_ever),
  diagnosis_at_start = first(diagnosis)
) %>% ungroup()

# calculate numerator for stabilized weights
combined_df = combined_df %>% mutate(
  prob_no_tx_start = 1 - predict(tx_model_first, newdata = combined_df, type = "response"),
  prob_no_tx_start = ifelse(delisting2>1,1,prob_no_tx_start))

# Create stabilized weight variable based on prob_no_tx and prob_no_tx_start
combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  sweight = cumprod(prob_no_tx_start)/(cumprod(prob_no_tx))
) %>% ungroup
```

## Recode BNP_NT_Pro as character
```{r}
combined_df = combined_df %>% mutate(
  BNP_NT_Pro_at_start = as.character(BNP_NT_Pro_at_start)
)
```


## Trim extreme values for model fitting
```{r trim}
# trim LVAD time to 10 years
combined_df = combined_df %>% mutate(
    lvad_years_at_start = ifelse(lvad_years_at_start>10,10,lvad_years_at_start))

# trim other continuous variables to 1st and 99th percentiles
trimOuter = function(x) {
  q99 = quantile(x,.99,na.rm=T)
  q01 = quantile(x,.01,na.rm=T)
  
  x[x>q99] = q99
  x[x<q01] = q01
  return(x)
}
combined_df = combined_df %>%
  mutate_at(vars(albumin_at_start,bilirubin_at_start,eGFR_at_start,
              sodium_at_start,cpo_at_start,api_at_start,papi_at_start),
            trimOuter) %>%
  group_by(BNP_NT_Pro_at_start) %>%
  mutate_at(vars(BNP_at_start),trimOuter) %>% ungroup
```

## Model selection
With the stabilized weights, starting from the US-CRS 1.0 base model, fit the following models:

   - Model A: US-CRS 1.0
   
   - Model B: US-CRS 1.0 + forward selection of nonlinear effects
   
   - Model C: US-CRS 1.0 + current Impella + current IABP
   
   - Model D: US-CRS 1.0 + ever Impella + ever IABP
   
   - Model E: US-CRS 1.0 with nonlinear effects (Model B) + forward selection of new main effects (Impella, IABP, time on LVAD, API, PAPi, CPO, diagnosis, CHD only)
   
   - Model F: Model E + forward selection of nonlinear effects for new variables
   
   - Model G: Model F + backward selection of 2-way interactions (between devices and labs, devices and selected hemodynamics)
   
   - Model H: Relaxed lasso
```{r}
### MODEL C: US-CRS 1.0 + current impella + current iabp ###
mortality_modelC = glm(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + iabp_at_start + impella_at_start, data = combined_df, weights = sweight, family = binomial)
AIC(mortality_modelC) # 72891.66
length(coef(mortality_modelC)) # 12

### Model F ###

# refit mortality_modelF with linear splines
combined_df = combined_df %>%
  mutate(api_at_start1.45 = ifelse(api_at_start > 1.45, api_at_start - 1.45, 0),
         api_at_start6 = ifelse(api_at_start > 6, api_at_start - 6, 0),
         cpo_at_start.8 = ifelse(cpo_at_start > .8, cpo_at_start - .8, 0),
         lvad_years_at_start3 = ifelse(lvad_years_at_start > 3, lvad_years_at_start - 3, 0),
         lvad_years_at_start5 = ifelse(lvad_years_at_start > 5, lvad_years_at_start - 5, 0),
          eGFR_at_start30 = ifelse(eGFR_at_start > 30, eGFR_at_start - 30, 0),

         )

mortality_modelF_linear = glm(
  outcome ~ 
    factor(week_from_baseline) + 
    durable_LVAD_at_start +
    lvad_years_at_start + lvad_years_at_start3 + lvad_years_at_start5 +
    iabp_at_start +
    impella_at_start + 
    short_mcs_ever_at_start +
    api_at_start + api_at_start1.45 + api_at_start6 +
    cpo_at_start + cpo_at_start.8 +
    diagnosis_at_start +
    albumin_at_start + albumin_at_start3.5 + albumin_at_start4.25 +
    eGFR_at_start + eGFR_at_start40 + eGFR_at_start90 +
    log_bili_at_start + log_bili_at_start.47 +
    log_BNP_at_start + log_BNP_at_start5.5 + log_BNP_at_start7 +
    BNP_NT_Pro_at_start +
    log_BNP_NT_Pro_at_start + log_BNP_NT_Pro_at_start5 + #log_BNP_NT_Pro_at_start5.5 +
    sodium_at_start + sodium_at_start137,
  data = combined_df,
  weights = sweight, family = binomial)
mortality_modelF_linear

```

