---
title: "mortality_model_dataset"
output: html_document
date: "2024-06-24"
---

## Load data and libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
load("discrete_final_data.RData") # df_final - created in discrete_data_prep.Rmd
load("df_intervals_setup.RData") # df_intervals - US-CRS 1.0
library(dplyr)
library(splines)
library(tidyr)
library(ggplot2)
library(rms)
library(survey) # for svydesign and svyCreateTableOne
library(tableone) # for CreateTableOne

### fix death to be pretransplant only and just one row ###

# fix extra rows w/ outcome=1
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    T ~ 0))
table(df_final$outcome)

# recode outcome as 0 if patient received transplant
## want outcome to be an indicator of pretx death
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(any_tx=any(transplant==1)) %>%
  ungroup() %>%
  mutate(outcome = ifelse(any_tx,0,outcome))
table(df_final$outcome)

# date for end of interval
# df_final = df_final %>% mutate(
#   current_date_end = current_date + pmin(last_active_date - current_date, 7)
# )

# time on LVAD at current_date
df_final = df_final %>% mutate(
  lvad_years = durable_lvad_time/365.25
)
```

## Summary of listing dates, current dates
The `df_final` dataset contains listings through 11/15/2023 and updates through 12/30/2023. The cutoff listing date was chosen to ensure all candidates had at least 6 weeks of follow-up.
```{r}
summary(df_final$can_listing_dt)
summary(df_final$current_date)
```

# Fit transplant models for IPW
The denominator of the inverse probability weights is a function of the time-varying probability of transplant. We estimated this probability by a discrete-time survival model (ie, logistic regression) for time to transplant, using the entire discrete-time dataset from listing to removal and allowing patient characteristics to update at the beginning of each week. (Note: The training/test split is designed to hold out mortality information in particular, which was not used in this step.)

We considered four specifications of the model for the denominator of the weights:

   - US-CRS 1.0 variables
   - US-CRS 1.0 variables + status
   - US-CRS 1.0 variables + status + hemodynamics
   - US-CRS 1.0 variables + status + hemodynamics + time since listing
   
The denominator weights should adjust for all time-varying confounders of transplant and mortality, and balance the distribution of confounders at each timepoint. Status was considered a potential confounder because it is a strong predictor of transplant under the current heart allocation system and may provide additional information about mortality.

We then assessed covariate balance after weighting each week of the discrete-time dataset by the inverse probability of receiving the observed treatment (transplant or no transplant) that week. The weighted distributions of all variables (US-CRS 1.0 variables + status + hemodynamics + time since listing) were compared between the groups "transplant" and "no transplant", where patients could contribute to both groups, joining the "transplant" group on their week of receiving transplant.

The model with US-CRS 1.0 variables + status and the model with US-CRS 1.0 variables + status + hemodynamics had comparable covariate balance, and the model without hemodynamics was selected for slightly better balance of key variables and parsimony.

## Compare denominator models
```{r}
## fit candidate models for denominator of IPW ##

## Model with only variables from US-CRS 1.0
tx_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro +
    durable_LVAD + short_MCS_ever), data = df_final, family = binomial)
tx_model
summary(tx_model)

## Model with US-CRS 1.0 vars and status
tx_status_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status), data = df_final, family = binomial)
tx_status_model
summary(tx_status_model)

## Model with US-CRS 1.0 vars, status, and hemodynamics
tx_status_hemo_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status + ns(cpo, df = 3) + ns(api, df = 3) + ns(papi, df = 3)), data = df_final, family = binomial)
tx_status_hemo_model
summary(tx_status_hemo_model)

## Model with US-CRS 1.0 vars, status, hemodynamics, and time since listing
tx_status_hemo_waitlist_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status + ns(cpo, df = 3) + ns(api, df = 3) + ns(papi, df = 3) + ns(interval_start, df = 3)), data = df_final, family = binomial)
tx_status_hemo_waitlist_model
summary(tx_status_hemo_waitlist_model)

## calculate weights for covariate balance assessment ##

## Predict probability of transplant
df_final$fit_prob_simple <- predict(tx_model, newdata=df_final, type = "response")
df_final$fit_prob_status <- predict(tx_status_model, newdata=df_final, type = "response")
df_final$fit_prob_hemo <- predict(tx_status_hemo_model, newdata=df_final, type = "response")
df_final$fit_prob_waitlist <- predict(tx_status_hemo_waitlist_model, newdata=df_final, type = "response")

## Construct weight variables for covariate balance assessment
df_final <- df_final %>%
  mutate(weight_simple = if_else(transplant == 1, 1 / fit_prob_simple, 1 / (1 - fit_prob_simple)),
         weight_status = if_else(transplant == 1, 1 / fit_prob_status, 1 / (1 - fit_prob_status)),
         weight_hemo = if_else(transplant == 1, 1 / fit_prob_hemo, 1 / (1 - fit_prob_hemo)),
         weight_waitlist = if_else(transplant == 1, 1 / fit_prob_waitlist, 1 / 1 - fit_prob_waitlist))

## assess covariate balance in weighted Table 1s ##

## Define variables for SMD tables
vars <- c("albumin", "bilirubin", "eGFR", "sodium", "BNP", "BNP_NT_Pro", "durable_LVAD", "short_MCS_ever", "status", "cpo", "api", "papi", "interval_start")

gc()

## remove people missing status
table_data = df_final %>% filter(!is.na(status)) %>%
  # randomly grab 100k rows or it takes too long
  ## WORKING: we should run with the full dataset (or at least full training set) for the paper
  slice_sample(n=100000)

# make unweighted table
tabUnweighted <- CreateTableOne(vars = vars, strata = "transplant", data = table_data, test = FALSE)
print(tabUnweighted, smd = TRUE) # exact = "stage", quote = TRUE, noSpaces = TRUE

## make weight_simple table
df_weights_simple <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_simple)
tabWeighted_simple <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_simple, test = FALSE)
print(tabWeighted_simple, smd = TRUE)

## make weight_status table
df_weights_status <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_status)
tabWeighted_status <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_status, test = FALSE)
print(tabWeighted_status, smd = TRUE)

## make weight_hemo table
df_weights_hemo <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_hemo)
tabWeighted_hemo <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_hemo, test = FALSE)
print(tabWeighted_hemo, smd = TRUE)

## make weight_waitlist table
df_weights_waitlist <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_waitlist)
tabWeighted_waitlist <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_waitlist, test = FALSE)
print(tabWeighted_waitlist, smd = TRUE)

# compare status and hemo models
AIC(tx_status_model) # 114330
AIC(tx_status_hemo_model) # 113961 - hemo does predict transplant

# with chosen model, set probability of not receiving a transplant each week
df_final = df_final %>%
  mutate(prob_no_tx = 1 - fit_prob_status)

# clean up
rm(df_weights_hemo,df_weights_simple,df_weights_status,df_weights_waitlist,
   table_data,tx_model,tx_status_hemo_model,tx_status_hemo_waitlist_model)
gc()
```

## Fit numerator model for stabilized weights
For stabilized weights, we also need to model the time-varying probability of transplant based on variables at baseline, which is, for us, at the start of a 6-week window. Any variables in the numerator model have to be included in the marginal structural model, so we restricted to the variables in US-CRS 1.0 (the base model) for this step. Again, we estimated this probability by a discrete-time survival model (ie, logistic regression) for time to transplant, using the entire discrete-time dataset from listing to removal, but divided into non-overlapping 6-week windows, with predictors collected at the start of the window.

Note that rows after removal/transplant are not needed for modeling the probability of transplant, because after removal, the probability of transplant is 0.
```{r}

## create numerator of stabilized weights ##
# we want Pr(transplant) given variables at start of 6-week window
# for simplicity (and to avoid correlated data), we will model this in non-overlapping 6-week windows of df_final

# create 6-week windows of df_final
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(window=ceiling(row_number()/6),
         week_from_baseline = (row_number() - 6*(window-1))) %>%
  ungroup()
# df_final %>% select(PX_ID,interval_start,interval_stop,transplant,window,week_from_baseline) %>% head(20)
df_final = df_final %>%
  group_by(PX_ID,window) %>%
  mutate(albumin_at_start = first(albumin),
         bilirubin_at_start = first(bilirubin),
         eGFR_at_start = first(eGFR),
         sodium_at_start = first(sodium),
         BNP_at_start = first(BNP),
         BNP_NT_Pro_at_start = first(BNP_NT_Pro),
         durable_LVAD_at_start = first(durable_LVAD),
         short_mcs_ever_at_start = first(short_MCS_ever)) %>%
  ungroup()

# fit transplant model
## just US-CRS variables bc anything in this numerator model has to be in the mortality model
tx_model_first = glm(
   transplant ~ 
    ns(albumin_at_start, df = 3) + ns(bilirubin_at_start, df = 3) + ns(eGFR_at_start, df = 3) + ns(sodium_at_start, df = 3) + ns(BNP_at_start, df = 3)*BNP_NT_Pro_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + factor(week_from_baseline), data = df_final, family = binomial)
summary(tx_model_first)

# remove new variables from df_final
df_final = df_final %>% select(-(window:short_mcs_ever_at_start))

```


## Split data into training / testing sets
All 1-week intervals that began between 1/1/2019 and 12/31/2022, for candidates listed by 11/15/2022, were retained in the model training data, while all other 1-week intervals were held out in a test set for model assessment.
```{r}
# confirm listing dates and current dates
with(df_final,tapply(can_listing_dt,train_test_split,summary))
with(df_final,tapply(current_date,train_test_split,summary))

#temporally split training and testing sets
train_data = df_final %>% filter(train_test_split == "train")
test_data = df_final %>% filter(train_test_split == "test")

save(test_data, file = "mortality_model_test_data.RData")

```


## Create week variable
```{r}
train_data = train_data %>% 
  group_by(PX_ID) %>%
  mutate(
    week = row_number()
) %>% relocate(week, .after = CAN_LISTING_CTR_CD) %>% ungroup
```


## Censor the training data on 12/31/2022
Note: This code just censors the follow-up for delisted candidates. There may be candidates whose final 1-week interval overlaps 1/1/2023 and includes a death at the start of 2023, if the interval started by 12/31/2022. That is fine as long as we don't include those deaths in the test data.
```{r}
library(lubridate)
censor_date_train = mdy("12-31-2022")
train_data = train_data %>%
  mutate(death_after_delisting_dt = 
           as.Date(can_listing_dt) + removal_time + time_until_death_after_delisting,
         delisting_dt = ifelse(!is.na(time_until_censor_after_delisting) |
                                 !is.na(time_until_death_after_delisting),
           as.Date(can_listing_dt) + removal_time,ymd(NA)),
         ## fix these
         # code as missing if delisted after censor_date_train
         time_until_death_after_delisting =
           ifelse(is.na(delisting_dt) |
                    death_after_delisting_dt > censor_date_train |
                    delisting_dt > censor_date_train,
                  NA,time_until_death_after_delisting),
         time_until_censor_after_delisting =
           ifelse(!is.na(delisting_dt) & delisting_dt <= censor_date_train &
                    is.na(time_until_death_after_delisting) &
                    !(CAN_REM_CD %in% 13),
                  censor_date_train - delisting_dt,NA)
          )
  
```


## Adjust delisting time variables to censor at 6 weeks and create additional rows after delisting
```{r}

# Adjust delisting time variables to censor at 6 weeks
# Depending on whether the week is even or odd, this translates to 4 or 5 weeks of follow up after delisting
# If candidates dies more than 4/5 weeks after delisting, they are censored
train_data = train_data %>% group_by(PX_ID) %>% 
  
  mutate(
    # if we have full follow-up for the week, follow for censoring (floor)
    time_until_censor_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_censor_after_delisting) ~ 
                  floor(pmax(0,removal_time + time_until_censor_after_delisting - interval_stop)/7)*7,
                T ~ NA),
    # follow for death even w/o full week (ceiling)
    time_until_death_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_death_after_delisting) ~ 
                  ceiling(pmax(0,removal_time + time_until_death_after_delisting - interval_stop)/7)*7,
                T ~ NA)) %>%
  
  fill(time_until_censor_after_delisting,time_until_death_after_delisting,.direction="downup") %>%

  # censor at 6 weeks
  mutate(time_until_censor_after_delisting = case_when(
    time_until_censor_after_delisting > 28 & max(week) %% 2 == 0 ~ 28,
    time_until_censor_after_delisting > 35 & max(week) %% 2 != 0 ~ 35,
    time_until_death_after_delisting > 28 & max(week) %% 2 == 0 ~ 28,
    time_until_death_after_delisting > 35 & max(week) %% 2 != 0 ~ 35,
    TRUE ~ time_until_censor_after_delisting
  ),
    # recode time to death as missing if censored
    time_until_death_after_delisting = ifelse(!is.na(time_until_censor_after_delisting),NA,
                                              time_until_death_after_delisting),
    outcome = ifelse(!is.na(time_until_censor_after_delisting),0,outcome)
    
)

## Determine delisting row and # of duplicates we need after delisting, until death or censoring
train_data = train_data %>% group_by(PX_ID) %>% mutate(
  delisting = case_when(
    (!is.na(time_until_death_after_delisting) | !is.na(time_until_censor_after_delisting)) & row_number() == n() ~ 1,
    TRUE ~ 0)
  ) %>% ungroup() %>% mutate(
  duplicates_after_delisting = case_when(
    delisting == 1 & time_until_death_after_delisting > 0 ~
      time_until_death_after_delisting/7,
    delisting == 1 & time_until_censor_after_delisting > 0 ~ 
      time_until_censor_after_delisting/7,
  TRUE ~ 0))

## repeat rows for the duplicates after delisting
# and recode interval_start, interval_stop
train_data = train_data %>%
  slice(rep(row_number(), duplicates_after_delisting+1)) %>%
  group_by(PX_ID) %>%
  mutate(week=row_number()) %>%
  group_by(PX_ID,duplicates_after_delisting) %>%
  mutate(dup=row_number()-1) %>%
  ungroup() %>%
  mutate(interval_start=ifelse(duplicates_after_delisting>0,interval_start+7*dup,interval_start),
         interval_stop=ifelse(duplicates_after_delisting>0,interval_stop+7*dup,interval_stop)) %>%
  select(-dup)

# code outcome for last row - now outcome is correct
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    row_number()==n() & !is.na(time_until_death_after_delisting) & is.na(time_until_censor_after_delisting) ~ 1,
    T ~ 0))
table(train_data$outcome)

# more than one death per px_id? no
train_data %>% group_by(PX_ID) %>% filter(sum(outcome)>1)
  
# more than one transplant per px_id? no
train_data %>% group_by(PX_ID) %>% filter(sum(transplant)>1)

# transplant or death not on last row? no
train_data %>% filter(outcome==1,row_number()!=n())
train_data %>% filter(transplant==1,row_number()!=n())

# Remove total_duplicates from train_data so columns match
# train_data = train_data %>% select(-any_tx,-duplicates_after_delisting)
```


## Adjust prob_no_tx variable after delisting
```{r}
# set prob_no_tx to 1 after delisting
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(delisting2 = cumsum(delisting),
         prob_no_tx = ifelse(delisting2>1,1,prob_no_tx))
table(train_data$transplant,train_data$delisting) # quick check
```



# Create duplicate observations for each landmark
```{r}
# save current version, without duplicates for landmarks
train_data_short = train_data

# First 4 weeks don't have 3 landmarks, so they have fewer duplicates
train_data = train_data %>% 
  ungroup() %>%
  mutate(
  duplicate_amount = case_when(
    week %in% c(1,2) ~ 0,
    week %in% c(3,4) ~ 1,
    TRUE ~ 2
  ))
  
# create duplicate rows
train_data = train_data %>%
  slice(rep(row_number(), duplicate_amount+1)) %>% relocate(duplicate_amount, .after = CAN_LISTING_CTR_CD)
```

## Define landmark variable as function of PX_ID, week, and duplicate number
Also, remove transplant rows, when candidates were censored mid-week.
```{r}

combined_df = train_data %>% group_by(PX_ID, week) %>%
  mutate(rown = row_number()) %>% ungroup() %>%
  mutate(landmark = case_when(
    week %in% c(1,2) ~ 1,
    rown == 1 & week %in% c(3,4,5,6) ~ 1,
    rown == 1 & !(week %in% c(1,2,3,4,5,6)) ~ ceiling(week/2) - 2,
    rown == 2 & week %in% c(3,4,5,6) ~ 2,
    rown == 2 & !(week %in% c(3,4,5,6)) ~ ceiling(week/2) - 1,
    rown == 3 & week %in% c(5,6) ~ 3,
    rown == 3 ~ ceiling(week/2)
  )) %>% arrange(PX_ID, landmark) %>% 
  relocate(landmark, .after = CAN_LISTING_CTR_CD) %>%
  rename(week_from_baseline = rown) %>%
  # remove landmarks where first row delisting2 > 1
  group_by(PX_ID,landmark) %>%
  mutate(rem=first(delisting2)>1) %>% ungroup() %>% filter(!rem) %>% select(-rem)

# is transplant always the last week? yes
combined_df %>% group_by(PX_ID,landmark) %>% filter(transplant==1,row_number()!=n())

# Create weight variable based on prob_no_tx
combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  weight = 1/(cumprod(prob_no_tx))
) %>% ungroup()

# save current version
combined_df_tx = combined_df

# remove transplant weeks
combined_df = combined_df %>% filter(transplant == 0)

```

## Landmark start variables and calculate stabilized weights
```{r}


combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  albumin_at_start = first(albumin),
  bilirubin_at_start = first(bilirubin),
  eGFR_at_start = first(eGFR),
  sodium_at_start = first(sodium),
  BNP_at_start = first(BNP),
  BNP_NT_Pro_at_start = first(BNP_NT_Pro),
  durable_LVAD_at_start = first(durable_LVAD),
  short_mcs_ever_at_start = first(short_MCS_ever),
  status_at_start = first(status),
  cpo_at_start = first(cpo),
  api_at_start = first(api),
  papi_at_start = first(papi), 
  lvad_years_at_start = first(lvad_years)
)

# calculate numerator for stabilized weights
combined_df = combined_df %>% ungroup() %>% mutate(
  prob_no_tx_start = 1 - predict(tx_model_first, newdata = combined_df, type = "response"),
  prob_no_tx_start = ifelse(delisting2>1,1,prob_no_tx_start))
with(combined_df,table(prob_no_tx==1,prob_no_tx_start==1))

# Create stabilized weight variable based on prob_no_tx and prob_no_tx_start
combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  sweight = cumprod(prob_no_tx_start)/(cumprod(prob_no_tx))
)


```

## Overall distribution of weights
```{r}
summary(combined_df$weight)
summary(combined_df$sweight)

# some people missing weights; these are people with weird statuses
combined_df %>% filter(is.na(weight)) %>%
  ungroup %>%
  count(PX_ID,status)
```

## Weight distributions by week within landmark
```{r}
with(combined_df,tapply(weight,week_from_baseline,summary))
with(combined_df,tapply(sweight,week_from_baseline,summary))

# how many rows w/ weights above 10 for each week?
with(combined_df,tapply(weight,week_from_baseline,function(x) table(x>10)))
with(combined_df,tapply(sweight,week_from_baseline,function(x) table(x>10)))
```

## Distribution of landmarks
WORKING: Might be better to show this with dates?
```{r}
# give some indication of which landmarks are contributing most strongly to estimates
## effectively averaging over this distribution

combined_df %>% ungroup() %>% filter(week_from_baseline==1) %>%
  count(interval_start/7)

ggplot(combined_df %>% ungroup() %>% filter(week_from_baseline==1),aes(interval_start/7)) +
  geom_bar() +
  coord_cartesian(xlim=c(0,500)) +
  # scale_x_continuous(breaks=seq(0,300,50)) +
  labs(x="Landmark start point (weeks from listing)",
       y="Sample size") +
  theme_bw()
```

# Function to calculate robust confidence intervals for odds ratios
```{r}
getHR <- function(model, vars, cov) {
  # Extract the coefficients for the variable of interest
  coef_var <- coef(model)[vars]
  
  # Compute the robust standard errors
  robust_se <- sqrt(diag(cov)[vars])
  
  # Calculate the odds ratio (OR) by exponentiating the coefficients
  hr <- exp(coef_var)
  
  # Calculate the 95% confidence intervals
  lowCI <- exp(coef_var - 1.96 * robust_se)
  highCI <- exp(coef_var + 1.96 * robust_se)
  
  # Return as a data frame with the OR and CIs
  return(data.frame(hr = hr, lowCI = lowCI, highCI = highCI))
}
```


## Recode BNP_NT_Pro as character
```{r}
combined_df = combined_df %>% mutate(
  BNP_NT_Pro_at_start = as.character(BNP_NT_Pro_at_start)
)
```

## Mortality model without weights
All linear effects, including hemodynamics.
```{r}
mortality_model_no_weights = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start), data = combined_df, family = binomial)
mortality_model_no_weights

# for landmark model, use robust SEs (bc of repeat observations from same px_id)
## also known as GEE with working independence model

# load package
library(sandwich)

# calculate robust VCOV matrix
mortality_model_no_weights_VCOV = vcovCL(mortality_model_no_weights, cluster = ~ PX_ID)
sqrt(diag(mortality_model_no_weights_VCOV))
summary(mortality_model_no_weights)$coef[,"Std. Error"] # smaller than robust SEs

# model summary with robust CIs
mortality_model_no_weights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_no_weights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_no_weights,vars=var_interest,cov=mortality_model_no_weights_VCOV))
}))
mortality_model_no_weights_sum

# first landmark only (from listing) - just curious
mortality_model_no_weights1 = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start +  sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start), data = subset(combined_df,landmark==1), family = binomial)
mortality_model_no_weights1
sqrt(diag(vcovCL(mortality_model_no_weights1, cluster = ~ PX_ID)))
summary(mortality_model_no_weights1)$coef[,"Std. Error"] # robust vs model-based variance are similar with only one landmark, which is as expected bc there is technically no correlated data with one landmark, just nonoverlapping survival intervals

```

## Fit US-CRS 1.0
```{r}
# lvad test variable for turning lvad variable into "currently qualifying by lvad"
df_intervals = df_intervals %>% dplyr::mutate(
  durable_LVAD = as.numeric(as.character(durable_LVAD)),
  durable_LVAD_test = case_when(
    durable_LVAD == 1 ~ 1,
    lag(PX_ID) == PX_ID & cumsum(durable_LVAD) > 0 ~ 1,
    TRUE ~ 0
  )
) 
             
df_intervals %>% select(PX_ID, durable_LVAD) %>% filter(PX_ID == 1251088)


uscrs_model = glm(formula = as.formula(
  death6week ~ 
    albumin + log(bilirubin + 1) + 
    eGFR + sodium + 
    (log(BNP + 1):BNP_NT_Pro) + 
    durable_LVAD_test + short_MCS_ever), 
  data = df_intervals[df_intervals$train_test == 'Train',], 
  family = binomial)

uscrs_model


```

## Mortality model with (unstabilized) weights
```{r}
mortality_model_weights = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start), data = combined_df, weights = weight, family = binomial)
mortality_model_weights

# summary with robust CIs
## note: the weighted model needs robust SEs whether we stack landmarks or not
mortality_model_weights_VCOV = vcovCL(mortality_model_weights, cluster = ~ PX_ID)
mortality_model_weights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_weights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_weights,vars=var_interest,cov=mortality_model_weights_VCOV))
}))
mortality_model_weights_sum
```

## Mortality model with stabilized weights
```{r}
mortality_model_sweights = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start), data = combined_df, weights = sweight, family = binomial)
mortality_model_sweights

# summary with robust CIs
mortality_model_sweights_VCOV = vcovCL(mortality_model_sweights, cluster = ~ PX_ID)
mortality_model_sweights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_sweights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_sweights,vars=var_interest,cov=mortality_model_sweights_VCOV))
}))
mortality_model_sweights_sum
```


## Compare unweighted and weighted model coefficients
```{r}
plotdat = bind_rows(mortality_model_no_weights_sum %>% mutate(model="Unweighted"),
          mortality_model_weights_sum %>% mutate(model="Unstabilized Weights"),
          mortality_model_sweights_sum %>% mutate(model="Stabilized Weights"))
ggplot(plotdat %>% filter(var!="(Intercept)"),
       aes(x = log(hr), y = var, color = model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = log(lowCI), xmax = log(highCI)), height = 0.2, position = position_dodge(width = 0.5)) +
  labs(x = "Coefficient Estimate", y = "Term", title = "Model Coefficients") +
  geom_vline(xintercept=0) +
  ## zoom in:
  # coord_cartesian(xlim=c(-1,1)) +
  theme_minimal()
```


## Truncate weights and rerun model
```{r}
# truncate stabilized weights at 1st and 99th
combined_df = combined_df %>% group_by(week_from_baseline) %>% mutate(
  tweight = ifelse(sweight > quantile(sweight,probs=.99,na.rm=T), 
                   quantile(sweight,probs=.99,na.rm=T), sweight),
  tweight = ifelse(sweight < quantile(sweight,probs=.01,na.rm=T), 
                   quantile(sweight,probs=.01,na.rm=T), tweight)
) %>% ungroup()
summary(combined_df$sweight)
summary(combined_df$tweight)
# a little low; mean weight < 1
with(combined_df,tapply(tweight,week_from_baseline,summary))

# fit model with truncated weights
mortality_model_tweights = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start), data = combined_df, weights = tweight, family = binomial)
mortality_model_tweights

# summary with robust CIs
mortality_model_tweights_VCOV = vcovCL(mortality_model_tweights, cluster = ~ PX_ID)
mortality_model_tweights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_tweights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_tweights,vars=var_interest,cov=mortality_model_tweights_VCOV))
}))
mortality_model_tweights_sum
```

# Compare unweighted, weighted, stabilized, truncated
```{r}
# compare unweighted and weighted
plotdat = bind_rows(mortality_model_no_weights_sum %>% mutate(model="Unweighted"),
          mortality_model_weights_sum %>% mutate(model="Unstabilized Weights"),
          mortality_model_sweights_sum %>% mutate(model="Stabilized Weights"),
          mortality_model_tweights_sum %>% mutate(model="Stabilized + Truncated Weights"))
ggplot(plotdat %>% filter(var!="(Intercept)"),
       aes(x = log(hr), y = var, color = model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = log(lowCI), xmax = log(highCI)), height = 0.2, position = position_dodge(width = 0.5)) +
  labs(x = "Coefficient Estimate", y = "Term", title = "Model Coefficients") +
  geom_vline(xintercept=0) +
  ## zoom in:
  # coord_cartesian(xlim=c(-1,1)) +
  theme_minimal()
# in the bias-variance tradeoff, truncated stabilized weights seem to have too much bias, almost identical to unweighted

```

## Model selection function
```{r}
#############################
# function to do forward or backward selection by AIC - more customized than MASS::stepAIC()
## output: returns the selected model formula
#############################
## fit: the model object, eg, lm, glm
## consider_vars: a named list where each element contains a possible change to the formula of fit. Should include + or - to indicate whether each variable is being added or removed from the formula.
mod_select = function(fit,consider_vars) {
  continue=T
  while (continue==T) {
    print("Current model:")
    print(formula(fit))
    print(paste("Current AIC:",AIC(fit)))
    AIC_test = lapply(consider_vars,function(var) {
      # update the model
      fit_test = stats::update(fit,as.formula(paste0(". ~ .",var)))
      return(AIC(fit_test))
    })
    if (AIC(fit) > min(as.numeric(AIC_test))) {
      select_var_name = names(consider_vars)[which.min(as.numeric(AIC_test))]
      print(paste("Selected variable for addition/removal:",select_var_name))
      select_var_full = consider_vars[select_var_name]
      fit = stats::update(fit,as.formula(paste0("~ .",select_var_full)))
      consider_vars = consider_vars[names(consider_vars) != select_var_name]
      continue=T
    } else {
      continue=F
      print("No variables selected for addition/removal")
    }
    if (length(consider_vars)==0) {print("No variables left to consider");continue=F}
  }
  print("Final fit:")
  print(formula(fit))
  print(paste("Final AIC:",AIC(fit)))
  return(formula(fit))
}
```

## Model selection
With the stabilized weights, perform forward selection, starting from the US-CRS 1.0 base model and considering whether the following reduce AIC:

   - Nonlinear effects of all continuous US-CRS 1.0 variables

   - Hemodynamics: PAPI, API, CPO
  
   - Two-way interactions of durable LVAD with sodium, albumin, bilirubin, eGFR, BNP

   - Total years on durable LVAD

WORKING: Try adding week (1-6), time since listing, previous values, interactions of current and previous values. Also, should we consider trimming to extreme quantiles (1st, 99th) to mitigate the impact of extreme (but plausible) values?
```{r}
# fit base model w/ stabilized weights
mortality_model_base = glm(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start, data = combined_df, weights = sweight, family = binomial)

# model selection by AIC
mort_mod_select = mod_select(mortality_model_base,list(
  cpo="+cpo_at_start",
  api="+api_at_start",
  papi="+papi_at_start",
  albumin_ns = "-albumin_at_start+ns(albumin_at_start,df=3)",
  bilirubin_ns = "-log(bilirubin_at_start+1)+ns(log(bilirubin_at_start+1),df=3)",
  eGFR_ns = "-eGFR_at_start+ns(eGFR_at_start,df=3)",
  BNP_ns = "-log(BNP_at_start + 1)*BNP_NT_Pro_at_start+ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start",
  sodium_ns = "-sodium_at_start+ns(sodium_at_start,df=3)",
  lvad_time = "+lvad_years_at_start",
  lvad_alb = "+durable_LVAD_at_start:albumin_at_start",
  lvad_bili = "+durable_LVAD_at_start:log(bilirubin_at_start+1)",
  lvad_eGFR = "+durable_LVAD_at_start:eGFR_at_start",
  lvad_sod = "+durable_LVAD_at_start:sodium_at_start"
  ))
## Base AIC: 66911
## Selected on Feb 18:
# lvad_time (66566), albumin_ns (66401), eGFR_ns (66251), api (66150), BNP_ns (66094), bilirubin_ns (66067), sodium_ns (66041), lvad_eGFR (66018), lvad_bili (65999), lvad_sod (65993)

# fit final model
mortality_model_selected = stats::update(mortality_model_base,mort_mod_select)
mortality_model_selected

# should api time be nonlinear?
# should LVAD time be nonlinear?
# should LVAD time be trimmed to 10 years to avoid impact of extreme values?

mortality_model_full = glm(
  outcome ~ 
    ns(albumin_at_start,df=3) + 
    ns(log(bilirubin_at_start + 1),df=3)*durable_LVAD_at_start + 
    ns(eGFR_at_start,df=3)*durable_LVAD_at_start + 
    ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start*durable_LVAD_at_start + 
    ns(sodium_at_start,df=3)*durable_LVAD_at_start + 
    durable_LVAD_at_start + 
    short_mcs_ever_at_start +
    ns(lvad_years_at_start,df=3) + #,knots=c(6,8)) +
    ns(api_at_start,df=3), 
  data = combined_df %>% mutate(
    lvad_years_at_start = ifelse(lvad_years_at_start>10,10,lvad_years_at_start)),
  weights = sweight, family = binomial)
AIC(mortality_model_full) # 64900.64 
mortality_model_full

# knots for LVAD years
attr(mortality_model_full$model$`ns(lvad_years_at_start, df = 3)`,"knots")
attr(mortality_model_full$model$`ns(lvad_years_at_start, df = 3)`,"Boundary.knots")


# get robust SEs

```


## Calculate predicted survival without transplant in training data
```{r}
# trim LVAD years and just take the first week of each landmark
pred_data = combined_df %>%
  mutate(lvad_years_at_start =
           ifelse(lvad_years_at_start>10,10,lvad_years_at_start)) %>%
  filter(week_from_baseline==1)

# predict hazard of death within the week
pred_data$haz_dth = predict(mortality_model_full,newdata=pred_data,type="response")

# convert to probability of surviving 6 weeks
pred_data$prob_surv_6wk = (1-pred_data$haz_dth)^6
summary(pred_data$prob_surv_6wk)
quantile(pred_data$prob_surv_6wk,probs=seq(0,.25,.05))
```

## Calculate 6-week survival by time on LVAD
```{r}
# Map the 6-week probability to the status system
mortality_model_status = glm(
  outcome ~ status_at_start,
  data = combined_df,
  weights = weight, # in the absence of transplant
  family = binomial)
haz_dth_status = predict(mortality_model_status,newdata=data.frame(status_at_start=factor(1:6,levels=levels(combined_df$status_at_start))),type="response")
prob_surv_6wk_status = (1-haz_dth_status)^6
prob_surv_6wk_status

# how many LVAD years do you need to move from status 4 to status 3?

# set values for predicting
new_data_lvad <- data.frame(
  albumin_at_start = mean(pred_data$albumin_at_start[pred_data$durable_LVAD_at_start==1]),
  bilirubin_at_start = mean(pred_data$bilirubin_at_start[pred_data$durable_LVAD_at_start==1]),
  eGFR_at_start = mean(pred_data$eGFR_at_start[pred_data$durable_LVAD_at_start==1]),
  BNP_at_start = mean(pred_data$BNP_at_start[pred_data$durable_LVAD_at_start==1 & 
                                               pred_data$BNP_NT_Pro_at_start=="0"]),
  BNP_NT_Pro_at_start = "0",
  sodium_at_start = mean(pred_data$sodium_at_start[pred_data$durable_LVAD_at_start==1]),
  durable_LVAD_at_start = 1,
  short_mcs_ever_at_start = 0,
  cpo_at_start = mean(pred_data$cpo_at_start[pred_data$durable_LVAD_at_start==1]),
  api_at_start = mean(pred_data$api_at_start[pred_data$durable_LVAD_at_start==1]),
  papi_at_start = mean(pred_data$papi_at_start[pred_data$durable_LVAD_at_start==1]),
  lvad_years_at_start = seq(0,10,.2)
)
head(new_data_lvad)
pred_lvad <- predict(mortality_model_full, 
                newdata = new_data_lvad,
                type = "response")
new_data_lvad$pred_surv_6_weeks = (1-pred_lvad)^6

# make plot
## add distribution of years on LVAD
## add lines for Status 3 and 4
ggplot(new_data_lvad, aes(x = lvad_years_at_start, y = pred_surv_6_weeks)) +
  geom_line(size = 1) +
  labs(
    title = "Effect of LVAD years on 6-week survival probability in durable LVAD patients",
    x = "Years with LVAD",
    y = "6-week survival without transplant"
  ) +
  theme_minimal() + 
  coord_cartesian(xlim=c(0,10),
                  ylim=c(.97,1)) +
  scale_x_continuous(breaks=seq(0,10,1)) +
  geom_hline(yintercept=prob_surv_6wk_status["3"]) +
  geom_text(aes(0,prob_surv_6wk_status["3"],label="Status 3",vjust=-1)) +
  geom_hline(yintercept=prob_surv_6wk_status["4"]) +
  geom_text(aes(0,prob_surv_6wk_status["4"],label="Status 4",vjust=-1))
prob_surv_6wk_status

```



## Forest plot
```{r,eval=F}
# Extract coefficients and confidence intervals
coef1 <- summary(mortality_model_no_weights)$coefficients
coef2 <- summary(uscrs_model)$coefficients


# Create data frames with the coefficients and confidence intervals
df1 <- data.frame(
  term = rownames(coef1),
  estimate = coef1[, "Estimate"],
  conf.low = coef1[, "Estimate"] - 1.96 * coef1[, "Std. Error"],
  conf.high = coef1[, "Estimate"] + 1.96 * coef1[, "Std. Error"],
  model = "Model without Weights"
)

# Same terms as US-CRS + hemodynamics
df2 <- data.frame(
  term = setdiff(rownames(coef1), c("papi_at_start", "cpo_at_start", "api_at_start")),
  estimate = coef2[, "Estimate"],
  conf.low = coef2[, "Estimate"] - 1.96 * coef2[, "Std. Error"],
  conf.high = coef2[, "Estimate"] + 1.96 * coef2[, "Std. Error"],
  model = "USCRS 1.0"
)

df1 <- df1[df1$term != "(Intercept)", ]
df2 <- df2[df2$term != "(Intercept)", ]

# Combine the two data frames
forest_df <- rbind(df1, df2)

coef2
df2
forest_df
class(df_intervals$BNP_NT_Pro)
setdiff(rownames(coef1), c("papi_at_start", "cpo_at_start", "api_at_start"))

# Rename and reorder variables
new_names <- c("sodium_at_start" = "Sodium", "short_mcs_ever_at_start" = "Short-term MCS", "papi_at_start" = "PAPI", "log(BNP_at_start + 1):BNP_NT_Pro_at_start0" = "Log BNP, Regular BNP", "log(BNP_at_start + 1):BNP_NT_Pro_at_start1" = "Log BNP, NT-proBNP", "log(bilirubin_at_start + 1)" = "Log Bilirubin", "eGFR_at_start" = "eGFR", "durable_LVAD_at_start" = "Durable LVAD", "cpo_at_start" = "CPO", "api_at_start" = "API", "albumin_at_start" = "Albumin" )
term_order <- c("PAPI", "API", "CPO", "Durable LVAD", "Sodium", "Albumin", "Log BNP, NT-proBNP", "Log BNP, Regular BNP", "eGFR", "Log Bilirubin", "Short-term MCS")
forest_df

forest_df <- forest_df %>%
  mutate(term = recode(term, !!!new_names))
forest_df <- forest_df %>%
  mutate(term = factor(term, levels = term_order))# Create the forest plot
ggplot(forest_df, aes(x = estimate, y = term, color = model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, position = position_dodge(width = 0.5)) +
  labs(x = "Coefficient Estimate", y = "Term", title = "Model Coefficients") +
  theme_minimal()

new_names
```


```{r}
mortality_model_sweights_lvad_time = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1):BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start + rcs(lvad_years_at_start, 5)), data = combined_df, weights = sweight, family = binomial)


mortality_model_sweights

# summary with robust CIs
mortality_model_sweights_VCOV = vcovCL(mortality_model_sweights_lvad_time, cluster = ~ PX_ID)
mortality_model_sweights_lvad_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_sweights_lvad_time)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_sweights_lvad_time,vars=var_interest,cov=mortality_model_sweights_VCOV))
}))

# compare lvad time vs no
plotdat = bind_rows(mortality_model_sweights_lvad_sum %>% mutate(model="Stabilized Weights + lvad time"),
          mortality_model_sweights_sum %>% mutate(model="Stabilized Weights"))

ggplot(plotdat %>% filter(var!="(Intercept)"),
       aes(x = log(hr), y = var, color = model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = log(lowCI), xmax = log(highCI)), height = 0.2, position = position_dodge(width = 0.5)) +
  labs(x = "Coefficient Estimate", y = "Term", title = "Model Coefficients") +
  geom_vline(xintercept=0)  +
  xlim(c(-5,5))
ggsave("stabilized_weights_lvad_time.png", height = 7, width = 16)
summary(combined_df$durable_lvad_time)


```
```{r}
combined_df$BNP_at_start
table(combined_df$BNP_NT_Pro_at_start)

new_data <- data.frame(
  albumin_at_start = mean(combined_df$albumin_at_start, na.rm = TRUE),
  bilirubin_at_start = mean(combined_df$bilirubin_at_start, na.rm = TRUE),
  eGFR_at_start = mean(combined_df$eGFR_at_start, na.rm = TRUE),
  BNP_at_start = mean(combined_df$BNP_at_start, na.rm = TRUE),
  BNP_NT_Pro_at_start = "0",
  sodium_at_start = mean(combined_df$sodium_at_start, na.rm = TRUE),
  durable_LVAD_at_start = 1, # Only generate data for LVAD patients
  short_mcs_ever_at_start = 0, # Example binary predictor
  cpo_at_start = mean(combined_df$cpo_at_start, na.rm = TRUE),
  api_at_start = mean(combined_df$api_at_start, na.rm = TRUE),
  papi_at_start = mean(combined_df$papi_at_start, na.rm = TRUE),
  lvad_years_at_start = seq(min(combined_df$lvad_years_at_start[combined_df$durable_LVAD_at_start == 1], na.rm = TRUE),
                            max(combined_df$lvad_years_at_start, na.rm = TRUE), length.out = 100)
)

pred <- predict(mortality_model_sweights_lvad_time, 
                newdata = new_data,
                type = "response")

new_data$pred_surv_6_weeks = (1-pred)^6
new_data$predicted <- pred

ggplot(new_data, aes(x = lvad_years_at_start, y = pred_surv_6_weeks)) +
  geom_line(size = 1) +
  labs(
    title = "Effect of LVAD Years on Survival Probability (Durable LVAD Patients)",
    x = "Years with LVAD",
    y = "Predicted Probability of 6-week Survival"
  ) +
  theme_minimal() + 
  xlim(c(0,5.5)) + 
  ylim(c(0.985,1))

combined_df$sweights_predictions = predict(mortality_model_sweights_lvad_time, newdata = combined_df, type = "response")
combined_df$surv_prob_6week_sweight = (1-combined_df$sweights_predictions)^6
new_data %>% select(lvad_years_at_start, predicted)

summary(combined_df$surv_prob_6week_sweight)

# survival probability histogram
ggplot(combined_df, aes(x = surv_prob_6week_sweight)) +
  geom_histogram(binwidth = 0.0001, fill = "blue", color = "black", alpha = 0.7) +
  scale_x_continuous(limits = c(0.99, 1)) + # Zoom in to the range [0.99, 1]
  labs(
    title = "Histogram of surv_prob_6week_sweight",
    x = "Survival Probability at 6 Weeks (Stabilized Weight)",
    y = "Count"
  ) +
  theme_minimal()

mortality_model_sweights_lvad_time

table(combined_df$outcome)
```


