---
title: "mortality_model_dataset"
output: html_document
date: "2024-06-24"
editor_options: 
  chunk_output_type: console
---

## Load data and libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# rm(list=ls())
load("discrete_final_data.RData") # df_final - created in discrete_data_prep.Rmd
# load("df_intervals_setup.RData") # df_intervals - US-CRS 1.0
library(dplyr)
library(splines)
library(tidyr)
library(ggplot2)
library(rms)
library(survey) # for svydesign and svyCreateTableOne
library(tableone) # for CreateTableOne
library(haven)
library(lubridate)
library(sandwich)

### fix death to be pretransplant only and just one row ###

# fix extra rows w/ outcome=1
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    T ~ 0))
table(df_final$outcome)

# recode outcome as 0 if patient received transplant
## want outcome to be an indicator of pretx death
df_final = df_final %>%
  group_by(PX_ID) %>%
  mutate(any_tx=any(transplant==1)) %>%
  ungroup() %>%
  mutate(outcome = ifelse(any_tx,0,outcome))
table(df_final$outcome)

# date for end of interval
# df_final = df_final %>% mutate(
#   current_date_end = current_date + pmin(last_active_date - current_date, 7)
# )

# fix API when negative (recode as missing, then impute w/ median)
summary(df_final$api)
df_final = df_final %>%
  mutate(api = ifelse(api<0,NA,api),
         api = ifelse(is.na(api),median(api,na.rm=T),api))
summary(df_final$api)

# time on LVAD at current_date
df_final = df_final %>% mutate(
  durable_LVAD = case_when(
    durable_lvad_time > 0 ~ 1,
    T ~ durable_LVAD
  ),
  durable_lvad_time = case_when(
    durable_LVAD == 1 & durable_lvad_implant_dt > current_date ~ as.numeric(current_date - as.Date(can_listing_dt)),
    TRUE ~ durable_lvad_time
  ),
  lvad_years = durable_lvad_time/365.25
)
```

## Summary of listing dates, current dates
The `df_final` dataset contains listings through 8/31/2025 and updates through 8/31/2025. All 1-week intervals that began between 1/1/2019 and 6/30/2023, for candidates listed by 5/15/2023, were retained in the model training data. The cutoff listing date was chosen to ensure all candidates had at least 6 weeks of follow-up.
```{r}
summary(df_final$can_listing_dt)
summary(df_final$current_date)

summary(df_final$can_listing_dt[df_final$train_test_split=="train"])
summary(df_final$current_date[df_final$train_test_split=="train"])
```

## Bring in justification form data
```{r excep}
## note: only works for Grace
curr_data_release = "//srtr-fs-p01/SAF/Internal/SAF2512/thor/ThoracicRegistration"
just_form_hr = read_sas(here::here(curr_data_release, "JustFormHR.sas7bdat"), NULL) %>%  
    zap_formats() %>% zap_labels()
just_form_hr_data_link = 
  read_sas(here::here(curr_data_release, "JustFormHRDataLink.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
risk_strat_data_hr = 
  read_sas(here::here(curr_data_release, "RiskStratDataHR.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
status_just_episode = 
  read_sas(here::here(curr_data_release, "StatusJustEpisode.sas7bdat"), NULL) %>%
  zap_formats() %>% zap_labels()
just_form_hr_stat1 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat1.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
just_form_hr_stat2 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat2.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
just_form_hr_stat3 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat3.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
just_form_hr_stat4 = 
  read_sas(here::here(curr_data_release, "JustFormHRStat4.sas7bdat"), NULL) %>%  
  zap_formats() %>% zap_labels()
thor_support_device = 
    read_sas(here::here(curr_data_release, "ThorSupportDevice.sas7bdat"), NULL) %>% 
    zap_formats() %>% zap_labels()

# join everything
justdat = inner_join(risk_strat_data_hr %>% select(px_id,wl_id,WlregAuditId),
                     just_form_hr_data_link %>% select(WlregAuditId,JustId),
                     by="WlregAuditId") %>%
  inner_join(.,
             status_just_episode %>% select(JustId,BeginDate,EndDate),
             by="JustId") %>%
  inner_join(.,
             just_form_hr %>% select(JustId,RequestedCandStatCd,Exception,ThorSupportDevicePrimary,ThorSupportDeviceSecondary),
             by="JustId") %>%
  left_join(.,
             bind_rows(just_form_hr_stat2 %>% select(JustId,CriteriaIabpSupport),
                       just_form_hr_stat3 %>% select(JustId,CriteriaIabpSupport)),
             by="JustId") %>%
  left_join(.,thor_support_device %>% select(DeviceId,DeviceTypeId,VadBrandId,TahBrandId,PercuBrandId,VentricularSupportId,ImplantDt),
            by=c("ThorSupportDevicePrimary"="DeviceId")) %>%
  mutate(BrandId = coalesce(VadBrandId,TahBrandId,PercuBrandId)) %>%
  # EndDate is missing for people who were still at that status when data were pulled
  mutate(EndDate = case_when(is.na(EndDate) ~ max(EndDate,na.rm=T),T ~ EndDate))

# special dataset for iabp and impella
# 237=Impella CP, 238=Impella RP, 239=Impella 5.5, 225=Impella Recover 2.5, 226=Impella Recover 5.0
justdat_device = justdat %>%
  mutate(iabp = case_when(DeviceTypeId %in% 8 ~ 1,T ~ 0),
         impella = case_when(BrandId %in% c(237,238,239,225,226) ~ 1,T ~ 0)) %>%
  arrange(px_id,BeginDate,EndDate) %>%
  group_by(px_id) %>%
  mutate(iabp_ever = ifelse(cumsum(iabp)>0,1,0),
         impella_ever = ifelse(cumsum(impella)>0,1,0)) %>%
  mutate(rown=row_number()==n()) %>%
  ungroup() %>%
  # define a BeginDate-EndDate range until the next form (including inactive time and status 5-6)
  mutate(EndDate = case_when(!rown ~ lead(BeginDate),
                             T ~ EndDate))
  
# keep justification forms for patients in study, on current_date (end of day)
justdat_join = justdat %>% right_join(.,df_final %>% select(PX_ID,current_date),by=c("px_id"="PX_ID")) %>%
  mutate(current_date_yms=current_date+days(1)-seconds(1)) %>% 
  filter(current_date_yms>=BeginDate,current_date_yms<=EndDate) %>% distinct # note: distinct not needed

# keep devices from forms for patients in study, on current_date (end of day)
justdat_device_join = justdat_device %>% right_join(.,df_final %>% select(PX_ID,current_date),by=c("px_id"="PX_ID")) %>%
  mutate(current_date_yms=current_date+days(1)-seconds(1)) %>% 
  filter(current_date_yms>=BeginDate,current_date_yms<=EndDate) %>% distinct # note: distinct not needed
justdat_device_join %>% count(iabp,iabp_ever,impella,impella_ever)

# join justification forms to df_final
names(justdat_join)[names(justdat_join) %in% names(df_final)]
names(justdat_device_join)[names(justdat_device_join) %in% names(df_final)]
df_final2 = df_final %>% rename(JustId_old=JustId,RequestedCandStatCd_old=RequestedCandStatCd,Exception_old=Exception) %>%
  left_join(.,justdat_join,by=c("PX_ID"="px_id","current_date")) %>%
  left_join(.,justdat_device_join %>% select(px_id,current_date,iabp,impella,iabp_ever,impella_ever),
            by=c("PX_ID"="px_id","current_date"))

## sanity checks
nrow(df_final2);nrow(df_final)
df_final2 %>% count(status,RequestedCandStatCd)
df_final2 %>% count(status,Exception)
df_final2 %>% count(iabp,impella)
df_final2 %>% count(iabp,iabp_ever,impella,impella_ever)
df_final2 %>% filter(is.na(iabp),is.na(impella)) %>% count(status)

# people missing both iabp and impella are people w/ no previous Status 1-4 form
justdat %>% right_join(.,df_final2 %>% filter(is.na(iabp),is.na(impella)) %>%
                                select(PX_ID,current_date_yms),by=c("px_id"="PX_ID")) %>%
  filter(BeginDate <= current_date_yms) %>%
  nrow

# recode missing iabp and impella as 0
df_final2 = df_final2 %>%
  mutate(iabp=ifelse(is.na(iabp),0,iabp),
         iabp_ever=ifelse(is.na(iabp_ever),0,iabp_ever),
         impella=ifelse(is.na(impella),0,impella),
         impella_ever=ifelse(is.na(impella_ever),0,impella_ever))
df_final2 %>% count(iabp,iabp_ever,impella,impella_ever)
df_final2 %>% count(status,iabp,impella) %>% print(n=30) # carry forward into next status, even if inactive, 5, or 6

# save df_final2 as df_final
df_final=df_final2
rm(df_final2)
rm(just_form_hr,just_form_hr_data_link,risk_strat_data_hr,status_just_episode,just_form_hr_stat1,just_form_hr_stat2,just_form_hr_stat3,just_form_hr_stat4,thor_support_device,justdat,justdat_device,justdat_join,justdat_device_join)
gc()
```

# Checks on balloon pump and impella variables
```{r iabp}
# some checks
with(df_final,table(iabp,iabp_ever))
with(df_final,table(impella,impella_ever))
df_final %>% filter(iabp==1) %>% count(iabp,status,Exception) # mostly status 2, standard or exception
df_final %>% filter(iabp==0,iabp_ever==1) %>% count(status,Exception) # a lot of status 4 standard and some status 2 exception
  # group_by(status) %>% summarise(propE = mean(Exception))
df_final %>% filter(impella==1) %>% count(impella,status,Exception) # mostly status 2, standard or exception
df_final %>% filter(impella==0,impella_ever==1) %>% count(status,Exception) # mostly status 4 by standard criteria

# are these candidates actually being downgraded (or upgraded)?
df_final %>% filter(train_test_split == "train") %>% filter(iabp==0,iabp_ever==1) %>% count(status,Exception,durable_LVAD,short_MCS_ever) %>% arrange(desc(n)) %>% print(n=15)
df_final %>% filter(train_test_split == "train") %>% filter(impella==0,impella_ever==1) %>% count(status,Exception,durable_LVAD,short_MCS_ever) %>% arrange(desc(n)) %>% print(n=15)
# are we confident the patient has transitioned? then we can use the currently qualifying variable

# look at someone who was inactive with iabp_ever=1 but iabp=0
## it's because they switched to Status 1 w/ a different device type (9=VA ECMO)
# df_final %>% filter(PX_ID==1027106) %>%
#   select(PX_ID,current_date,iabp,iabp_ever,status,durable_LVAD,JustId) %>%
#   filter(current_date>=ymd("2019-04-20")) %>% head(20)
# justdat %>% filter(px_id==1027106,BeginDate>ymd("2019-04-17")) %>%
#   select(px_id,BeginDate,EndDate,CriteriaIabpSupport,RequestedCandStatCd,BrandId,DeviceTypeId)
```

# Calculate diagnosis group variables
```{r diag}
# diagnosis group - CHD
  ## the four big ones including CHD
  # restrictive cardiomyopathy are ineligible to receive certain devices
# definitely include retransplant
table(df_final$diagnosis,useNA="ifany")

# read in SAF to get CAN_DGN
## only works for Grace
curr_data_release = "//srtr-fs-p01/SAF/Internal/SAF2512/thor"
df_cand_thor <- haven::read_sas(file.path(curr_data_release,'cand.sas7bdat'),
                                  col_select=c(PX_ID, CAN_DGN, CAN_DGN2))
df_final = left_join(df_final,df_cand_thor,by="PX_ID")

# look at CAN_DGN for each diagnosis
df_final %>% filter(CAN_DGN %in% c(1208,1209,1100,1199)) %>% count(diagnosis,CAN_DGN)

# recode 1100 and 1199 as Re-Transplant
df_final = df_final %>%
  mutate(diagnosis = case_when(CAN_DGN %in% c(1100,1199) ~ "Re-Transplant",
                               T ~ diagnosis))
table(df_final$diagnosis,useNA="ifany")

# recode 1208 and 1209 as Dilated_CM
df_final = df_final %>%
  mutate(diagnosis = case_when(CAN_DGN %in% c(1208,1209) ~ "Dilated_CM",
                               T ~ diagnosis))
table(df_final$diagnosis,useNA="ifany")

# look at CAN_DGN for each diagnosis
df_final %>% filter(CAN_DGN %in% c(1208,1209,1100,1199)) %>% count(diagnosis,CAN_DGN)

# confirm no missing CAN_DGN being coded as Other
df_final %>% filter(is.na(CAN_DGN)) %>% nrow

# one observation per PX_ID in training data
df_final1 = df_final %>% filter(train_test_split=="train") %>% distinct(PX_ID,diagnosis,CAN_DGN)
nrow(df_final1);dim(table(df_final1$PX_ID))
table(df_final1$diagnosis,useNA="ifany")
rm(df_final1)

## note: could pool Dilated, Restricted, Amyloid, and Hypertrophic as "Cardiomyopathy" like SRTR


## Comparison: heart-pipeline versus PSR ##
# 1000-1049 but not 1007: Dilated Cardiomyopathy - just Cardiomyopathy
# 1050-1099 but not 1051: Restricted - just Cardiomyopathy
# 1051: Amyloid - just Cardiomyopathy
# 1201: Hypertrophic - just Cardiomyopathy
# 1202: Valvular
# 1007, 1200: Ischemic - Coronary Artery Disease
# ??1208,1209
#   He has them as congenital; PSR has them as cardiomyopathy
#   After talking to Nikhil: they should be dilated cardiomyopathy
# ??1100,1199
#   He has them as Other; they should be Retransplant

## heart_pipeline_final.rmd ##
# diagnosis = case_when(
#     CAN_DGN >= 1000 & CAN_DGN <= 1049 & CAN_DGN != 1007 ~ 'Dilated_CM',
#     CAN_DGN >= 1050 & CAN_DGN <= 1099 & CAN_DGN != 1051 ~ 'Restricted',
#     CAN_DGN == 1051 ~ 'Amyloid',
#     CAN_DGN >= 1101 & CAN_DGN <= 1106 ~ 'Re-Transplant',
#     CAN_DGN == 1201 ~ 'Hypertrophic',
#     CAN_DGN == 1202 ~ 'Valvular',
#     CAN_DGN == 1203 | (CAN_DGN >= 1205 & CAN_DGN <= 1209) ~ 'Congenital', 
#     CAN_DGN == 1200 | CAN_DGN == 1007 ~ 'Ischemic',
#     TRUE ~ 'Other'), 

## PSR ##
# if cdgn in (1000 1001 1002 1003 1004 1005 1006 1049 1050 1051 1052 1053 1054 1099 1201) then DGN_PRIM = 'Cardiomyopathy';
# 	else if cdgn in (1100 1101 1102 1103 1104 1105 1106 1199) then DGN_PRIM = '.';	/** Retranspant/graft failure **/
# 	else if cdgn in (1007 1200) then DGN_PRIM = 'Coronary Artery Disease';
# 	else if cdgn in (1202) then DGN_PRIM = 'Valvular Heart Disease';
# 	else if cdgn in (1203 1205 1206 1207) then DGN_PRIM = 'Congenital Heart Disease';
# 	else if cdgn in (1204 1497 1498) then DGN_PRIM = 'Other';
# substr(can_dgn,1,4) %in% c(1008,1009,1010,1208,1209) ~ "Cardiomyopathy",
# substr(can_dgn,1,4) %in% c(1100,1101,1102,1103,1104,1105,1106,1199) ~ "Retransplant",

```

## Make new Table 1 - working
```{r}
# save data by week, including all weeks w/ no landmarks or rows after removal
save(df_final,file="discrete_data_preSplit.Rdata")


```

## Split data into training / testing sets
All 1-week intervals that began between 1/1/2019 and 6/30/2023, for candidates listed by 5/15/2023, were retained in the model training data, while all other 1-week intervals were held out in validation and test sets for model assessment.
```{r}
# confirm listing dates and current dates
with(df_final,tapply(can_listing_dt,train_test_split,summary))
with(df_final,tapply(current_date,train_test_split,summary))

#temporally split training and testing sets
train_data = df_final %>% filter(train_test_split == "train")
test_data = df_final %>% filter(train_test_split == "test")

save(test_data, file = "mortality_model_test_data.RData")

```


# Fit transplant models for IPW
The denominator of the inverse probability weights is a function of the time-varying probability of transplant. We estimated this probability by a discrete-time survival model (ie, logistic regression) for time to transplant, using the training discrete-time dataset from listing to removal and allowing patient characteristics to update at the beginning of each week.

We are fitting this model in the training data, even though no mortality information are used, because transplant probabilities shift over time and we have data over a long time period, from 2019 through August 2025. Weights for model evaluation should therefore be based on transplant models fit to the validation and/or test data.

We considered four specifications of the model for the denominator of the weights:

   - US-CRS 1.0 variables
   - US-CRS 1.0 variables + status
   - US-CRS 1.0 variables + status + hemodynamics
   - US-CRS 1.0 variables + status + hemodynamics + time since listing
   
The denominator weights should adjust for all time-varying confounders of transplant and mortality, and balance the distribution of confounders at each timepoint. Status was considered a potential confounder because it is a strong predictor of transplant under the current heart allocation system and may provide additional information about mortality.

We then assessed covariate balance after weighting each week of the discrete-time dataset by the inverse probability of receiving the observed treatment (transplant or no transplant) that week. The weighted distributions of all variables (US-CRS 1.0 variables + status + hemodynamics + time since listing) were compared between the groups "transplant" and "no transplant", where patients could contribute to both groups, joining the "transplant" group on their week of receiving transplant.

The model with US-CRS 1.0 variables + status + hemodynamics had the best covariate balance.

## Compare denominator models
```{r}
## fit candidate models for denominator of IPW ##

## Model with only variables from US-CRS 1.0
tx_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro +
    durable_LVAD + short_MCS_ever), data = train_data, family = binomial)
tx_model
summary(tx_model)

## Model with US-CRS 1.0 vars and status
tx_status_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status), data = train_data, family = binomial)
tx_status_model
summary(tx_status_model)

## Model with US-CRS 1.0 vars, status, and hemodynamics
tx_status_hemo_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status + ns(cpo, df = 3) + ns(api, df = 3) + ns(papi, df = 3)), data = train_data, family = binomial)
tx_status_hemo_model
summary(tx_status_hemo_model)

## Model with US-CRS 1.0 vars, status, hemodynamics, and time since listing
# (Cox model w/ left truncation)
tx_status_hemo_waitlist_model = glm(formula = as.formula(
   transplant ~ 
    ns(albumin, df = 3) + ns(bilirubin, df = 3) + ns(eGFR, df = 3) + ns(sodium, df = 3) + 
    ns(BNP, df = 3)*BNP_NT_Pro + 
    durable_LVAD + short_MCS_ever + status + ns(cpo, df = 3) + ns(api, df = 3) + ns(papi, df = 3) + ns(interval_start, df = 3)), data = train_data, family = binomial)
tx_status_hemo_waitlist_model
summary(tx_status_hemo_waitlist_model)

## calculate weights for covariate balance assessment ##

## Predict probability of transplant
train_data$fit_prob_simple <- predict(tx_model, newdata=train_data, type = "response")
train_data$fit_prob_status <- predict(tx_status_model, newdata=train_data, type = "response")
train_data$fit_prob_hemo <- predict(tx_status_hemo_model, newdata=train_data, type = "response")
train_data$fit_prob_waitlist <- predict(tx_status_hemo_waitlist_model, newdata=train_data, type = "response")

## Construct weight variables for covariate balance assessment
train_data <- train_data %>%
  mutate(weight_simple = if_else(transplant == 1, 1 / fit_prob_simple, 1 / (1 - fit_prob_simple)),
         weight_status = if_else(transplant == 1, 1 / fit_prob_status, 1 / (1 - fit_prob_status)),
         weight_hemo = if_else(transplant == 1, 1 / fit_prob_hemo, 1 / (1 - fit_prob_hemo)),
         weight_waitlist = if_else(transplant == 1, 1 / fit_prob_waitlist, 1 / 1 - fit_prob_waitlist))

## assess covariate balance in weighted Table 1s ##

## Define variables for SMD tables
vars <- c("albumin", "bilirubin", "eGFR", "sodium", "BNP", "BNP_NT_Pro", "durable_LVAD", "short_MCS_ever", "cpo", "api", "papi", "interval_start", "status_num")

gc()

## remove people missing status
table_data = train_data %>% filter(!is.na(status)) %>%
  mutate(status_num = case_when(
    status %in% "inactive" ~ "7",
    T ~ status
  ),
  status_num = as.numeric(status_num))
with(table_data,table(status_num,status))

# make unweighted table
tabUnweighted <- CreateTableOne(vars = vars, strata = "transplant", data = table_data, test = FALSE)
print(tabUnweighted, smd = TRUE)
printUnweighted = print(tabUnweighted, smd = TRUE, exact = "stage", quote = FALSE, noSpaces = TRUE, printToggle = F)
write.csv(printUnweighted, file = "tabUnweighted.csv")

## make weight_simple table
df_weights_simple <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_simple)
tabWeighted_simple <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_simple, test = FALSE)
print(tabWeighted_simple, smd = TRUE)
printSimple = print(tabWeighted_simple, smd = TRUE, exact = "stage", quote = FALSE, noSpaces = TRUE, printToggle = F)
write.csv(printSimple, file = "tabSimple.csv")


## make weight_status table
df_weights_status <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_status)
tabWeighted_status <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_status, test = FALSE)
print(tabWeighted_status, smd = TRUE)
printStatus = print(tabWeighted_status, smd = TRUE, exact = "stage", quote = FALSE, noSpaces = TRUE, printToggle = F)
write.csv(printStatus, file = "tabStatus.csv")


## make weight_hemo table
df_weights_hemo <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_hemo)
tabWeighted_hemo <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_hemo, test = FALSE)
print(tabWeighted_hemo, smd = TRUE)
printHemo = print(tabWeighted_hemo, smd = TRUE, exact = "stage", quote = FALSE, noSpaces = TRUE, printToggle = F)
write.csv(printHemo, file = "tabHemo.csv")

## make weight_waitlist table
df_weights_waitlist <- svydesign(ids = ~ 1, data = table_data, weights = ~ weight_waitlist)
tabWeighted_waitlist <- svyCreateTableOne(vars = vars, strata = "transplant", data = df_weights_waitlist, test = FALSE)
print(tabWeighted_waitlist, smd = TRUE)
printWaitlist = print(tabWeighted_waitlist, smd = TRUE, exact = "stage", quote = FALSE, noSpaces = TRUE, printToggle = F)
write.csv(printWaitlist, file = "tabWaitlist.csv")

# compare status and hemo models
AIC(tx_status_model) # 100286.1
AIC(tx_status_hemo_model) # 99945.18 - hemo does predict transplant

# any reason not to do hemo? positivity violations? no
summary(train_data$fit_prob_status)
summary(train_data$fit_prob_hemo)

# with chosen model, set probability of not receiving a transplant each week
train_data = train_data %>%
  mutate(prob_no_tx = 1 - fit_prob_hemo)

# clean up
rm(df_weights_hemo,df_weights_simple,df_weights_status,df_weights_waitlist,
   table_data,tx_model,tx_status_hemo_model,tx_status_hemo_waitlist_model)
gc()
```

## Fit numerator model for stabilized weights
For stabilized weights, we also need to model the time-varying probability of transplant based on variables at baseline, which is, for us, at the start of a 6-week window. Any variables in the numerator model have to be included in the marginal structural model, so we restricted to the variables in US-CRS 1.0 (the base model) for this step. Again, we estimated this probability by a discrete-time survival model (ie, logistic regression) for time to transplant, using the training discrete-time dataset from listing to removal, but divided into non-overlapping 6-week windows, with predictors collected at the start of the window.

Note that rows after removal/transplant are not needed for modeling the probability of transplant, because after removal, the probability of transplant is 0.
```{r}

## create numerator of stabilized weights ##
# we want Pr(transplant) given variables at start of 6-week window
# for simplicity (and to avoid correlated data), we will model this in non-overlapping 6-week windows of train_data

# create 6-week windows of train_data
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(window=ceiling(row_number()/6),
         week_from_baseline = (row_number() - 6*(window-1))) %>%
  ungroup()
# train_data %>% select(PX_ID,interval_start,interval_stop,transplant,window,week_from_baseline) %>% head(20)
train_data = train_data %>%
  group_by(PX_ID,window) %>%
  mutate(albumin_at_start = first(albumin),
         bilirubin_at_start = first(bilirubin),
         eGFR_at_start = first(eGFR),
         sodium_at_start = first(sodium),
         BNP_at_start = first(BNP),
         BNP_NT_Pro_at_start = first(BNP_NT_Pro),
         durable_LVAD_at_start = first(durable_LVAD),
         short_mcs_ever_at_start = first(short_MCS_ever)) %>%
  ungroup()

# fit transplant model
## just US-CRS variables bc anything in this numerator model has to be in the mortality model
tx_model_first = glm(
   transplant ~ 
    ns(albumin_at_start, df = 3) + ns(bilirubin_at_start, df = 3) + ns(eGFR_at_start, df = 3) + ns(sodium_at_start, df = 3) + ns(BNP_at_start, df = 3)*BNP_NT_Pro_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + factor(week_from_baseline), data = train_data, family = binomial)
summary(tx_model_first)

# remove new variables from train_data
train_data = train_data %>% select(-(window:short_mcs_ever_at_start))

```





## Create week variable
```{r}
train_data = train_data %>% 
  group_by(PX_ID) %>%
  mutate(
    week = row_number()
) %>% relocate(week, .after = CAN_LISTING_CTR_CD) %>% ungroup
```


## Censor the training data on 8/15/2023
Note: This code just censors the follow-up after removal for delisted candidates.
```{r}
censor_date_train = mdy("8-15-2023")
train_data = train_data %>%
  mutate(death_after_delisting_dt = 
           as.Date(can_listing_dt) + removal_time + time_until_death_after_delisting,
         delisting_dt = ifelse(!is.na(time_until_censor_after_delisting) |
                                 !is.na(time_until_death_after_delisting),
           as.Date(can_listing_dt) + removal_time,ymd(NA)),
         ## fix these
         # code as missing if delisted after censor_date_train
         time_until_death_after_delisting =
           ifelse(is.na(delisting_dt) |
                    death_after_delisting_dt > censor_date_train |
                    delisting_dt > censor_date_train,
                  NA,time_until_death_after_delisting),
         time_until_censor_after_delisting =
           ifelse(!is.na(delisting_dt) & delisting_dt <= censor_date_train &
                    is.na(time_until_death_after_delisting) &
                    !(CAN_REM_CD %in% 13),
                  censor_date_train - delisting_dt,NA)
          )
  
```


## Adjust delisting time variables to censor at 6 weeks and create additional rows after delisting
```{r}

# Adjust delisting time variables to censor at 6 weeks
# Depending on whether the week is even or odd, this translates to 4 or 5 weeks of follow up after delisting
# If candidates dies more than 4/5 weeks after delisting, they are censored
train_data = train_data %>% group_by(PX_ID) %>% 
  
  mutate(
    # if we have full follow-up for the week, follow for censoring (floor)
    time_until_censor_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_censor_after_delisting) ~ 
                  floor(pmax(0,removal_time + time_until_censor_after_delisting - interval_stop)/7)*7,
                T ~ NA),
    # follow for death even w/o full week (ceiling)
    time_until_death_after_delisting = 
      case_when(row_number() == n() & !is.na(time_until_death_after_delisting) ~ 
                  ceiling(pmax(0,removal_time + time_until_death_after_delisting - interval_stop)/7)*7,
                T ~ NA)) %>%
  
  fill(time_until_censor_after_delisting,time_until_death_after_delisting,.direction="downup") %>%

  # censor at 6 weeks
  mutate(time_until_censor_after_delisting = case_when(
    time_until_censor_after_delisting > 28 & max(week) %% 2 == 0 ~ 28,
    time_until_censor_after_delisting > 35 & max(week) %% 2 != 0 ~ 35,
    time_until_death_after_delisting > 28 & max(week) %% 2 == 0 ~ 28,
    time_until_death_after_delisting > 35 & max(week) %% 2 != 0 ~ 35,
    TRUE ~ time_until_censor_after_delisting
  ),
    # recode time to death as missing if censored
    time_until_death_after_delisting = ifelse(!is.na(time_until_censor_after_delisting),NA,
                                              time_until_death_after_delisting),
    outcome = ifelse(!is.na(time_until_censor_after_delisting),0,outcome)
    
)

## Determine delisting row and # of duplicates we need after delisting, until death or censoring
train_data = train_data %>% group_by(PX_ID) %>% mutate(
  delisting = case_when(
    (!is.na(time_until_death_after_delisting) | !is.na(time_until_censor_after_delisting)) & row_number() == n() ~ 1,
    TRUE ~ 0)
  ) %>% ungroup() %>% mutate(
  duplicates_after_delisting = case_when(
    delisting == 1 & time_until_death_after_delisting > 0 ~
      time_until_death_after_delisting/7,
    delisting == 1 & time_until_censor_after_delisting > 0 ~ 
      time_until_censor_after_delisting/7,
  TRUE ~ 0))

## repeat rows for the duplicates after delisting
# and recode interval_start, interval_stop
train_data = train_data %>%
  slice(rep(row_number(), duplicates_after_delisting+1)) %>%
  group_by(PX_ID) %>%
  mutate(week=row_number()) %>%
  group_by(PX_ID,duplicates_after_delisting) %>%
  mutate(dup=row_number()-1) %>%
  ungroup() %>%
  mutate(interval_start=ifelse(duplicates_after_delisting>0,interval_start+7*dup,interval_start),
         interval_stop=ifelse(duplicates_after_delisting>0,interval_stop+7*dup,interval_stop)) %>%
  select(-dup)

# code outcome for last row - now outcome is correct
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(outcome = case_when(
    row_number()==n() & outcome==1 ~ 1,
    row_number()==n() & !is.na(time_until_death_after_delisting) & is.na(time_until_censor_after_delisting) ~ 1,
    T ~ 0))
table(train_data$outcome) # 1936 
table(train_data$outcome,train_data$CAN_REM_CD) # 1077 removals due to deteriorated condition

# more than one death per px_id? no
train_data %>% group_by(PX_ID) %>% filter(sum(outcome)>1) %>% nrow
  
# more than one transplant per px_id? no
train_data %>% group_by(PX_ID) %>% filter(sum(transplant)>1) %>% nrow

# transplant or death not on last row? no
train_data %>% filter(outcome==1,row_number()!=n()) %>% nrow
train_data %>% filter(transplant==1,row_number()!=n()) %>% nrow

# Remove total_duplicates from train_data so columns match
# train_data = train_data %>% select(-any_tx,-duplicates_after_delisting)
```


## Adjust prob_no_tx variable after delisting
```{r}
# set prob_no_tx to 1 after delisting
train_data = train_data %>%
  group_by(PX_ID) %>%
  mutate(delisting2 = cumsum(delisting),
         prob_no_tx = ifelse(delisting2>1,1,prob_no_tx))
table(train_data$transplant,train_data$delisting) # quick check
```



# Create duplicate observations for each landmark
```{r}
# save current version, without duplicates for landmarks
train_data_short = train_data

# First 4 weeks don't have 3 landmarks, so they have fewer duplicates
train_data = train_data %>% 
  ungroup() %>%
  mutate(
  duplicate_amount = case_when(
    week %in% c(1,2) ~ 0,
    week %in% c(3,4) ~ 1,
    TRUE ~ 2
  ))
  
# create duplicate rows
train_data = train_data %>%
  slice(rep(row_number(), duplicate_amount+1)) %>% relocate(duplicate_amount, .after = CAN_LISTING_CTR_CD)
```

## Define landmark variable as function of PX_ID, week, and duplicate number
Also, remove transplant rows, when candidates were censored mid-week.
```{r}

combined_df = train_data %>% group_by(PX_ID, week) %>%
  mutate(rown = row_number()) %>% ungroup() %>%
  mutate(landmark = case_when(
    week %in% c(1,2) ~ 1,
    rown == 1 & week %in% c(3,4,5,6) ~ 1,
    rown == 1 & !(week %in% c(1,2,3,4,5,6)) ~ ceiling(week/2) - 2,
    rown == 2 & week %in% c(3,4,5,6) ~ 2,
    rown == 2 & !(week %in% c(3,4,5,6)) ~ ceiling(week/2) - 1,
    rown == 3 & week %in% c(5,6) ~ 3,
    rown == 3 ~ ceiling(week/2)
  )) %>% arrange(PX_ID, landmark) %>% 
  relocate(landmark, .after = CAN_LISTING_CTR_CD) %>%
  group_by(PX_ID,landmark) %>%
  mutate(week_from_baseline = row_number(),
         rem=first(delisting2)>1) %>% ungroup() %>% filter(!rem) %>% select(-rem)

# is transplant always the last week? yes
combined_df %>% group_by(PX_ID,landmark) %>% filter(transplant==1,row_number()!=n()) %>% nrow

# Create weight variable based on prob_no_tx
combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  weight = 1/(cumprod(prob_no_tx))
) %>% ungroup()

# save current version
combined_df_tx = combined_df

# remove transplant weeks
combined_df = combined_df %>% filter(transplant == 0)

```

## Landmark start variables and calculate stabilized weights
```{r}


combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  albumin_at_start = first(albumin),
  bilirubin_at_start = first(bilirubin),
  eGFR_at_start = first(eGFR),
  sodium_at_start = first(sodium),
  BNP_at_start = first(BNP),
  BNP_NT_Pro_at_start = first(BNP_NT_Pro),
  durable_LVAD_at_start = first(durable_LVAD),
  short_mcs_ever_at_start = first(short_MCS_ever),
  status_at_start = first(status),
  cpo_at_start = first(cpo),
  api_at_start = first(api),
  papi_at_start = first(papi), 
  lvad_years_at_start = first(lvad_years),
  impella_at_start = first(impella),
  iabp_at_start = first(iabp),
  impella_ever_at_start = first(impella_ever),
  iabp_ever_at_start = first(iabp_ever),
  diagnosis_at_start = first(diagnosis)
) %>% ungroup()

# calculate numerator for stabilized weights
combined_df = combined_df %>% mutate(
  prob_no_tx_start = 1 - predict(tx_model_first, newdata = combined_df, type = "response"),
  prob_no_tx_start = ifelse(delisting2>1,1,prob_no_tx_start))
with(combined_df,table(prob_no_tx==1,prob_no_tx_start==1))

# Create stabilized weight variable based on prob_no_tx and prob_no_tx_start
combined_df = combined_df %>% group_by(PX_ID, landmark) %>% mutate(
  sweight = cumprod(prob_no_tx_start)/(cumprod(prob_no_tx))
) %>% ungroup


```

## Overall distribution of weights
```{r}
summary(combined_df$weight)
summary(combined_df$sweight)

# some people missing weights; these are people with weird statuses
combined_df %>% filter(is.na(weight)) %>%
  count(PX_ID,status)
```

## Weight distributions by week within landmark
```{r}
with(combined_df,tapply(weight,week_from_baseline,summary))
with(combined_df,tapply(sweight,week_from_baseline,summary))

# how many rows w/ weights above 10 for each week?
with(combined_df,tapply(weight,week_from_baseline,function(x) table(x>10)))
with(combined_df,tapply(sweight,week_from_baseline,function(x) table(x>10)))
```

## Distribution of landmarks
For the supplement, to show which landmarks are contributing most strongly to estimates.
```{r}
# give some indication of which landmarks are contributing most strongly to estimates
## effectively averaging over this distribution

combined_df %>% ungroup() %>% filter(week_from_baseline==1) %>%
  count(interval_start/7)

ggplot(combined_df %>% ungroup() %>% filter(week_from_baseline==1),aes(interval_start/365.25)) +
  geom_bar() +
  coord_cartesian(xlim=c(0,5)) +
  # scale_x_continuous(breaks=seq(0,300,50)) +
  labs(x="Landmark start time (years from listing)",
       y="Number of candidates") +
  theme_bw()
ggsave("landmark-distribution.png",width=7.5,height=5,dpi=300)
```

# Function to calculate robust confidence intervals for odds ratios
```{r}
#############################
# function to get HR for a linear combination of betas and normal-based CIs (helpful w/ interaction models)
#############################
## mod: the model object, eg, lm, glm, coxph
## vars: names of the coefficients to be combined in a linear combination
## comb: numeric vector (same length as vars) with multipliers for the linear combination, typically -1 or 1; default is 1 for a sum across coefficients in vars
## cov: covariance matrix for the coefficients; default is vcov(mod) but could be provided by the user (eg, an empirical covariance matrix derived from bootstrapping)
## alpha: significance level which determines two-sided confidence interval coverage; default is 0.05 for a 95% CI
getHR = function(mod,vars,comb=rep(1,length(vars)),cov=vcov(mod),transform=T,alpha=0.05) {
  # get beta from mod object
  beta = coef(mod)
  beta[is.na(beta)] = 0
  
  # get combination vector
  c = rep(0,length(beta))
  names(c) = names(beta)
  c[vars] = comb
  
  # compute the HR
  xb = c %*% beta
  hr = exp(xb)
  
  if (!all.equal(names(c),colnames(cov))) {
    stop("Mismatched names bw beta and var-cov matrix")
  }
  
  # compute the 95% CI
  se = sqrt(t(c) %*% cov %*% c)
  lowCI = xb - qnorm(1-alpha/2)*se
  lowCI_hr = exp(lowCI)
  highCI = xb + qnorm(1-alpha/2)*se
  highCI_hr = exp(highCI)
  p = pnorm(abs(xb/se),lower.tail=F)*2
  if (transform) {
    df = data.frame(hr,lowCI=lowCI_hr,highCI=highCI_hr,p)
  } else {
    df = data.frame(xb,se,lowCI,highCI,p)
  }
  return(df)
}

#############################
# function to format p-values
#############################
pval_format = function(p,include.p=F) {
  if (include.p) {
    p = ifelse(p<0.001,"p<0.001",
               paste0("p=",ifelse(p<0.005,round(p,3),format(round(p,2),nsmall=2))))
  } else {
    p = ifelse(p<0.001,"<0.001",
               ifelse(p<0.005,round(p,3),format(round(p,2),nsmall=2)))
  }
  return(p)
}

```


## Recode BNP_NT_Pro as character
```{r}
combined_df = combined_df %>% mutate(
  BNP_NT_Pro_at_start = as.character(BNP_NT_Pro_at_start)
)
```

## Mortality model without weights
All linear effects, including hemodynamics.
```{r}
# make standardized values
combined_df = combined_df %>%
  mutate(albumin_at_start_s = scale(albumin_at_start),
         bilirubin_at_start_s = scale(log(bilirubin_at_start+1)),
         eGFR_at_start_s = scale(eGFR_at_start),
         BNP_at_start_s = scale(log(BNP_at_start+1)),
         sodium_at_start_s = scale(sodium_at_start),
         cpo_at_start_s = scale(cpo_at_start),
         api_at_start_s = scale(api_at_start),
         papi_at_start_s = scale(papi_at_start))
basic_form = as.formula(outcome ~  albumin_at_start_s + bilirubin_at_start_s + eGFR_at_start_s + BNP_at_start_s*BNP_NT_Pro_at_start + sodium_at_start_s + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start_s + api_at_start_s + papi_at_start_s)         

mortality_model_no_weights = glm(basic_form, data = combined_df, family = binomial)
mortality_model_no_weights

# for landmark model, use robust SEs (bc of repeat observations from same px_id)
## also known as GEE with working independence model

# calculate robust VCOV matrix
mortality_model_no_weights_VCOV = vcovCL(mortality_model_no_weights, cluster = ~ PX_ID)
sqrt(diag(mortality_model_no_weights_VCOV))
summary(mortality_model_no_weights)$coef[,"Std. Error"] # smaller than robust SEs

# model summary with robust CIs
mortality_model_no_weights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_no_weights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_no_weights,vars=var_interest,cov=mortality_model_no_weights_VCOV))
}))
mortality_model_no_weights_sum

# first landmark only (from listing) - just curious
mortality_model_no_weights1 = glm(formula = as.formula(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start +  sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + cpo_at_start + api_at_start + papi_at_start), data = subset(combined_df,landmark==1), family = binomial)
mortality_model_no_weights1
sqrt(diag(vcovCL(mortality_model_no_weights1, cluster = ~ PX_ID)))
summary(mortality_model_no_weights1)$coef[,"Std. Error"] # robust vs model-based variance are similar with only one landmark, which is as expected bc there is technically no correlated data with one landmark, just nonoverlapping survival intervals

```

## Fit US-CRS 1.0
```{r,eval=F}
# lvad test variable for turning lvad variable into "currently qualifying by lvad"
df_intervals = df_intervals %>% dplyr::mutate(
  durable_LVAD = as.numeric(as.character(durable_LVAD)),
  durable_LVAD_test = case_when(
    durable_LVAD == 1 ~ 1,
    lag(PX_ID) == PX_ID & cumsum(durable_LVAD) > 0 ~ 1,
    TRUE ~ 0
  )
) 
             
df_intervals %>% select(PX_ID, durable_LVAD) %>% filter(PX_ID == 1251088)


uscrs_model = glm(formula = as.formula(
  death6week ~ 
    albumin + log(bilirubin + 1) + 
    eGFR + sodium + 
    (log(BNP + 1):BNP_NT_Pro) + 
    durable_LVAD_test + short_MCS_ever), 
  data = df_intervals[df_intervals$train_test == 'Train',], 
  family = binomial)

uscrs_model


```

## Mortality model with (unstabilized) weights
```{r}
mortality_model_weights = glm(formula = basic_form, data = combined_df, weights = weight, family = binomial)
mortality_model_weights

# summary with robust CIs
## note: the weighted model needs robust SEs whether we stack landmarks or not
mortality_model_weights_VCOV = vcovCL(mortality_model_weights, cluster = ~ PX_ID)
mortality_model_weights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_weights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_weights,vars=var_interest,cov=mortality_model_weights_VCOV))
}))
mortality_model_weights_sum
```

## Mortality model with stabilized weights
```{r}
mortality_model_sweights = glm(basic_form, data = combined_df, weights = sweight, family = binomial)
mortality_model_sweights

# summary with robust CIs
mortality_model_sweights_VCOV = vcovCL(mortality_model_sweights, cluster = ~ PX_ID)
mortality_model_sweights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_sweights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_sweights,vars=var_interest,cov=mortality_model_sweights_VCOV))
}))
mortality_model_sweights_sum
```


## Compare unweighted and weighted model coefficients
```{r}
# new_names <- c("sodium_at_start" = "Sodium", "short_mcs_ever_at_start" = "Short-term MCS", "papi_at_start" = "PAPi", "log(BNP_at_start + 1)" = "Log BNP", "BNP_NT_Pro_at_start1" = "BNP NT Pro", "log(BNP_at_start + 1):BNP_NT_Pro_at_start1" = "Log BNP * BNP NT Pro", "log(bilirubin_at_start + 1)" = "Log Bilirubin", "eGFR_at_start" = "eGFR", "durable_LVAD_at_start" = "Durable LVAD", "cpo_at_start" = "CPO", "api_at_start" = "API", "albumin_at_start" = "Albumin" )
# plotdat = bind_rows(mortality_model_no_weights_sum %>% mutate(model="Unweighted"),
#           mortality_model_weights_sum %>% mutate(model="Unstabilized Weights"),
#           mortality_model_sweights_sum %>% mutate(model="Stabilized Weights")) %>%
#   mutate(var = recode(var, !!!new_names))
# ggplot(plotdat %>% filter(var!="(Intercept)"),
#        aes(x = log(hr), y = var, color = model)) +
#   geom_point(position = position_dodge(width = 0.5)) +
#   geom_errorbarh(aes(xmin = log(lowCI), xmax = log(highCI)), height = 0.2, position = position_dodge(width = 0.5)) +
#   labs(x = "Coefficient Estimate", y = "Term", title = "Model Coefficients") +
#   geom_vline(xintercept=0) +
#   ## zoom in:
#   # coord_cartesian(xlim=c(-1,1)) +
#   theme_minimal()
```


## Truncate weights and rerun model
```{r}
# truncate stabilized weights at 1st and 99th
combined_df = combined_df %>% group_by(week_from_baseline) %>% mutate(
  tweight = ifelse(sweight > quantile(sweight,probs=.99,na.rm=T), 
                   quantile(sweight,probs=.99,na.rm=T), sweight),
  tweight = ifelse(sweight < quantile(sweight,probs=.01,na.rm=T), 
                   quantile(sweight,probs=.01,na.rm=T), tweight)
) %>% ungroup()
summary(combined_df$sweight)
summary(combined_df$tweight)
# a little low; mean weight < 1
with(combined_df,tapply(tweight,week_from_baseline,summary))

# make a table 1 of weight distributions by week
weighttab = combined_df %>% select(week_from_baseline,weight,sweight,tweight)
weighttab$week_from_baseline = paste("Week",weighttab$week_from_baseline)
label(weighttab$weight) = "Unstabilized"
label(weighttab$sweight) = "Stabilized"
label(weighttab$tweight) = "Stabilized + Truncated"
table1(~ weight + sweight + tweight | week_from_baseline, 
       data=weighttab %>% filter(!is.na(weight)), overall = F)

# fit model with truncated weights
mortality_model_tweights = glm(formula = basic_form, data = combined_df, weights = tweight, family = binomial)
mortality_model_tweights

# summary with robust CIs
mortality_model_tweights_VCOV = vcovCL(mortality_model_tweights, cluster = ~ PX_ID)
mortality_model_tweights_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_tweights)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_tweights,vars=var_interest,cov=mortality_model_tweights_VCOV))
}))
mortality_model_tweights_sum
```

# Compare unweighted, weighted, stabilized, truncated
```{r}
new_names <- c("sodium_at_start_s" = "Sodium", "short_mcs_ever_at_start" = "Short-term MCS", "papi_at_start_s" = "PAPi", "BNP_at_start_s" = "Log BNP", "BNP_NT_Pro_at_start1" = "BNP NT Pro", "BNP_at_start_s:BNP_NT_Pro_at_start1" = "Log BNP * BNP NT Pro", "bilirubin_at_start_s" = "Log Bilirubin", "eGFR_at_start_s" = "eGFR", "durable_LVAD_at_start" = "Durable LVAD", "cpo_at_start_s" = "CPO", "api_at_start_s" = "API", "albumin_at_start_s" = "Albumin" )
term_order <- c("Albumin","Log Bilirubin","eGFR", "Log BNP", "BNP NT Pro", "Log BNP * BNP NT Pro", "Sodium", "Durable LVAD", "Short-term MCS", "CPO", "API", "PAPi")

# compare unweighted and weighted
plotdat = bind_rows(mortality_model_no_weights_sum %>% mutate(model="Unweighted"),
          mortality_model_weights_sum %>% mutate(model="Unstabilized Weights"),
          mortality_model_sweights_sum %>% mutate(model="Stabilized Weights"),
          mortality_model_tweights_sum %>% mutate(model="Stabilized + Truncated\nWeights")) %>%
  mutate(var = recode(var, !!!new_names),
         var = factor(var,levels=term_order),
         model = factor(model,levels=c("Unweighted","Unstabilized Weights","Stabilized Weights","Stabilized + Truncated\nWeights")))

ggplot(plotdat %>% filter(var!="(Intercept)"),
       aes(x = log(hr), y = var, color = model, linetype=model)) +
  geom_point(position = position_dodge(width = 0.5),size=1) +
  geom_errorbarh(aes(xmin = log(lowCI), xmax = log(highCI)), height = 0.2, position = position_dodge(width = 0.5),linewidth=.4) +
  labs(x = "Coefficient Estimate", y = "Term", color="Model",linetype="Model") +
  geom_vline(xintercept=0) +
  scale_color_viridis_d() +
  ## zoom in:
  # coord_cartesian(xlim=c(-.25,.25)) +
  theme_bw() +
  theme(text=element_text(size=12)) +
  guides(color = guide_legend(reverse = TRUE),
         linetype = guide_legend(reverse = TRUE))
# in the bias-variance tradeoff, truncated stabilized weights seem to have too much bias, almost identical to unweighted
ggsave("wts-compare.png",width=9,height=5)

```

## Model selection function
```{r}
#############################
# function to do forward or backward selection by AIC - more customized than MASS::stepAIC()
## output: returns the selected model formula
#############################
## fit: the model object, eg, lm, glm
## consider_vars: a named list where each element contains a possible change to the formula of fit. Should include + or - to indicate whether each variable is being added or removed from the formula.
mod_select = function(fit,consider_vars) {
  continue=T
  while (continue==T) {
    print("Current model:")
    print(formula(fit))
    print(paste("Current AIC:",AIC(fit)))
    mod_test = lapply(consider_vars,function(var) {
      # update the model
      fit_test = stats::update(fit,as.formula(paste0(". ~ .",var)))
      return(fit_test)
    })
    AIC_test = lapply(mod_test,function(mod) return(AIC(mod)))
    if ((round(AIC(fit))-3) > round(min(as.numeric(AIC_test)))) {
      select_var_name = names(consider_vars)[which.min(as.numeric(AIC_test))]
      print(paste("Selected variable for addition/removal:",select_var_name))
      select_var_full = consider_vars[select_var_name]
      fit = mod_test[[which(names(consider_vars)==select_var_name)]] #stats::update(fit,as.formula(paste0("~ .",select_var_full)))
      consider_vars = consider_vars[names(consider_vars) != select_var_name]
      continue=T
    } else {
      continue=F
      print("No variables selected for addition/removal")
    }
    if (length(consider_vars)==0) {print("No variables left to consider");continue=F}
  }
  print("Final fit:")
  print(formula(fit))
  print(paste("Final AIC:",AIC(fit)))
  return(formula(fit))
}
```

## Trim extreme values for model fitting
```{r trim}
# trim LVAD time to 10 years
combined_df = combined_df %>% mutate(
    lvad_years_at_start = ifelse(lvad_years_at_start>10,10,lvad_years_at_start))

# trim other continuous variables to 1st and 99th percentiles
trimOuter = function(x) {
  q99 = quantile(x,.99,na.rm=T)
  q01 = quantile(x,.01,na.rm=T)
  
  x[x>q99] = q99
  x[x<q01] = q01
  return(x)
}
combined_df = combined_df %>%
  mutate_at(vars(albumin_at_start,bilirubin_at_start,eGFR_at_start,
              sodium_at_start,cpo_at_start,api_at_start,papi_at_start),
            trimOuter) %>%
  group_by(BNP_NT_Pro_at_start) %>%
  mutate_at(vars(BNP_at_start),trimOuter) %>% ungroup
trim_values = bind_rows(
  combined_df %>%
    summarise_at(vars(albumin_at_start,bilirubin_at_start,eGFR_at_start,sodium_at_start,cpo_at_start,api_at_start,papi_at_start),.funs=c("min"=min,"max"=max)) %>%
    pivot_longer(cols = everything(), 
         names_to = c("var", ".value"), names_sep="_at_start_"),
  combined_df %>%
    group_by(BNP_NT_Pro_at_start) %>%
    summarise_at(vars(BNP_at_start),.funs=c("min"=min,"max"=max)) %>%
    mutate(var="BNP"))
write.csv(trim_values,file="trim_values.csv",row.names=F)
```

## Save training data for modeling
```{r}
save(combined_df,file="mortality_model_train_data.Rdata")
```

## Model selection
With the stabilized weights, starting from the US-CRS 1.0 base model, fit the following models:

   - Model A: US-CRS 1.0
   
   - Model B: US-CRS 1.0 + forward selection of nonlinear effects
   
   - Model C: US-CRS 1.0 + current Impella + current IABP
   
   - Model D: US-CRS 1.0 + ever Impella + ever IABP
   
   - Model E: US-CRS 1.0 with nonlinear effects (Model B) + forward selection of new main effects (Impella, IABP, time on LVAD, API, PAPi, CPO, diagnosis, CHD only)
   
   - Model F: Model E + forward selection of nonlinear effects for new variables
   
   - Model G: Model F + backward selection of 2-way interactions (between devices and labs, devices and selected hemodynamics)
   
   - Model H: Relaxed lasso
```{r}
### MODEL A: US-CRS 1.0 ###
# fit base model w/ stabilized weights
mortality_modelA = glm(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start, data = combined_df, weights = sweight, family = binomial)
AIC(mortality_modelA) # 73145.38
length(coef(mortality_modelA)) # 10

### MODEL C: US-CRS 1.0 + current impella + current iabp ###
mortality_modelC = glm(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + iabp_at_start + impella_at_start, data = combined_df, weights = sweight, family = binomial)
AIC(mortality_modelC) # 72956.31
length(coef(mortality_modelC)) # 12

### MODEL D: US-CRS 1.0 + ever impella + ever iabp ###
mortality_modelD = glm(
  outcome ~ 
    albumin_at_start + log(bilirubin_at_start + 1) + eGFR_at_start + log(BNP_at_start + 1)*BNP_NT_Pro_at_start + sodium_at_start + durable_LVAD_at_start + short_mcs_ever_at_start + iabp_ever_at_start + impella_ever_at_start, data = combined_df, weights = sweight, family = binomial)
AIC(mortality_modelD) # 73036.53
length(coef(mortality_modelD)) # 12

### MODEL B: US-CRS 1.0 + forward selection of nonlinear effects ###
mort_mod_selectB = mod_select(mortality_modelA,list(
  albumin_ns = "-albumin_at_start+ns(albumin_at_start,df=3)",
  bilirubin_ns = "-log(bilirubin_at_start+1)+ns(log(bilirubin_at_start+1),df=3)",
  eGFR_ns = "-eGFR_at_start+ns(eGFR_at_start,df=3)",
  BNP_ns = "-log(BNP_at_start + 1)*BNP_NT_Pro_at_start+ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start",
  sodium_ns = "-sodium_at_start+ns(sodium_at_start,df=3)"
  ))
mortality_modelB = stats::update(mortality_modelA,mort_mod_selectB)
AIC(mortality_modelB) # 72862.55
length(coef(mortality_modelB)) # 22
print(formula(mortality_modelB))
# outcome ~ durable_LVAD_at_start + short_mcs_ever_at_start + ns(eGFR_at_start, 
    # df = 3) + ns(sodium_at_start, df = 3) + ns(log(BNP_at_start + 
    # 1), df = 3) + BNP_NT_Pro_at_start + ns(albumin_at_start, 
    # df = 3) + ns(log(bilirubin_at_start + 1), df = 3) + ns(log(BNP_at_start + 
    # 1), df = 3):BNP_NT_Pro_at_start

### Model E ###
# US-CRS 1.0 with nonlinear effects (Model B) + forward selection of new main effects (Impella, IABP, time on LVAD, API, PAPi, CPO, diagnosis, CHD only)

combined_df = combined_df %>% 
  mutate(chd_at_start = case_when(
    diagnosis_at_start %in% "Congenital" ~ 1,
    T ~ 0,
  ))

mort_mod_selectE = mod_select(mortality_modelB,list(
  week="+factor(week_from_baseline)",
  impella="+impella_at_start",
  iabp="+iabp_at_start",
  diagnosis="+diagnosis_at_start",
  chd="+chd_at_start",
  lvad_time = "+lvad_years_at_start",
  cpo="+cpo_at_start",
  api="+api_at_start",
  papi="+papi_at_start"
  ))
mortality_modelE = stats::update(mortality_modelA,mort_mod_selectE)
AIC(mortality_modelE) # 72249.18
length(coef(mortality_modelE)) # 40
print(formula(mortality_modelE))
# outcome ~ durable_LVAD_at_start + short_mcs_ever_at_start + ns(eGFR_at_start, 
#     df = 3) + ns(sodium_at_start, df = 3) + ns(log(BNP_at_start + 
#     1), df = 3) + BNP_NT_Pro_at_start + ns(albumin_at_start, 
#     df = 3) + ns(log(bilirubin_at_start + 1), df = 3) + lvad_years_at_start + 
#     iabp_at_start + api_at_start + diagnosis_at_start + 
#     impella_at_start + factor(week_from_baseline) + cpo_at_start + 
#     ns(log(BNP_at_start + 1), df = 3):BNP_NT_Pro_at_start

### Model F ###
# Model E + forward selection of nonlinear effects for new variables
## lvad_years_at_start, api_at_start, cpo_at_start
mort_mod_selectF = mod_select(mortality_modelE,list(
  lvad_ns = "-lvad_years_at_start+ns(lvad_years_at_start,df=3,knots=c(1,3))",
  cpo_ns="-cpo_at_start+ns(cpo_at_start,df=3)",
  api_ns="-api_at_start+ns(api_at_start,df=3)"
  ))
mortality_modelF = stats::update(mortality_modelA,mort_mod_selectF)
AIC(mortality_modelF) # 72102.94
length(coef(mortality_modelF)) # 46
print(formula(mortality_modelF))
# outcome ~ durable_LVAD_at_start + short_mcs_ever_at_start + ns(eGFR_at_start, 
#     df = 3) + ns(sodium_at_start, df = 3) + ns(log(BNP_at_start + 
#     1), df = 3) + BNP_NT_Pro_at_start + ns(albumin_at_start, 
#     df = 3) + ns(log(bilirubin_at_start + 1), df = 3) + iabp_at_start + 
#     diagnosis_at_start + impella_at_start + factor(week_from_baseline) + 
#     ns(lvad_years_at_start, df = 3, knots = c(1, 3)) + ns(api_at_start, 
#     df = 3) + ns(cpo_at_start, df = 3) + ns(log(BNP_at_start + 
#     1), df = 3):BNP_NT_Pro_at_start


### Model G ###

# Model F + forward selection of 2-way interactions (between devices and labs, devices and selected hemodynamics)
## NOTE: Didn't finish due to maxxing out on memory - vector could not be allocated
# mort_mod_selectG = mod_select(mortality_modelF,list(
#   lvad_alb = "+durable_LVAD_at_start:ns(albumin_at_start,df=3)",
#   lvad_bili = "+durable_LVAD_at_start:ns(log(bilirubin_at_start+1),df=3)",
#   lvad_eGFR = "+durable_LVAD_at_start:ns(eGFR_at_start,df=3)",
#   lvad_sod = "+durable_LVAD_at_start:ns(sodium_at_start,df=3)",
#   lvad_bnp = "-log(BNP_at_start + 1)*BNP_NT_Pro_at_start - durable_LVAD_at_start + ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start*durable_LVAD_at_start",
#   lvad_api = "+durable_LVAD_at_start:ns(api_at_start,df=3)",
#   lvad_cpo = "+durable_LVAD_at_start:ns(cpo_at_start,df=3)",
#   impella_alb = "+impella_ever_at_start:ns(albumin_at_start,df=3)",
#   impella_bili = "+impella_ever_at_start:ns(log(bilirubin_at_start+1),df=3)",
#   impella_eGFR = "+impella_ever_at_start:ns(eGFR_at_start,df=3)",
#   impella_sod = "+impella_ever_at_start:ns(sodium_at_start,df=3)",
#   impella_bnp = "-log(BNP_at_start + 1)*BNP_NT_Pro_at_start - impella_ever_at_start + ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start*impella_ever_at_start",
#   impella_api = "+impella_ever_at_start:ns(api_at_start,df=3)",
#   impella_cpo = "+impella_ever_at_start:ns(cpo_at_start,df=3)",
#   iabp_alb = "+iabp_ever_at_start:ns(albumin_at_start,df=3)",
#   iabp_bili = "+iabp_ever_at_start:ns(log(bilirubin_at_start+1),df=3)",
#   iabp_eGFR = "+iabp_ever_at_start:ns(eGFR_at_start,df=3)",
#   iabp_sod = "+iabp_ever_at_start:ns(sodium_at_start,df=3)",
#   iabp_bnp = "-log(BNP_at_start + 1)*BNP_NT_Pro_at_start - iabp_ever_at_start + ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start*iabp_ever_at_start",
#   iabp_api = "+iabp_ever_at_start:ns(api_at_start,df=3)",
#   iabp_cpo = "+iabp_ever_at_start:ns(cpo_at_start,df=3)"
#   ))
## Result before error:
# LVAD*BNP, IABP*bilirubin, LVAD*API, Impella*sodium, ERROR
# Final AIC: 71968



### Model G ###
# Model F + backward selection of 2-way interactions (between devices and labs, devices and selected hemodynamics)

mortality_modelG_start = glm(
  outcome ~ 
    factor(week_from_baseline) + 
    (durable_LVAD_at_start + short_mcs_ever_at_start + iabp_at_start + impella_at_start) *
    (ns(eGFR_at_start, df = 3) + ns(sodium_at_start, df = 3) + 
       ns(log(BNP_at_start + 1), df = 3)*BNP_NT_Pro_at_start + 
       ns(albumin_at_start,df = 3) +
       ns(log(bilirubin_at_start + 1), df = 3) + 
       ns(api_at_start,df=3) + ns(cpo_at_start,df=3)) +
    diagnosis_at_start + 
    ns(lvad_years_at_start,df=3,knots=c(1,3)),
  data = combined_df,
  weights = sweight, family = binomial)
gc()

## Wald test for interaction significance

# get variance-covariance matrix for coefficients
mortality_modelG_VCOV = vcovCL(mortality_modelG_start, cluster = ~ PX_ID)
gc()

#############################
# function to perform a Wald test for a vector of betas when null hypothesis is no effect
#############################
## mod: the model object, eg, lm, glm, coxph
## vars: names of the coefficients to test null hypothesis of no effect
## cov: covariance matrix for the coefficients; default is vcov(mod) but could be provided by the user (eg, an empirical covariance matrix derived from bootstrapping)
wald_test = function(mod,vars,cov=vcov(mod)) {
  c = matrix(0,ncol=length(coef(mod)),nrow=length(vars))
  colnames(c) = names(coef(mod))
  c[,vars] = diag(length(vars))
  teststat = t(c %*% coef(mod)) %*% solve(c %*% cov %*% t(c)) %*% (c %*% coef(mod))
  P = pchisq(teststat,df=length(vars),lower.tail=F)
  return(P)
}

# test interactions

## LVAD - only API
wald_test(mortality_modelG_start,paste0("durable_LVAD_at_start:ns(eGFR_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .059
wald_test(mortality_modelG_start,paste0("durable_LVAD_at_start:ns(sodium_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .112
wald_test(mortality_modelG_start,c(paste0("durable_LVAD_at_start:ns(log(BNP_at_start + 1), df = 3)",1:3),"durable_LVAD_at_start:BNP_NT_Pro_at_start1"),mortality_modelG_VCOV) # .154
wald_test(mortality_modelG_start,paste0("durable_LVAD_at_start:ns(albumin_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .842
wald_test(mortality_modelG_start,paste0("durable_LVAD_at_start:ns(log(bilirubin_at_start + 1), df = 3)",1:3),mortality_modelG_VCOV) # .271
wald_test(mortality_modelG_start,paste0("durable_LVAD_at_start:ns(api_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .001
wald_test(mortality_modelG_start,paste0("durable_LVAD_at_start:ns(cpo_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .463

## Short-term MCS (ECMO) - BNP, albumin, API, CPO
wald_test(mortality_modelG_start,paste0("short_mcs_ever_at_start:ns(eGFR_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .274
wald_test(mortality_modelG_start,paste0("short_mcs_ever_at_start:ns(sodium_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .158
wald_test(mortality_modelG_start,c(paste0("short_mcs_ever_at_start:ns(log(BNP_at_start + 1), df = 3)",1:3),"short_mcs_ever_at_start:BNP_NT_Pro_at_start1"),mortality_modelG_VCOV) # .02
wald_test(mortality_modelG_start,paste0("short_mcs_ever_at_start:ns(albumin_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .019
wald_test(mortality_modelG_start,paste0("short_mcs_ever_at_start:ns(log(bilirubin_at_start + 1), df = 3)",1:3),mortality_modelG_VCOV) # .70
wald_test(mortality_modelG_start,paste0("short_mcs_ever_at_start:ns(api_at_start, df = 3)",1:3),mortality_modelG_VCOV) # <.001
wald_test(mortality_modelG_start,paste0("short_mcs_ever_at_start:ns(cpo_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .03

## Impella - only sodium
wald_test(mortality_modelG_start,paste0("impella_at_start:ns(eGFR_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .50
wald_test(mortality_modelG_start,paste0("impella_at_start:ns(sodium_at_start, df = 3)",1:3),mortality_modelG_VCOV) # <.001
wald_test(mortality_modelG_start,c(paste0("impella_at_start:ns(log(BNP_at_start + 1), df = 3)",1:3),"impella_at_start:BNP_NT_Pro_at_start1"),mortality_modelG_VCOV) # .92
wald_test(mortality_modelG_start,paste0("impella_at_start:ns(albumin_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .07
wald_test(mortality_modelG_start,paste0("impella_at_start:ns(log(bilirubin_at_start + 1), df = 3)",1:3),mortality_modelG_VCOV) # .08
wald_test(mortality_modelG_start,paste0("impella_at_start:ns(api_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .12
wald_test(mortality_modelG_start,paste0("impella_at_start:ns(cpo_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .23

## IABP - only bilirubin
wald_test(mortality_modelG_start,paste0("iabp_at_start:ns(eGFR_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .63
wald_test(mortality_modelG_start,paste0("iabp_at_start:ns(sodium_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .25
wald_test(mortality_modelG_start,c(paste0("iabp_at_start:ns(log(BNP_at_start + 1), df = 3)",1:3),"iabp_at_start:BNP_NT_Pro_at_start1"),mortality_modelG_VCOV) # .74
wald_test(mortality_modelG_start,paste0("iabp_at_start:ns(albumin_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .70
wald_test(mortality_modelG_start,paste0("iabp_at_start:ns(log(bilirubin_at_start + 1), df = 3)",1:3),mortality_modelG_VCOV) # .04
wald_test(mortality_modelG_start,paste0("iabp_at_start:ns(api_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .5
wald_test(mortality_modelG_start,paste0("iabp_at_start:ns(cpo_at_start, df = 3)",1:3),mortality_modelG_VCOV) # .98

# fit Model G w/ selected interactions
mortality_modelG = glm(
  outcome ~ 
    factor(week_from_baseline) + 
    durable_LVAD_at_start*ns(api_at_start,df=3) +
    short_mcs_ever_at_start*(
      ns(log(BNP_at_start + 1), df = 3)*BNP_NT_Pro_at_start + 
       ns(albumin_at_start,df = 3) +
        ns(api_at_start,df=3) + ns(cpo_at_start,df=3)) +
    iabp_at_start*ns(log(bilirubin_at_start + 1), df = 3) +
    impella_at_start*ns(sodium_at_start, df = 3) +
    ns(eGFR_at_start, df = 3) +
    diagnosis_at_start + 
    ns(lvad_years_at_start,df=3,knots=c(1,3)),
  data = combined_df,
  weights = sweight, family = binomial)
gc()

AIC(mortality_modelG) # 71622.36
length(coef(mortality_modelG)) # 71
print(formula(mortality_modelG))
# outcome ~ factor(week_from_baseline) + durable_LVAD_at_start * 
#     ns(api_at_start, df = 3) + short_mcs_ever_at_start * (ns(log(BNP_at_start + 
#     1), df = 3) * BNP_NT_Pro_at_start + ns(albumin_at_start, 
#     df = 3) + ns(api_at_start, df = 3) + ns(cpo_at_start, df = 3)) + 
#     iabp_at_start * ns(log(bilirubin_at_start + 1), df = 3) + 
#     impella_at_start * ns(sodium_at_start, df = 3) + ns(eGFR_at_start, 
#     df = 3) + diagnosis_at_start + ns(lvad_years_at_start, df = 3, 
#     knots = c(1, 3))

# save models
save(mortality_modelA,mortality_modelB,mortality_modelC,
     mortality_modelD,mortality_modelE,mortality_modelF,
     mortality_modelG,file="mortality-model-options.Rdata") # mortality_modelG_VCOV

```

### Model H: Relaxed lasso

```{r}
library(glmnet)

# set full formula
full_formula <- outcome ~ 
    factor(week_from_baseline) + 
    (durable_LVAD_at_start + short_mcs_ever_at_start + iabp_at_start + impella_at_start) *
    (ns(eGFR_at_start, df = 3) + ns(sodium_at_start, df = 3) + 
       ns(log(BNP_at_start + 1), df = 3)*BNP_NT_Pro_at_start + 
       ns(albumin_at_start,df = 3) +
       ns(log(bilirubin_at_start + 1), df = 3) + 
       ns(api_at_start,df=3) + ns(papi_at_start,df=3) + 
       ns(cpo_at_start,df=3)) +
    diagnosis_at_start + 
    ns(lvad_years_at_start,df=3,knots=c(1,3))

# which rows have nonmissing IPW?
lasso_rows = which(!is.na(combined_df$sweight))

# save the terms (knots) for prediction
mf_train <- model.frame(full_formula, data = combined_df[lasso_rows, ])
orig_terms <- attr(mf_train, "terms")

# put data in model.matrix
x_matrix <- model.matrix(orig_terms, data = combined_df)[, -1]

# fit CV lasso to select predictors
cv_fit <- cv.glmnet(x=x_matrix[lasso_rows,], 
                    y=combined_df$outcome[lasso_rows], 
                    weights=combined_df$sweight[lasso_rows],
                    family = "binomial",
                    alpha = 1)

# get coefficients at lambda.1se
lasso_coefs <- coef(cv_fit, s = "lambda.1se")
print(lasso_coefs[which(lasso_coefs!=0),])
selected_vars <- rownames(lasso_coefs)[which(lasso_coefs != 0)][-1]

# fit (unpenalized) glm with selected predictors
x_selected <- x_matrix[, selected_vars, drop=F]
mortality_modelH = glm(
  combined_df$outcome ~ x_selected,
  weights = combined_df$sweight, family = binomial)
AIC(mortality_modelH) # 71728.83
length(coef(mortality_modelH)) # 49

# save lasso model
save(mortality_modelH,file="mortality-model-lasso-options.Rdata")

```


## Function to calculate predicted survival
```{r}
## function to calculate predicted survival
### note: predicted 6-week survival will be calculated for each row of newdata; if the data are prepared for landmark modeling, they should be filtered to week 1 from each landmark; predictions will be made based on that row's covariates, assuming those are the covariates at the start of the 6-week period

## need in the environment for lasso:
# lasso_orig_terms
# lasso_selected_vars

lasso_orig_terms = orig_terms; lasso_selected_vars = selected_vars
pred_data_fun = function(newdata,mod,lasso=F,linear_spline=F) {
  # trim LVAD years to 10
  newdata = newdata %>%
    ungroup() %>%
    mutate(lvad_years_at_start =
             ifelse(lvad_years_at_start>10,10,lvad_years_at_start))
  
  # create an ID variable
  newdata$tmp_id = 1:nrow(newdata)
  
  if (linear_spline) {
    newdata = newdata %>%
      mutate(albumin_at_start3.5 = ifelse(albumin_at_start > 3.5, albumin_at_start - 3.5, 0),
         albumin_at_start4.25 = ifelse(albumin_at_start > 4.25, albumin_at_start - 4.25, 0),
         eGFR_at_start40 = ifelse(eGFR_at_start > 40, eGFR_at_start - 40, 0),
         eGFR_at_start90 = ifelse(eGFR_at_start > 90, eGFR_at_start - 90, 0),
         log_bili_at_start = log(bilirubin_at_start + 1),
         log_bili_at_start.47 = ifelse(log_bili_at_start > .47, log_bili_at_start - .47, 0),
         log_BNP_at_start = ifelse(BNP_NT_Pro_at_start==0,log(BNP_at_start + 1),0),
         log_BNP_at_start5.5 = ifelse(BNP_NT_Pro_at_start==0,
                                   ifelse(log_BNP_at_start>5.5,log_BNP_at_start-5.5,0),0),
         log_BNP_at_start7 = ifelse(BNP_NT_Pro_at_start==0,
                                   ifelse(log_BNP_at_start>7,log_BNP_at_start-7,0),0),
         log_BNP_NT_Pro_at_start = ifelse(BNP_NT_Pro_at_start==1,log(BNP_at_start + 1),0),
         log_BNP_NT_Pro_at_start5 = ifelse(BNP_NT_Pro_at_start==1,
                                           ifelse(log_BNP_NT_Pro_at_start>5,log_BNP_NT_Pro_at_start-5,0),0),
         sodium_at_start137 = ifelse(sodium_at_start>137,sodium_at_start-137,0)
  )
  }
  
  # expand each row to six rows and create week_from_baseline
  pred_data = newdata[rep(1:nrow(newdata), each = 6), ]  # Duplicate each row 6 times
  pred_data$week_from_baseline = rep(1:6, times = nrow(newdata))
  
  if (lasso) {
    mf_pred_data <- model.frame(delete.response(lasso_orig_terms), data=pred_data)
    x_pred_data <- model.matrix(delete.response(lasso_orig_terms), data = mf_pred_data)[, -1]
    x_pred_data_selected <- x_pred_data[, lasso_selected_vars, drop = F]
    pred_data$haz_dth <- predict(mod,newdata=list(x_selected = x_pred_data_selected),type="response")
  } else {
    # predict hazard of death within the week
    pred_data$haz_dth = predict(mod,newdata=pred_data,type="response")
  }
  
  # calculate probability of surviving 6 weeks
  pred_data = pred_data %>%
    group_by(tmp_id) %>%
    mutate(prob_surv_6wk = cumprod(1-haz_dth),
           haz_dth_nowk = first(haz_dth)) %>%
    slice(n()) %>%
    ungroup() %>%
    mutate(week_from_baseline=1,
           prob_outcome_6wk = 1 - prob_surv_6wk) %>% 
    select(-tmp_id)
  
  # return pred_data
  return(pred_data)
}
save(pred_data_fun,lasso_orig_terms,lasso_selected_vars,file="mort-pred-fun.Rdata")

## get predicted 6-week survival from training data, including all landmarks
# pred_data = pred_data_fun(combined_df %>% filter(week_from_baseline==1))
# summary(pred_data$prob_surv_6wk)
# quantile(pred_data$prob_surv_6wk,probs=seq(0,.25,.05))

```

Before the next step, the models are compared in the validation data, in the model_eval.Rmd script.

Models C and E had the best performance in validation data and will be compared in the test data.

## Convert natural cubic splines to linear splines
For Model E, we need to refit with linear splines instead of natural cubic splines.
```{r}
# bilirubin, eGFR, sodium, albumin, BNP

## create pred_data object for lab value ranges
pred_data = combined_df %>% filter(week_from_baseline==1)

## calculate vcov for Model E
mortality_modelE_VCOV = vcovCL(mortality_modelE, cluster = ~ PX_ID)

## bilirubin
bili_range = with(pred_data,seq(min(log(bilirubin_at_start+1)),max(log(bilirubin_at_start+1)),length.out=200))
bili_spl = ns(bili_range,df=3,
              knots=attr(mortality_modelE$model$`ns(log(bilirubin_at_start + 1), df = 3)`,"knots"),
              Boundary.knots=attr(mortality_modelE$model$`ns(log(bilirubin_at_start + 1), df = 3)`,"Boundary.knots"))
bili_rr = cbind.data.frame(
  bilirubin_trans = bili_range,
  do.call(rbind.data.frame,apply(bili_spl,1,function(row) {
              getHR(mortality_modelE,
                vars=names(coef(mortality_modelE))[grepl("bili",names(coef(mortality_modelE)))][1:3],
                comb=row,
                cov=mortality_modelE_VCOV)}))) %>%
  mutate(bilirubin=exp(bilirubin_trans)-1)

## eGFR
egfr_range = with(pred_data,seq(min(eGFR_at_start),max(eGFR_at_start),length.out=200))
egfr_spl = ns(egfr_range,df=3,
              knots=attr(mortality_modelE$model$`ns(eGFR_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_modelE$model$`ns(eGFR_at_start, df = 3)`,"Boundary.knots"))
egfr_rr = cbind.data.frame(
  egfr = egfr_range,
  do.call(rbind.data.frame,apply(egfr_spl,1,function(row) {
              getHR(mortality_modelE,
                vars=names(coef(mortality_modelE))[grepl("eGFR",names(coef(mortality_modelE)))][1:3],
                comb=row,
                cov=mortality_modelE_VCOV)})))

# sodium
sod_range = with(pred_data,seq(min(sodium_at_start),max(sodium_at_start),length.out=200))
sod_spl = ns(sod_range,df=3,
              knots=attr(mortality_modelE$model$`ns(sodium_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_modelE$model$`ns(sodium_at_start, df = 3)`,"Boundary.knots"))
sod_rr = cbind.data.frame(
  sodium = sod_range,
  do.call(rbind.data.frame,apply(sod_spl,1,function(row) {
              getHR(mortality_modelE,
                vars=names(coef(mortality_modelE))[grepl("sodium",names(coef(mortality_modelE)))][1:3],
                comb=row,
                cov=mortality_modelE_VCOV)})))

## albumin
alb_range = with(pred_data,seq(min(albumin_at_start),max(albumin_at_start),length.out=200))
alb_spl = ns(alb_range,df=3,
              knots=attr(mortality_modelE$model$`ns(albumin_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_modelE$model$`ns(albumin_at_start, df = 3)`,"Boundary.knots"))
alb_rr = cbind.data.frame(
  albumin = alb_range,
  do.call(rbind.data.frame,apply(alb_spl,1,function(row) {
              getHR(mortality_modelE,
                vars=names(coef(mortality_modelE))[grepl("albumin",names(coef(mortality_modelE)))][1:3],
                comb=row,
                cov=mortality_modelE_VCOV)})))

# BNP
bnp_range = with(pred_data %>% filter(BNP_NT_Pro_at_start==0),
                  seq(min(log(BNP_at_start+1)),max(log(BNP_at_start+1)),length.out=200))
bnp_nt_range = with(pred_data %>% filter(BNP_NT_Pro_at_start==1),
                  seq(min(log(BNP_at_start+1)),max(log(BNP_at_start+1)),length.out=200))
bnp_spl = ns(c(bnp_range,bnp_nt_range),df=3,
              knots=attr(mortality_modelE$model$`ns(log(BNP_at_start + 1), df = 3)`,"knots"),
              Boundary.knots=attr(mortality_modelE$model$`ns(log(BNP_at_start + 1), df = 3)`,"Boundary.knots"))
bnp_rr = cbind.data.frame(
  bnp_trans = c(bnp_range,bnp_nt_range),
  bnp_nt_pro = c(rep(0,length(bnp_range)),rep(1,length(bnp_nt_range))),
  rbind.data.frame(
    # BNP_NT_Pro=0
    do.call(rbind.data.frame,apply(bnp_spl[1:length(bnp_range),],1,function(row) {
                getHR(mortality_modelE,
                  vars=names(coef(mortality_modelE))[grepl("BNP",names(coef(mortality_modelE))) & !grepl("BNP_NT",names(coef(mortality_modelE)))][1:3],
                  comb=row,
                  cov=mortality_modelE_VCOV)})),
    # BNP_NT_Pro=1
    do.call(rbind.data.frame,apply(bnp_spl[(length(bnp_range)+1):(length(bnp_range)+length(bnp_nt_range)),],1,function(row) {
                getHR(mortality_modelE,
                  vars=names(coef(mortality_modelE))[grepl("BNP",names(coef(mortality_modelE)))][1:7],
                  # note: subtracting off the BNP_NT_Pro intercept; start at 0 at minimum BNP NT Pro
                  comb=c(row,1,row)-c(bnp_spl[(length(bnp_range)+1),],1,bnp_spl[(length(bnp_range)+1)]),
                  cov=mortality_modelE_VCOV)})))) %>%
  mutate(bnp=exp(bnp_trans)-1)

# plot them all together
ggplot(bind_rows(bili_rr %>% mutate(value=bilirubin_trans,lab="ln(Bilirubin)"),
                 egfr_rr %>% mutate(value=egfr,lab="eGFR"),
                 bnp_rr %>% mutate(value=bnp_trans,lab=case_when(
                   bnp_nt_pro==0 ~ "ln(BNP)",T ~ "ln(NT-proBNP)")),
                 alb_rr %>% mutate(value=albumin,lab="Albumin"),
                 sod_rr %>% mutate(value=sodium,lab="Sodium")),
       aes(x=value,y=hr)) +
  geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
  geom_line(color="black",linewidth=.75) +
  geom_hline(yintercept=1,linetype=2) +
  scale_y_continuous(trans='log2',minor_breaks=NULL,
                     breaks=2^((-3):6),limits=c(.125/2,32),
                     oob = scales::rescale_none) +
  facet_wrap(~lab,scales="free_x",nrow=2) +
  scale_fill_viridis_d() +
  # scale_x_continuous(breaks=seq(0,1,.2),limits=c(0,1),oob = scales::rescale_none) +
  labs(x="Lab value",y="Adjusted hazard ratio (95% CI)") +
  theme_bw() +
  theme(text=element_text(size=14))
# ggsave("lab-coefs-by-lvad-NEW.png",width=7.5,height=5,dpi=300)

# now plot the XB (not exponentiated) with no y log scale
ggplot(bind_rows(bili_rr %>% mutate(value=bilirubin_trans,lab="ln(Bilirubin)"),
                 egfr_rr %>% mutate(value=egfr,lab="eGFR"),
                 bnp_rr %>% mutate(value=bnp_trans,lab=case_when(
                   bnp_nt_pro==0 ~ "ln(BNP)",T ~ "ln(NT-proBNP)")),
                 alb_rr %>% mutate(value=albumin,lab="Albumin"),
                 sod_rr %>% mutate(value=sodium,lab="Sodium")),# %>%
         # filter(lab=="Albumin"),
       aes(x=value,y=log(hr))) +
  geom_ribbon(aes(ymin = log(lowCI), ymax = log(highCI)),alpha=.5) +
  geom_line(color="black",linewidth=.75) +
  geom_hline(yintercept=1,linetype=2) +
  scale_y_continuous(#trans='log2',
                     minor_breaks=NULL,
                     #breaks=2^((-3):6),
                     # limits=c(-2.5,0),
                     oob = scales::rescale_none) +
  facet_wrap(~lab,scales="free",nrow=2) +
  scale_x_continuous(oob = scales::rescale_none) +
  labs(x="Lab value",y="Partial (95% CI)") +
  theme_bw() +
  theme(text=element_text(size=14))

## knots for linear splines:
# albumin: 3.5, 4.25
# eGFR: 40, 90
# log bili: 0.47
# log BNP: 5.5, 7
# log NT-proBNP: 5, 5.5
# sodium: 137

# refit mortality_modelE with linear splines
combined_df = combined_df %>%
  mutate(albumin_at_start3.5 = ifelse(albumin_at_start > 3.5, albumin_at_start - 3.5, 0),
         albumin_at_start4.25 = ifelse(albumin_at_start > 4.25, albumin_at_start - 4.25, 0),
         eGFR_at_start40 = ifelse(eGFR_at_start > 40, eGFR_at_start - 40, 0),
         eGFR_at_start90 = ifelse(eGFR_at_start > 90, eGFR_at_start - 90, 0),
         log_bili_at_start = log(bilirubin_at_start + 1),
         log_bili_at_start.47 = ifelse(log_bili_at_start > .47, log_bili_at_start - .47, 0),
         log_BNP_at_start = ifelse(BNP_NT_Pro_at_start==0,log(BNP_at_start + 1),0),
         log_BNP_at_start5.5 = ifelse(BNP_NT_Pro_at_start==0,
                                   ifelse(log_BNP_at_start>5.5,log_BNP_at_start-5.5,0),0),
         log_BNP_at_start7 = ifelse(BNP_NT_Pro_at_start==0,
                                   ifelse(log_BNP_at_start>7,log_BNP_at_start-7,0),0),
         log_BNP_NT_Pro_at_start = ifelse(BNP_NT_Pro_at_start==1,log(BNP_at_start + 1),0),
         log_BNP_NT_Pro_at_start5 = ifelse(BNP_NT_Pro_at_start==1,
                                           ifelse(log_BNP_NT_Pro_at_start>5,log_BNP_NT_Pro_at_start-5,0),0),
         # log_BNP_NT_Pro_at_start5.5 = ifelse(BNP_NT_Pro_at_start==1,
         #                           ifelse(log_BNP_NT_Pro_at_start>5.5,log_BNP_NT_Pro_at_start-5.5,0),0),
         sodium_at_start137 = ifelse(sodium_at_start>137,sodium_at_start-137,0)
  )
mortality_modelE_linear = glm(
  outcome ~ 
    factor(week_from_baseline) + 
    durable_LVAD_at_start +
    lvad_years_at_start +
    iabp_at_start +
    impella_at_start + 
    short_mcs_ever_at_start +
    api_at_start + cpo_at_start +
    diagnosis_at_start +
    albumin_at_start + albumin_at_start3.5 + albumin_at_start4.25 +
    eGFR_at_start + eGFR_at_start40 + eGFR_at_start90 +
    log_bili_at_start + log_bili_at_start.47 +
    log_BNP_at_start + log_BNP_at_start5.5 + log_BNP_at_start7 +
    BNP_NT_Pro_at_start +
    log_BNP_NT_Pro_at_start + log_BNP_NT_Pro_at_start5 + #log_BNP_NT_Pro_at_start5.5 +
    sodium_at_start + sodium_at_start137,
  data = combined_df,
  weights = sweight, family = binomial)
mortality_modelE_linear
mortality_modelE_linearVCOV = vcovCL(mortality_modelE_linear, cluster = ~ PX_ID)

# calculate RRs with linear spline model

# 1. Bilirubin (Linear)
bili_rr_lin = cbind.data.frame(
  bilirubin_trans = bili_range,
  do.call(rbind.data.frame,
          apply(data.frame(bili_range - min(bili_range), 
                           pmax(bili_range - 0.47, 0)), 1, function(row) {
                  getHR(mortality_modelE_linear,
                        vars=c("log_bili_at_start", "log_bili_at_start.47"),
                        comb=as.numeric(row),
                        cov=mortality_modelE_linearVCOV)}))
)

# 2. eGFR (Linear)
egfr_rr_lin = cbind.data.frame(
  egfr = egfr_range,
  do.call(rbind.data.frame,
          apply(data.frame(egfr_range - min(egfr_range), 
                           pmax(egfr_range - 40, 0), 
                           pmax(egfr_range - 90, 0)), 1, function(row) {
                  getHR(mortality_modelE_linear,
                        vars=c("eGFR_at_start", "eGFR_at_start40", "eGFR_at_start90"),
                        comb=as.numeric(row),
                        cov=mortality_modelE_linearVCOV)}))
)

# 3. BNP (Linear - Non-NT subset)
bnp_rr_lin = cbind.data.frame(
  bnp_trans = bnp_range,
  do.call(rbind.data.frame,
          apply(data.frame(bnp_range - min(bnp_range), 
                           pmax(bnp_range - 5.5, 0), 
                           pmax(bnp_range - 7, 0)), 1, function(row) {
                  getHR(mortality_modelE_linear,
                        vars=c("log_BNP_at_start", "log_BNP_at_start5.5", "log_BNP_at_start7"),
                        comb=as.numeric(row),
                        cov=mortality_modelE_linearVCOV)}))
)

# 4. NT-proBNP (Linear)
# no intercept for NT-proBNP for this plot; start at 0
bnp_nt_rr_lin = cbind.data.frame(
  bnp_nt_trans = bnp_nt_range,
  do.call(rbind.data.frame,
          apply(data.frame(bnp_nt_range - min(bnp_nt_range), 
                           pmax(bnp_nt_range - 5, 0)), 1, function(row) {
                  getHR(mortality_modelE_linear,
                        vars=c("log_BNP_NT_Pro_at_start", 
                               "log_BNP_NT_Pro_at_start5"),
                        comb=row,
                        cov=mortality_modelE_linearVCOV)}))
)

# 5. Albumin (Linear)
alb_rr_lin = cbind.data.frame(
  albumin = alb_range,
  do.call(rbind.data.frame,
          apply(data.frame(alb_range - min(alb_range), 
                           pmax(alb_range - 3.5, 0), 
                           pmax(alb_range - 4.25, 0)), 1, function(row) {
                  getHR(mortality_modelE_linear,
                        vars=c("albumin_at_start", "albumin_at_start3.5", "albumin_at_start4.25"),
                        comb=as.numeric(row),
                        cov=mortality_modelE_linearVCOV)}))
)

# 6. Sodium (Linear)
sod_rr_lin = cbind.data.frame(
  sodium = sod_range,
  do.call(rbind.data.frame,
          apply(data.frame(sod_range - min(sod_range), 
                           pmax(sod_range - 137, 0)), 1, function(row) {
                  getHR(mortality_modelE_linear,
                        vars=c("sodium_at_start", "sodium_at_start137"),
                        comb=as.numeric(row),
                        cov=mortality_modelE_linearVCOV)}))
)

# plot them all together
ggplot(bind_rows(bili_rr %>% mutate(value=bilirubin_trans,lab="ln(Bilirubin)",model="Cubic"),
                 bili_rr_lin %>% mutate(value=bilirubin_trans,lab="ln(Bilirubin)",model="Linear"),
                 egfr_rr %>% mutate(value=egfr,lab="eGFR",model="Cubic"),
                 egfr_rr_lin %>% mutate(value=egfr,lab="eGFR",model="Linear"),
                 bnp_rr %>% mutate(value=bnp_trans,lab=case_when(
                   bnp_nt_pro==0 ~ "ln(BNP)",T ~ "ln(NT-proBNP)"),model="Cubic"),
                 bnp_rr_lin %>% mutate(value=bnp_trans,lab="ln(BNP)",model="Linear"),
                 bnp_nt_rr_lin %>% mutate(value=bnp_nt_trans,lab="ln(NT-proBNP)",model="Linear"),
                 alb_rr %>% mutate(value=albumin,lab="Albumin",model="Cubic"),
                 alb_rr_lin %>% mutate(value=albumin,lab="Albumin",model="Linear"),
                 sod_rr %>% mutate(value=sodium,lab="Sodium",model="Cubic"),
                 sod_rr_lin %>% mutate(value=sodium,lab="Sodium",model="Linear")),
       aes(x=value,y=hr, linetype=model, fill=model)) +
  geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
  geom_line(color="black",linewidth=.75) +
  geom_hline(yintercept=1,linetype=2) +
  scale_y_continuous(trans='log2',minor_breaks=NULL,
                     breaks=2^((-3):6),limits=c(.125,8),
                     oob = scales::rescale_none) +
  facet_wrap(~lab,scales="free_x",nrow=2) +
  scale_fill_viridis_d() + # direction=-1
  # scale_x_continuous(breaks=seq(0,1,.2),limits=c(0,1),oob = scales::rescale_none) +
  labs(x="Lab value",y="Adjusted hazard ratio (95% CI)",
       fill="Spline",linetype="Spline") +
  theme_bw() +
  theme(text=element_text(size=14))
ggsave("lab-coefs-by-splinetype.png",width=7.5,height=5,dpi=300)

# save mortality_modelE_linear
save(mortality_modelE_linear,file="mortality-modelE-linear.Rdata")

```





# Old

## Save final model and calculate robust standard errors
```{r}

mortality_model_full = glm(
  outcome ~ 
    factor(week_from_baseline) +
    ns(albumin_at_start,df=3) + 
    ns(log(bilirubin_at_start + 1),df=3)*durable_LVAD_at_start + 
    ns(eGFR_at_start,df=3)*durable_LVAD_at_start + 
    ns(log(BNP_at_start + 1),df=3)*BNP_NT_Pro_at_start*durable_LVAD_at_start + 
    ns(sodium_at_start,df=3) + 
    durable_LVAD_at_start + 
    short_mcs_ever_at_start +
    ns(lvad_years_at_start,df=3,knots=c(1,3)) + #,knots=c(6,8)) +
    ns(api_at_start,df=3) +
    ns(papi_at_start,df=3), 
  data = combined_df,
  weights = sweight, family = binomial)
AIC(mortality_model_full) # 62172.65

# look at model
summary(mortality_model_full)

# look at knots for LVAD years
attr(mortality_model_full$model$`ns(lvad_years_at_start, df = 3, knots = c(1, 3))`,"knots")
attr(mortality_model_full$model$`ns(lvad_years_at_start, df = 3, knots = c(1, 3))`,"Boundary.knots")

# make dataset of knots
knotsdat = lapply(1:length(colnames(mortality_model_full$model)),function(i) {
  coef_name = colnames(mortality_model_full$model)[i]
  inner.knots = attr((mortality_model_full$model)[[coef_name]],"knots")
  outer.knots = attr((mortality_model_full$model)[[coef_name]],"Boundary.knots")
  if (!is.null(inner.knots)) {
    return(data.frame(coef_name,knot1=outer.knots[1],knot2=inner.knots[1],
                    knot3=inner.knots[2],knot4=outer.knots[2]))
  }
})
knotsdat = do.call(rbind.data.frame,knotsdat)
write.csv(knotsdat,file="knots.csv",row.names=F)

# get robust SEs
mortality_model_full_VCOV = vcovCL(mortality_model_full, cluster = ~ PX_ID)
mortality_model_full_sum = do.call(rbind.data.frame,lapply(names(coef(mortality_model_full)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_full,vars=var_interest,cov=mortality_model_full_VCOV))
}))
mortality_model_full_sum
mortality_model_sum_coef = do.call(rbind.data.frame,lapply(names(coef(mortality_model_full)),function(var_interest) {
  bind_cols(var=var_interest,getHR(mortality_model_full,vars=var_interest,cov=mortality_model_full_VCOV,transform=F))
}))
mortality_model_sum_coef

# make supplemental table of coefficients
write.csv(mortality_model_sum_coef %>% rename(coef=xb) %>% mutate_at(.vars=vars(coef,se,lowCI,highCI),round,digits=3) %>% mutate(p=pval_format(p)),
          file="mortality-model-coefs.csv",row.names=F)
```

## Device hazard ratios: ECMO/surgical MCS, impella, IABP
```{r stmcs}
mortality_model_full_sum[mortality_model_full_sum$var=="short_mcs_ever_at_start",]

# model w/ impella and iabp
mort_impella_sum[mort_impella_sum$var=="short_mcs_ever_at_start",]
mort_impella_sum[mort_impella_sum$var=="impella_at_start",]
mort_impella_sum[mort_impella_sum$var=="iabp_at_start",]
```

## Durable LVAD hazard ratio at average lab values
```{r lvad}
# all the terms with durable LVAD
lvad_termnames = (names(coef(mortality_model_full)))[grepl("LVAD",names(coef(mortality_model_full)))]
lvad_termnames

## first get the splines for average lab values
knotsdat$coef_name
rownames(knotsdat) = c("albumin","bilirubin","eGFR","BNP","sodium","lvad_years","api","papi")
bili_avg_ns = ns(log(mean(combined_df$bilirubin_at_start)+1),3,knots=as.numeric(knotsdat["bilirubin",c("knot2","knot3")]),Boundary.knots=as.numeric(knotsdat["bilirubin",c("knot1","knot4")]))
egfr_avg_ns = ns(mean(combined_df$eGFR_at_start),3,knots=as.numeric(knotsdat["eGFR",c("knot2","knot3")]),Boundary.knots=as.numeric(knotsdat["eGFR",c("knot1","knot4")]))
bnp_avg_ns = ns(log(mean(combined_df$BNP_at_start[combined_df$BNP_NT_Pro==0])+1),3,knots=as.numeric(knotsdat["BNP",c("knot2","knot3")]),Boundary.knots=as.numeric(knotsdat["BNP",c("knot1","knot4")]))
sod_avg_ns = ns(mean(combined_df$sodium_at_start),3,knots=as.numeric(knotsdat["sodium",c("knot2","knot3")]),Boundary.knots=as.numeric(knotsdat["sodium",c("knot1","knot4")]))

# now get hazard ratio for LVAD (just started, years=0) at average lab values
lvad_termnames
getHR(mortality_model_full,
      vars=lvad_termnames,
      comb=c(1,bili_avg_ns,egfr_avg_ns,bnp_avg_ns,0,0,0,0),
      cov=mortality_model_full_VCOV)

```

## Calculate predicted survival without transplant in training data
```{r}
# get predicted 6-week survival from training data, including all landmarks
pred_data = pred_data_fun(combined_df %>% filter(week_from_baseline==1))
summary(pred_data$prob_surv_6wk)
quantile(pred_data$prob_surv_6wk,probs=seq(0,.25,.05))

## SAVE MORTALITY MODEL
# also, function to calculate predicted 6-week survival
# also, model for numerator of stabilized weights
save(mortality_model_full,pred_data_fun,tx_model_first,mort_impella,file="mortality-model.RData")
save(combined_df,file="mortality-df.RData")

# summarize survival probability by status
with(pred_data,tapply(prob_surv_6wk,status_at_start,summary))

# plot distributions
library(ggforce)
ggplot(pred_data %>% filter(!is.na(status),!(status=="inactive")),
       aes(x = status_at_start, y = prob_surv_6wk)) +
  geom_boxplot(alpha=.25) + 
  facet_zoom(y = prob_surv_6wk > .95) +  
  labs(x = "Status", y = "6-Week Survival Probability") +
  # title = "Figure 1: Box plots of predicted 6-week survival without transplant\nin adult heart candidates, 2019-2022, by status"
  theme_bw()
ggsave("pred-by-status-NEW.png",width=6,height=4)
```

## Calculate 6-week survival by time on LVAD
```{r}
# Map the 6-week probability to the status system
mortality_model_status = glm(
  outcome ~ status_at_start,
  data = combined_df,
  weights = weight, # in the absence of transplant
  family = binomial)
haz_dth_status = predict(mortality_model_status,newdata=data.frame(status_at_start=factor(1:6,levels=levels(combined_df$status_at_start))),type="response")
prob_surv_6wk_status = (1-haz_dth_status)^6
prob_surv_6wk_status

# how many LVAD years do you need to move from status 4 to status 3?

# set values for predicting
new_data_lvad <- data.frame(
  albumin_at_start = mean(pred_data$albumin_at_start[pred_data$durable_LVAD_at_start==1]),
  bilirubin_at_start = mean(pred_data$bilirubin_at_start[pred_data$durable_LVAD_at_start==1]),
  eGFR_at_start = mean(pred_data$eGFR_at_start[pred_data$durable_LVAD_at_start==1]),
  BNP_at_start = mean(pred_data$BNP_at_start[pred_data$durable_LVAD_at_start==1 & 
                                               pred_data$BNP_NT_Pro_at_start=="0"]),
  BNP_NT_Pro_at_start = "0",
  sodium_at_start = mean(pred_data$sodium_at_start[pred_data$durable_LVAD_at_start==1]),
  durable_LVAD_at_start = 1,
  short_mcs_ever_at_start = 0,
  cpo_at_start = mean(pred_data$cpo_at_start[pred_data$durable_LVAD_at_start==1]),
  api_at_start = mean(pred_data$api_at_start[pred_data$durable_LVAD_at_start==1]),
  papi_at_start = mean(pred_data$papi_at_start[pred_data$durable_LVAD_at_start==1]),
  lvad_years_at_start = seq(0,10,.2),
  impella_at_start=0,iabp_at_start=0
)
head(new_data_lvad)
pred_lvad <- pred_data_fun(new_data_lvad)

# make plot
## add distribution of years on LVAD
## add lines for Status 3 and 4
ggplot(pred_lvad, aes(x = lvad_years_at_start, y = prob_surv_6wk)) +
  geom_line(size = 1) +
  labs(
    #title = "Effect of LVAD years on 6-week survival probability in durable LVAD patients",
    x = "Years with LVAD",
    y = "6-Week Survival Probability"#survival without transplant"
  ) +
  theme_bw() + 
  coord_cartesian(xlim=c(0,10),
                  ylim=c(.97,1)) +
  scale_x_continuous(breaks=seq(0,10,1)) +
  geom_hline(yintercept=prob_surv_6wk_status["3"]) +
  geom_text(aes(4,prob_surv_6wk_status["3"],label="Status 3",vjust=-1)) +
  geom_hline(yintercept=prob_surv_6wk_status["4"]) +
  geom_text(aes(4,prob_surv_6wk_status["4"],label="Status 4",vjust=-1))
ggsave("pred-by-LVADtime-NEW.png",width=6,height=4)

# distribution of time on LVAD in training data, including all landmarks
summary(pred_data$lvad_years_at_start[pred_data$durable_LVAD_at_start==1])
```


## Summarize spline effects by LVAD vs not
```{r}
## LVAD interactions with: bilirubin, BNP, eGFR

## bilirubin
bili_range = with(pred_data,seq(min(log(bilirubin_at_start+1)),max(log(bilirubin_at_start+1)),length.out=200))
bili_spl = ns(bili_range,df=3,
              knots=attr(mortality_model_full$model$`ns(log(bilirubin_at_start + 1), df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(log(bilirubin_at_start + 1), df = 3)`,"Boundary.knots"))
bili_rr = cbind.data.frame(
  bilirubin_trans = rep(bili_range,2),
  durable_lvad = c(rep(0,length(bili_range)),rep(1,length(bili_range))),
  rbind.data.frame(
    do.call(rbind.data.frame,apply(bili_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("bili",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})),
    do.call(rbind.data.frame,apply(bili_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("bili",names(coef(mortality_model_full)))],
                  comb=rep(row,2),
                  cov=mortality_model_full_VCOV)})))) %>%
  mutate(bilirubin=exp(bilirubin_trans)-1)
# ggplot(bili_rr %>% mutate(durable_lvad=factor(durable_lvad)),
#        aes(x=exp(bilirubin_trans)-1,y=hr,linetype=durable_lvad,color=durable_lvad,fill=durable_lvad)) +
#   geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
#   geom_line(color="black",linewidth=.75) +
#   geom_hline(yintercept=1,linetype=2) +
#   scale_y_continuous(trans='log2',minor_breaks=NULL,
#                      breaks=c(.25,.5,1,2,4,8),limits=c(.25,8),
#                      oob = scales::rescale_none) +
#   scale_x_continuous(breaks=seq(0,1,.2),limits=c(0,1),oob = scales::rescale_none) +
#   labs(x="Bilirubin",y="Adjusted hazard ratio (95% CI)",
#        linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
#   theme_bw() +
#   theme(text=element_text(size=14))

## eGFR
egfr_range = with(pred_data,seq(min(eGFR_at_start),max(eGFR_at_start),length.out=200))
egfr_spl = ns(egfr_range,df=3,
              knots=attr(mortality_model_full$model$`ns(eGFR_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(eGFR_at_start, df = 3)`,"Boundary.knots"))
egfr_rr = cbind.data.frame(
  egfr = rep(egfr_range,2),
  durable_lvad = c(rep(0,length(egfr_range)),rep(1,length(egfr_range))),
  rbind.data.frame(
    do.call(rbind.data.frame,apply(egfr_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("eGFR",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})),
    do.call(rbind.data.frame,apply(egfr_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("eGFR",names(coef(mortality_model_full)))],
                  comb=rep(row,2),
                  cov=mortality_model_full_VCOV)}))))
# ggplot(egfr_rr %>% mutate(durable_lvad=factor(durable_lvad)),
#        aes(x=egfr,y=hr,linetype=durable_lvad,color=durable_lvad,fill=durable_lvad)) +
#   geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
#   geom_line(color="black",linewidth=.75) +
#   geom_hline(yintercept=1,linetype=2) +
#   scale_y_continuous(trans='log2',minor_breaks=NULL,
#                      breaks=c(.25,.5,1,2,4,8),limits=c(.25,8),
#                      oob = scales::rescale_none) +
#   scale_x_continuous(breaks=seq(0,120,20),limits=c(0,120),oob = scales::rescale_none) +
#   labs(x="eGFR",y="Adjusted hazard ratio (95% CI)",
#        linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
#   theme_bw() +
#   theme(text=element_text(size=14))

# sodium
sod_range = with(pred_data,seq(min(sodium_at_start),max(sodium_at_start),length.out=200))
sod_spl = ns(sod_range,df=3,
              knots=attr(mortality_model_full$model$`ns(sodium_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(sodium_at_start, df = 3)`,"Boundary.knots"))
sod_rr = cbind.data.frame(
  sodium = rep(sod_range,2),
  durable_lvad = c(rep(0,length(sod_range)),rep(1,length(sod_range))),
  rbind.data.frame(
    do.call(rbind.data.frame,apply(sod_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("sodium",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})),
    do.call(rbind.data.frame,apply(sod_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("sodium",names(coef(mortality_model_full)))],
                  comb=row,#rep(row,2),
                  cov=mortality_model_full_VCOV)}))))
# ggplot(sod_rr %>% mutate(durable_lvad=factor(durable_lvad)),
#        aes(x=sodium,y=hr,linetype=durable_lvad,color=durable_lvad,fill=durable_lvad)) +
#   geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
#   geom_line(color="black",linewidth=.75) +
#   geom_hline(yintercept=1,linetype=2) +
#   # scale_y_continuous(trans='log2',minor_breaks=NULL,
#   #                    breaks=2^(-8:1),limits=c(.0005,2),
#   #                    oob = scales::rescale_none) +
#   # scale_x_continuous(breaks=seq(100,200,20),limits=c(100,200),oob = scales::rescale_none) +
#   labs(x="Sodium",y="Adjusted hazard ratio (95% CI)",
#        linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
#   theme_bw() +
#   theme(text=element_text(size=14))

## albumin
alb_range = with(pred_data,seq(min(albumin_at_start),max(albumin_at_start),length.out=200))
alb_spl = ns(alb_range,df=3,
              knots=attr(mortality_model_full$model$`ns(albumin_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(albumin_at_start, df = 3)`,"Boundary.knots"))
alb_rr = cbind.data.frame(
  albumin = rep(alb_range,2),
  durable_lvad = c(rep(0,length(alb_range)),rep(1,length(alb_range))),
  rbind.data.frame(
    do.call(rbind.data.frame,apply(alb_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("albumin",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})),
    do.call(rbind.data.frame,apply(alb_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("albumin",names(coef(mortality_model_full)))],
                  comb=row,#rep(row,2),
                  cov=mortality_model_full_VCOV)}))))
# ggplot(alb_rr %>% mutate(durable_lvad=factor(durable_lvad)),
#        aes(x=albumin,y=hr,linetype=durable_lvad,color=durable_lvad,fill=durable_lvad)) +
#   geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
#   geom_line(color="black",linewidth=.75) +
#   geom_hline(yintercept=1,linetype=2) +
#   scale_y_continuous(trans='log2',minor_breaks=NULL,
#                      breaks=c(.125/2,.125,.25,.5,1,2),limits=c(.01,2),
#                      oob = scales::rescale_none) +
#   # scale_x_continuous(breaks=seq(100,200,20),limits=c(100,200),oob = scales::rescale_none) +
#   labs(x="Albumin",y="Adjusted hazard ratio (95% CI)",
#        linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
#   theme_bw() +
#   theme(text=element_text(size=14))

# BNP
bnp_range = with(pred_data %>% filter(BNP_NT_Pro_at_start==0),
                  seq(min(log(BNP_at_start+1)),max(log(BNP_at_start+1)),length.out=200))
bnp_nt_range = with(pred_data %>% filter(BNP_NT_Pro_at_start==1),
                  seq(min(log(BNP_at_start+1)),max(log(BNP_at_start+1)),length.out=200))
bnp_spl = ns(c(bnp_range,bnp_nt_range),df=3,
              knots=attr(mortality_model_full$model$`ns(log(BNP_at_start + 1), df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(log(BNP_at_start + 1), df = 3)`,"Boundary.knots"))
bnp_rr = cbind.data.frame(
  bnp_trans = rep(c(bnp_range,bnp_nt_range),2),
  bnp_nt_pro = rep(c(rep(0,length(bnp_range)),rep(1,length(bnp_nt_range))),2),
  durable_lvad = c(rep(0,length(c(bnp_range,bnp_nt_range))),rep(1,length(c(bnp_range,bnp_nt_range)))),
  rbind.data.frame(
    # BNP_NT_Pro=0, LVAD=0
    do.call(rbind.data.frame,apply(bnp_spl[1:200,],1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("BNP",names(coef(mortality_model_full))) & !grepl("BNP_NT",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})),
    # BNP_NT_Pro=1, LVAD=0
    do.call(rbind.data.frame,apply(bnp_spl[201:400,],1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("BNP",names(coef(mortality_model_full)))][1:7],
                  comb=c(row,1,row),
                  cov=mortality_model_full_VCOV)})),
    # BNP_NT_Pro=0, LVAD=1
    do.call(rbind.data.frame,apply(bnp_spl[1:200,],1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("BNP",names(coef(mortality_model_full))) & !grepl("BNP_NT",names(coef(mortality_model_full)))],
                  comb=rep(row,2),
                  cov=mortality_model_full_VCOV)})),
    # BNP_NT_Pro=1, LVAD=1
    do.call(rbind.data.frame,apply(bnp_spl[201:400,],1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("BNP",names(coef(mortality_model_full)))],
                  comb=rep(c(row,1,row),2),
                  cov=mortality_model_full_VCOV)}))
    )) %>%
  mutate(bnp=exp(bnp_trans)-1)
ggplot(bnp_rr %>% mutate(durable_lvad=factor(durable_lvad)),
       aes(x=bnp_trans,y=hr,linetype=durable_lvad,color=durable_lvad,fill=durable_lvad)) +
  geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
  geom_line(color="black",linewidth=.75) +
  geom_hline(yintercept=1,linetype=2) +
  scale_y_continuous(trans='log2',minor_breaks=NULL,
                     # breaks=c(.25,.5,1,2,4,8),
                     #limits=c(.3,200),
                     oob = scales::rescale_none) +
  facet_grid(~bnp_nt_pro,scales="free_x") +
  # scale_x_continuous(breaks=seq(0,1,.2),limits=c(0,1),oob = scales::rescale_none) +
  labs(x="BNP",y="Adjusted hazard ratio (95% CI)",
       linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
  theme_bw() +
  theme(text=element_text(size=14))

# plot them all together
ggplot(bind_rows(bili_rr %>% mutate(value=bilirubin_trans,lab="ln(Bilirubin)"),
                 egfr_rr %>% mutate(value=egfr,lab="eGFR"),
                 bnp_rr %>% mutate(value=bnp_trans,lab=case_when(
                   bnp_nt_pro==0 ~ "ln(BNP)",T ~ "ln(BNP NT Pro)")),
                 alb_rr %>% mutate(value=albumin,lab="Albumin"),
                 sod_rr %>% mutate(value=sodium,lab="Sodium")) %>%
         mutate(durable_lvad=factor(durable_lvad,levels=c(0,1),labels=c("No","Yes"))),
       aes(x=value,y=hr,linetype=durable_lvad,fill=durable_lvad)) +
  geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
  geom_line(color="black",linewidth=.75) +
  geom_hline(yintercept=1,linetype=2) +
  scale_y_continuous(trans='log2',minor_breaks=NULL,
                     breaks=2^((-3):6),limits=c(.125/2,32),
                     oob = scales::rescale_none) +
  facet_wrap(~lab,scales="free_x",nrow=2) +
  scale_fill_viridis_d() +
  # scale_x_continuous(breaks=seq(0,1,.2),limits=c(0,1),oob = scales::rescale_none) +
  labs(x="Lab value",y="Adjusted hazard ratio (95% CI)",
       linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
  theme_bw() +
  theme(text=element_text(size=14))
ggsave("lab-coefs-by-lvad-NEW.png",width=7.5,height=5,dpi=300)




```


## Summarize hemodynamic effects
```{r}

## api
api_range = with(pred_data,seq(min(api_at_start),max(api_at_start),length.out=200))
api_spl = ns(api_range,df=3,
              knots=attr(mortality_model_full$model$`ns(api_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(api_at_start, df = 3)`,"Boundary.knots"))
api_rr = cbind.data.frame(
  api = api_range,
  do.call(rbind.data.frame,apply(api_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("api",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})))

## papi
papi_range = with(pred_data,seq(min(papi_at_start),quantile(papi_at_start,.95),length.out=200))
papi_spl = ns(papi_range,df=3,
              knots=attr(mortality_model_full$model$`ns(papi_at_start, df = 3)`,"knots"),
              Boundary.knots=attr(mortality_model_full$model$`ns(papi_at_start, df = 3)`,"Boundary.knots"))
papi_rr = cbind.data.frame(
  papi = papi_range,
  do.call(rbind.data.frame,apply(papi_spl,1,function(row) {
                getHR(mortality_model_full,
                  vars=names(coef(mortality_model_full))[grepl("papi",names(coef(mortality_model_full)))][1:3],
                  comb=row,
                  cov=mortality_model_full_VCOV)})))

## cpo
# cpo_range = with(pred_data,seq(min(cpo_at_start),quantile(cpo_at_start,.99),length.out=200))
# cpo_spl = ns(cpo_range,df=3,
#               knots=attr(mortality_model_full$model$`ns(cpo_at_start, df = 3)`,"knots"),
#               Boundary.knots=attr(mortality_model_full$model$`ns(cpo_at_start, df = 3)`,"Boundary.knots"))
# cpo_rr = cbind.data.frame(
#   cpo = cpo_range,
#   do.call(rbind.data.frame,apply(cpo_spl,1,function(row) {
#                 getHR(mortality_model_full,
#                   vars=names(coef(mortality_model_full))[grepl("cpo",names(coef(mortality_model_full)))][1:3],
#                   comb=row,
#                   cov=mortality_model_full_VCOV)})))

# make the plot
ggplot(bind_rows(api_rr %>% rename(value=api) %>% mutate(hemo="API"),
                 papi_rr %>% rename(value=papi) %>% mutate(hemo="PAPi")) %>%
         mutate(hemo=factor(hemo,levels=c("API","PAPi"))),
       aes(x=value,y=hr)) +
  geom_ribbon(aes(ymin = lowCI, ymax = highCI),alpha=.5) +
  geom_line(color="black",linewidth=.75) +
  geom_hline(yintercept=1,linetype=2) +
  scale_y_continuous(trans='log2',minor_breaks=NULL,
                     breaks=c(.25,.5,1,1.5,2,4,8),
                     limits=c(.25,2),
                     oob = scales::rescale_none) +
  facet_grid(~hemo,scales="free") +
  # scale_x_continuous(breaks=seq(0,1,.2),limits=c(0,1),oob = scales::rescale_none) +
  labs(x="Pulsatility index value",y="Adjusted hazard ratio (95% CI)") +
       # linetype="Durable LVAD",color="Durable LVAD",fill="Durable LVAD") +
  theme_bw() +
  theme(text=element_text(size=14))
ggsave("hemo-coefs-NEW.png",width=6,height=4)
```


## Forest plot
```{r,eval=F}
# Extract coefficients and confidence intervals
coef1 <- summary(mortality_model_no_weights)$coefficients
coef2 <- summary(uscrs_model)$coefficients


# Create data frames with the coefficients and confidence intervals
df1 <- data.frame(
  term = rownames(coef1),
  estimate = coef1[, "Estimate"],
  conf.low = coef1[, "Estimate"] - 1.96 * coef1[, "Std. Error"],
  conf.high = coef1[, "Estimate"] + 1.96 * coef1[, "Std. Error"],
  model = "Model without Weights"
)

# Same terms as US-CRS + hemodynamics
df2 <- data.frame(
  term = setdiff(rownames(coef1), c("papi_at_start", "cpo_at_start", "api_at_start")),
  estimate = coef2[, "Estimate"],
  conf.low = coef2[, "Estimate"] - 1.96 * coef2[, "Std. Error"],
  conf.high = coef2[, "Estimate"] + 1.96 * coef2[, "Std. Error"],
  model = "USCRS 1.0"
)

df1 <- df1[df1$term != "(Intercept)", ]
df2 <- df2[df2$term != "(Intercept)", ]

# Combine the two data frames
forest_df <- rbind(df1, df2)

coef2
df2
forest_df
class(df_intervals$BNP_NT_Pro)
setdiff(rownames(coef1), c("papi_at_start", "cpo_at_start", "api_at_start"))

# Rename and reorder variables
new_names <- c("sodium_at_start" = "Sodium", "short_mcs_ever_at_start" = "Short-term MCS", "papi_at_start" = "PAPI", "log(BNP_at_start + 1):BNP_NT_Pro_at_start0" = "Log BNP, Regular BNP", "log(BNP_at_start + 1):BNP_NT_Pro_at_start1" = "Log BNP, NT-proBNP", "log(bilirubin_at_start + 1)" = "Log Bilirubin", "eGFR_at_start" = "eGFR", "durable_LVAD_at_start" = "Durable LVAD", "cpo_at_start" = "CPO", "api_at_start" = "API", "albumin_at_start" = "Albumin" )
term_order <- c("PAPI", "API", "CPO", "Durable LVAD", "Sodium", "Albumin", "Log BNP, NT-proBNP", "Log BNP, Regular BNP", "eGFR", "Log Bilirubin", "Short-term MCS")
forest_df

forest_df <- forest_df %>%
  mutate(term = recode(term, !!!new_names))
forest_df <- forest_df %>%
  mutate(term = factor(term, levels = term_order))# Create the forest plot
ggplot(forest_df, aes(x = estimate, y = term, color = model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, position = position_dodge(width = 0.5)) +
  labs(x = "Coefficient Estimate", y = "Term", title = "Model Coefficients") +
  theme_minimal()

new_names
```

